---
title: "MELTING POTES PROJECT : SDM-NEO MODELS Linear Quantile Regression"
author: "Amélie Lehuen"
date: "2022-12-14"
description: "ANALYSIS WITH KOENKER'S QUANTREG PACKAGE"
toc: true # table of content at beginning of document
number-sections: true
highlight-style: pygments
format: 
  html: 
    self-contained: true
    code-fold: false
    html-math-method: katex # displays equations
  pdf:
    code-line-numbers: true
    geometry: 
      - top=20mm
      - left=20mm
  docx:  
    reference-doc: default

bibliography: references.bib

editor: visual
execute:
  eval: true
  echo: false
  message: false
  warning: false
  output: true
  include: false
  cache: false #create cache of results of chunks
---

# LINEAR QUANTILE REGRESSION ANALYSIS WITH KOENKER'S QUANTREG PACKAGE

## Introduction

Linear quantile regression analysis with Koenker's `quantreg` package

## Script preparation

### Packages

```{r}
#| label: load-packages
#| code-summary: "Packages"

library(readxl) ; library(openxlsx); library(beepr) # Edition d'un fichier Excel
library(tidyverse); library(reshape2); library(rlist) # the one; melt; list.append
library(ggpubr); #library(GGally); # stat_compare_means ;
library(scales); library(RColorBrewer); library(wesanderson); library(grafify); library(colorspace); library(ggsci)# show_col and colors colors colors!
library(quantreg);# library(visreg)
library(plotly); library(plot3D);  # graphiques 3D plot 3D for mesh library(pracma)

# GIS Packages
library(sf); library(sfheaders); # st_as_sf ; sf_to_df
library(htmlwidgets) # library(leaflet) # saveWidget ; for interactive maps
```

### Working Environment

```{r}
#| label: workenvir

rm(list=ls())
# tsk <- "A_SDM_NEO/"
wdpath <- "../" #paste(pc,"Melting Pot/BDD/",tsk,sep="")
wdwork <- paste(wdpath,"Matrices/",sep="")
wdgraph <- paste(wdpath,"Graphiques/",sep="")
wdres <- paste(wdpath,"Resultats/",sep="")
# setwd("./")

pc <- "C:/Users/lehuen201/Nextcloud/" # "E:/" #
wdgraphEx<-wdgraph #paste(pc,"Copie-HD/Melting Potes/",tsk,"Graphiques/",sep="")
wdGIS <- paste(pc,"Melting Pot/SIG/",sep="")

etude <- "CSLN_Mars"
```

### Graphic charter

```{r}
#| label: graphchart

theme_set(theme_bw(base_size = 16)) # theme_gray() theme_bw() theme_light()

pal_cspx <- function(x) {divergingx_hcl(x,palette = "Zissou 1")}; # show_col(pal_cspx(6)) # colorspace
colRQ<-pal_cspx(4); colBin<-c(pal_cspx(6)[3],pal_cspx(6)[5]); blank<-alpha("#dae8ed",0.1); # show_col(colRQ)
Scalc_rq<- function() {scale_colour_manual(values=colRQ)}
Scalf_rq2d <- function() {scale_fill_gradientn(colours=colRQ)} # function(x) {scale_fill_material("teal")} # 

pal_ggscc <- pal_material("teal"); # show_col(pal_ggscc(9)) # ggsci
colSum <- c(pal_ggscc(9)[3],pal_ggscc(9)[7],pal_ggscc(9)[1]) # show_col(colSum)
colInliers<-pal_material("teal")(10)[5] ; colOutliers = pal_material("deep-orange")(10)[5] # show_col(colOutliers)
Scalc_misc <- function() {scale_color_discrete_diverging(palette = "Blue-Red 3")}
Scale_map<- function() {scale_fill_distiller(palette = "Spectral")}
```

### Home made functions

```{r}
#| label: functmade


```

## Load of External data and Basic Variables

### External data

```{r}
#| label: externdata

rdatain <- sprintf("%sCSLN_Mars_BDD.RData",wdwork)
load(rdatain)
```

### Output save binder and Rdata

```{r}
#| label: outfiles

binderout <- sprintf("%sCSLN_BDD.xlsx",wdres)
rdataout  <- sprintf("%s%s_RQ_BDD.RData",wdwork,etude)
# if exists
# load(rdataout)
```

### Basic variables

```{r}
#| label: basicvar

analysis <- "RQ Linear"

choixttt <- 1#23 # treatment choices 1: RQ 1 to 4 factors coeff and AICm; 2: RQ graphs on demand; 3 : Mars SDM calculation

# Choice of predictors used according to autocorrelation study, taus...
Var_choosen<-c("flow_mxd","inunt","sal_dtd","temp_m","mudrate_m","tenfon_mxd")
pred_red <- predict %>% 
  subset(Var %in% Var_choosen) %>% 
  arrange(match(Var,Var_choosen))
taus <- c(0.5,0.9,0.95,0.975) # c(0.5,0.85,0.9,0.95,0.975,0.99) #c(1:19/20,0.975,0.99)
spe <- 1#:nrow(speciesMP) # 1:CERED 2:CORVO 3:HEDDI 4:LIMBA 5:PERUL 6:SCRPL
reponse<-reponse %>% 
  filter(rvar %in% c("Biomass_gAFDWm2","Density_indm2"))
answ <- 1:nrow(reponse)
sai <- 1#:nrow(saison) # 1:Year 2:Winter 3:Summer
explo <- 1:nrow(pred_red) # Choice of the domain of exploration of the reduced predictors
graphfine<-50 # graph resolution for 2D and 3D graphs surfaces

```

# ONE to FOUR FACTOR LINEAR RQ with AICm and Rone calculation

A litterature ref is done by adding an "Insert anything" tool (CTRL+/) : "/Citation". Points to Zotero, looks like [@cozzoli2018] and creates the reference table. Can also inserted by the button insert.

You can add an equation @eq-one to illustrate the calculation : "/Inline Math" or "Display math"

$$
price = \hat{\beta}_0 + \hat{\beta}_1 \times area + \epsilon
$$ {#eq-one}

```{r}
#| label: calculation_1
#| cache: true

if (choixttt==1 | choixttt==12 | choixttt==13 | choixttt==123){
  modelq_list <- list()
  smrq_t_list <- list(); RqMod_list <- list(); # rqResid_t_list <- list()
  rqlim_t_list <- list(); rqinf_t_list <- list(); rqsup_t_list <- list();
  sm_plot_list <-list(); qqres_plot_list<-list(); dp_plot_list<-list()
  dp2dF_plot_list <- list(); dp2d_plot_list <- list()
  dp3d_plot_list <- list()
  dp3dS_plot_list <- list(); dp3dF_plot_list <- list()
  for (sp in spe) { # sp=1
    df <- CSLN_Mars[which(CSLN_Mars$SPCourt == speciesMP$SPCourt[sp]),]
    for (sa in sai) {# sa=1
      for (rep in answ){ # rep=1
        for (k in explo) { # k=1 

          # ONE FACTOR RQ ----
          # var def for interation
          yt = reponse[rep,1]
          yl = sprintf("%s (%s)",reponse[rep,2],reponse[rep,3])
          zt = "Zone"
          x1t = sprintf("%s%s",pred_red[k,1],saison[sa,1])
          x1l = sprintf("%s%s (%s)",pred_red[k,2],saison[sa,1],pred_red[k,3])
          dfrq <-df[,c(yt,x1t,zt)]
          dfrq <-as.data.frame(na.omit(dfrq)) %>% 
            rename(x1=!!x1t,y=!!yt,z=!!zt)
          y <- dfrq$y <- jitter(dfrq$y)
          x1 <- dfrq$x1 <- jitter(dfrq$x1)
          z <- dfrq$z
          
          x1mod <- seq(min(x1),max(x1),length.out=graphfine)
          x1Mod <- data.frame(x1mod) %>% rename(x1=x1mod)
          titleG <- sprintf("%s %s in %s",analysis,speciesMP$Taxon_SNa[sp],saison[sa,2])
          sdmname<-sprintf("RQ1_%g%g%g%g0",sp,sa,rep,k)
         
          # RQ calculation with error management ----
          tryCatch({
            modelq<-rq(y~x1, tau=taus);
            xt<-x1t; 
            xl<-x1l
            subtitleG <- sprintf("%s vs %s",yl,xl)
            modelq0<-rq(y ~ 1, tau=taus) # for Rone calculation # Often a warning "Solution may be nonunique"
            
            # Summary calculation and pimp
            smrq <- summary(modelq, se="boot") #, se="boot" or "boot" # avec boot, p value et std err
            smrq_t <- list(); 
            for (t in 1:length(taus)){
              smrq_t[[t]] <- as.data.frame(smrq[[t]][["coefficients"]]) %>%
                mutate(Var=rownames(.)) %>% relocate(Var) %>% remove_rownames(.) %>%
                mutate(AIC=round(AIC(modelq)[t],1),
                       Rone=round((1 - modelq$rho[t]/modelq0$rho[t]),5)) %>%
                mutate(tau=taus[t], taust=sprintf("tau= %s",tau),
                       sdmname=sdmname,
                       mode=paste("Linear","1var",sep=' '),
                       formula=format(modelq$formula),
                       type=as.character(modelq$formula[[1]]),
                       tau=taus[t],
                       Sp=speciesMP$SPCourt[sp],Season=saison[sa,2],
                       reponse=yl,reponset=yt,
                       Predictor1=pred_red[k,1],Predictor2="None",
                       predict=xl,predictt=xt)
            }
            
            # Out of limits points calculation
            rqlim_t <- as.data.frame(predict(modelq)) %>%
              rename_all(~sprintf("%.3f",taus)) %>% 
              mutate(across(everything(), 
                            function(x){replace(x, which(x<0), NA)})) %>% 
              bind_cols(tibble(x1=x1),tibble(y=y))
            rqlim_t <- melt(rqlim_t,id.vars =c("x1","y"),
                            variable.name = "tau",value.name = "RqLim") %>% 
              mutate(sdmname=sdmname,
                     taust=sprintf("tau= %s",tau),
                     tau=as.numeric(as.character(tau)))
            rqsup_t <- rqlim_t %>% 
              mutate(across(c(x1,y),
                            ~replace(.x, which(y<RqLim), NA))) %>% 
              drop_na()
            rqinf_t <- rqlim_t %>% 
              mutate(across(c(x1,y),
                            ~replace(.x, which(y>=RqLim), NA))) %>% 
              drop_na()
            
            # Surface model calculation
            RqMod_t <- as.data.frame(predict(modelq,newdata=x1Mod)) %>%
              rename_all(~sprintf("%.3f",taus)) %>% 
              mutate(across(everything(), 
                            function(x){replace(x, which(x<0), NA)})) %>% 
              bind_cols(x1Mod) 
            RqMod_t <- melt(RqMod_t,id.vars =c("x1"),
                            variable.name = "tau",
                            value.name = "RqMod") %>% 
              mutate(sdmname=sdmname,
                     taust=sprintf("tau= %s",tau),
                     tau=as.numeric(as.character(tau)))
            
            modelq_list[[sdmname]] <- modelq
            smrq_t_list[[sdmname]] <- smrq_t <- smrq_t %>% map_dfr(., ~.)
            RqMod_list[[sdmname]] <- RqMod_t
            rqlim_t_list[[sdmname]] <- rqlim_t
            rqsup_t_list[[sdmname]] <- rqsup_t
            rqinf_t_list[[sdmname]] <- rqinf_t
            
            # GRAPHIC PART ----
            # SUMMARY GRAPHS
            png(file=sprintf("%s%s/%s/RQ1/%s_%s_RQ1_sm_%s_%s_%s.tiff",
                             wdgraph,speciesMP$SPCourt[sp],
                             analysis,etude,speciesMP$SPCourt[sp],
                             saison[sa,2],yt,xt),
                width=3600, height=3600, res = 600)
            plot(smrq, main=c("Beta 0",x1l),cex=.7,pch=19,lcol=colSum[1],col=colSum[2:3])
            title(sub = titleG)
            dev.off()
            
            # QQPLOT RESIDUALS OF RQ MODEL
            rqResid <- melt(residuals(modelq)) %>% 
              rename(tau=Var2, residuals=value) #plot(density(rqResid$value)) #rqFit <- fitted(modelq)
            qqres_plot_list[[sdmname]] <- ggplot(rqResid, aes(sample=residuals, color=tau)) + 
              stat_qq() + stat_qq_line() + 
              facet_wrap(~tau) +
              Scalc_rq() +
              theme(legend.position="bottom")+
              labs(title=titleG,
                   subtitle=paste("Residuals for ",subtitleG,sep=""),
                   color="Quantile"); #qqres_plot_list[[sdmname]]
            ggsave(sprintf("%s%s/%s/RQ1/%s_%s_RQ1Resid_%s_%s_%s.tiff",
                           wdgraph,speciesMP$SPCourt[sp],
                           analysis,etude,speciesMP$SPCourt[sp],
                           saison[sa,2],yt,x1t),
                   plot = qqres_plot_list[[sdmname]], width = 12, height = 8, dpi=600)
            
            # RQ GRAPHS
            dp_plot_list[[sdmname]] <- ggplot(dfrq, aes(x=x1, y=y)) + 
              geom_point(aes(color=z)) +
              Scalc_misc() +
              theme(legend.position="bottom") +
              labs(title=titleG,
                   x=paste(x1l," (",pred_red[k,3],")",sep=""),
                   y=yl,color=zt)
            for (t in 1:length(taus)){ # required to have diff color for rq lines
              dp_plot_list[[sdmname]] <- dp_plot_list[[sdmname]] + 
                geom_quantile(quantiles = taus[t], 
                              formula=y ~ x,
                              colour = colRQ[t], size = 0.8, alpha = 0.8)+
                annotate("text",x=max(x1,na.rm=TRUE)*0.9,y=max(y)*((t/length(taus))*0.9),
                         colour=colRQ[t],size=4,fontface=2,hjust=0, label = paste(taus[t]))
            }
            ggsave(sprintf("%s%s/%s/RQ1/%s_%s_RQ1_%s_%s_%s.tiff",
                           wdgraph,speciesMP$SPCourt[sp],
                           analysis,etude,speciesMP$SPCourt[sp],
                           saison[sa,2],yt,x1t),
                   plot = dp_plot_list[[sdmname]], width = 8, height = 8, dpi=600)

            modelq<-NULL; smrq<-NULL;
          },error = function(e) {print(e)})#,finally = {})
          # ----
          
          # TWO FACTORS LOOP ----
          for (k2 in explo[-c(1:k)]) { #k2=2
            # var def for interation
            x2t = sprintf("%s%s",pred_red[k2,1],saison[sa,1])
            x2l = sprintf("%s%s (%s)",pred_red[k2,2],saison[sa,1],pred_red[k2,3])
            dfrq <-df[,c(yt,x1t,x2t)]
            dfrq <-as.data.frame(na.omit(dfrq)) %>% 
              rename(x1=!!x1t,x2=!!x2t,y=!!yt)
            y <- dfrq$y <- jitter(dfrq$y)
            x1 <- dfrq$x1 <- jitter(dfrq$x1)
            x2 <- dfrq$x2 <- jitter(dfrq$x2)
            z <- dfrq$z
            
            x2mod <- seq(min(x2),max(x2),length.out=graphfine)
            x1x2mod <- data.frame(cbind(x1mod,x2mod)) %>% 
              rename(x1=x1mod,x2=x2mod)
            gridx1x2 <- expand.grid(x1mod,x2mod) %>% 
              rename(x1=Var1,x2=Var2)
            gridx1x2mat <- mesh(x1mod,x2mod)
            titleG <- sprintf("%s %s in %s",analysis,speciesMP$Taxon_SNa[sp],saison[sa,2])
            
            # Two factors loop Addition----
            sdmname<-sprintf("RQ2add_%g%g%g%g%g",sp,sa,rep,k,k2)
            # RQ calculation with error management
            tryCatch({
              modelq<-rq(y~x1+x2, tau=taus); 
              xt<-paste(x1t,x2t,sep=paste(' ',modelq$formula[[3]][[1]],' ')); 
              xl<-paste(x1l,x2l,sep=paste(' ',modelq$formula[[3]][[1]],' '))
              subtitleG <- sprintf("%s vs %s",yl,xl)
              modelq0<-rq(y ~ 1, tau=taus)

              # Summary calculation and pimp
              smrq <- summary(modelq, se="boot") 
              smrq_t <- list(); 
              for (t in 1:length(taus)){
               smrq_t[[t]] <- as.data.frame(smrq[[t]][["coefficients"]]) %>%
                  mutate(Var=rownames(.)) %>% 
                 relocate(Var) %>% 
                 remove_rownames(.) %>%
                  mutate(AIC=round(AIC(modelq)[t],1),
                         Rone=round((1 - modelq$rho[t]/modelq0$rho[t]),5)) %>%
                  mutate(tau=taus[t], 
                         taust=sprintf("tau= %s",tau),
                         sdmname=sdmname,
                         mode=paste("Linear",modelq$formula[[3]][[1]],"2var",sep=' '),
                         formula=format(modelq$formula),
                         type=as.character(modelq$formula[[3]][[1]]),
                         tau=taus[t],
                         Sp=speciesMP$SPCourt[sp],Season=saison[sa,2],
                         reponse=yl,reponset=yt,
                         Predictor1=pred_red[k,1],Predictor2=pred_red[k2,1],
                         predict=xl,predictt=xt)
              }
              
              # Out of limits points calculation
              rqlim_t <- as.data.frame(predict(modelq)) %>%
                rename_all(~sprintf("%.3f",taus)) %>% 
                mutate(across(everything(), function(x){replace(x, which(x<0), NA)})) %>% 
                bind_cols(tibble(x1=x1),tibble(x2=x2),tibble(y=y))
              rqlim_t <- melt(rqlim_t,id.vars =c("x1","x2","y"),
                              variable.name = "tau",value.name = "RqLim") %>% 
                mutate(sdmname=sdmname,
                       taust=sprintf("tau= %s",tau),
                       tau=as.numeric(as.character(tau)))
              rqsup_t <- rqlim_t %>% 
                mutate(across(c(x1,x2,y),~replace(.x, which(y<RqLim), NA))) %>% 
                drop_na()
              rqinf_t <- rqlim_t %>% 
                mutate(across(c(x1,x2,y),~replace(.x, which(y>=RqLim), NA))) %>% 
                drop_na()
              
              # Surface model calculation
              RqMod_t <- as.data.frame(predict(modelq,newdata=gridx1x2)) %>%
                rename_all(~sprintf("%.3f",taus)) %>% 
                mutate(across(everything(), 
                              function(x){replace(x, which(x<0), NA)})) %>% 
                bind_cols(gridx1x2) 
              RqMod_t <- melt(RqMod_t,id.vars=c("x1","x2"),
                              variable.name = "tau",value.name = "RqMod") %>% 
                mutate(sdmname=sdmname,
                       taust=sprintf("tau= %s",tau),
                       tau=as.numeric(as.character(tau)))

              modelq_list[[sdmname]] <- modelq
              smrq_t_list[[sdmname]] <- smrq_t <- smrq_t %>% map_dfr(., ~.)
              RqMod_list[[sdmname]] <- RqMod_t
              rqlim_t_list[[sdmname]] <- rqlim_t
              rqsup_t_list[[sdmname]] <- rqsup_t
              rqinf_t_list[[sdmname]] <- rqinf_t

              # GRAPHIC PART ----
              # SUMMARY GRAPHS
              png(file=sprintf("%s%s/%s/RQ2add/%s_%s_RQ2add_sm_%s_%s_%s_%s.tiff",
                               wdgraph,speciesMP$SPCourt[sp],
                               analysis,etude,speciesMP$SPCourt[sp],
                               saison[sa,2],yt,x1t,x2t),
                  width=3600, height=3600, res = 600)
              plot(smrq, main=c("Beta 0",x1l,x2l),
                   cex=.7,pch=19,lcol=colSum[1],col=colSum[2:3])
              title(sub = titleG)
              dev.off()
              # QQPLOT RESIDUALS OF RQ MODEL
              rqResid <- melt(residuals(modelq)) %>% 
                rename(tau=Var2, residuals=value) 
              
              qqres_plot_list[[sdmname]] <- ggplot(rqResid, aes(sample=residuals, color=tau)) + 
                stat_qq() + stat_qq_line() + 
                facet_wrap(~tau) +
                Scalc_rq() +
                theme(legend.position="bottom")+
                labs(title=titleG,
                     subtitle=paste("Residuals for ",subtitleG,sep=""),
                     color="Quantile")
              ggsave(sprintf("%s%s/%s/RQ2add/%s_%s_RQ2addResid_%s_%s_%s_%s.tiff",
                             wdgraph,speciesMP$SPCourt[sp],
                             analysis,etude,speciesMP$SPCourt[sp],
                             saison[sa,2],yt,x1t,x2t),
                     plot = qqres_plot_list[[sdmname]], width = 12, height = 8, dpi=600)
              
              # RQ GRAPHS
              # 2D static graphic : RASTER
              dp2dF_plot_list[[sdmname]] <- ggplot(RqMod_t) +
                geom_raster(aes(x = x1, y = x2, fill = RqMod),alpha=0.7) + 
                geom_point(data=rqinf_t,aes(x = x1, y = x2, alpha=y), 
                           shape = 4, size = 1, color = colInliers, na.rm = TRUE) +
                geom_point(data=rqsup_t,aes(x = x1, y = x2, alpha=y), 
                           shape = 5, size = 1, color = colOutliers, na.rm = TRUE) +
                labs(title=titleG,x=x1l, y=x2l,
                     fill = paste("SDM-NEO\n",reponse[rep,3],sep="")) +
                guides(alpha = "none") +
                Scalf_rq2d() +
                facet_wrap(taust ~ .); print(dp2dF_plot_list[[sdmname]])
              ggsave(sprintf("%s%s/%s/RQ2add/%s_%s_RQ2add2d_%s_%s_%s_%s.tiff",
                             wdgraph,speciesMP$SPCourt[sp],
                             analysis,etude,speciesMP$SPCourt[sp],
                             saison[sa,2],yt,x1t,x2t),
                     plot = dp2d_plot_list[[sdmname]], width = 10, height = 7, dpi=600)
              
              # Experimental points on 3D graphic
              myscene<-list(camera = list(eye = list(x = -1.5, y = 1.5, z = 0.3)),
                            aspectmode='cube', # define standard layout scene
                            xaxis = list(title = x1l), 
                            yaxis = list(title = x2l), 
                            zaxis = list(title = yl))
              
              dp3dS_plot_list[[sdmname]] <- plot_ly(showlegend=FALSE) %>% 
                add_trace(data=dfrq, x = x1, y = x2, z = y,
                          mode = "markers", type = "scatter3d",
                          marker = list(size = .5, color = "blue", 
                                        opacity = 0.5, symbol = 104)) %>%
                layout(title = titleG, scene = myscene)
              
              dp2d_plot_t <- list(); dp3d_plot_t <- list()
              # Define surfaces for each tau
              for (t in 1:length(taus)){ # t=4
                RqMod_mat <- RqMod_t %>% 
                  filter(tau==taus[t])
                RqMod_mat <- array(RqMod_mat$RqMod,dim=c(graphfine,graphfine))
                rqlim <- rqlim_t %>% 
                  filter(tau==taus[t]) %>% 
                  select(RqLim)
                rqsup<-rqsup_t %>% 
                  filter(tau==taus[t])
                rqinf<-rqinf_t %>% 
                  filter(tau==taus[t])
                
                dp2d_plot_t[[t]] <- ggplot(RqMod_t %>% 
                  filter(tau==taus[t])) +
                  geom_raster(aes(x = x1, y = x2, fill = RqMod), alpha=0.7) + 
                  geom_point(data=rqinf_t,aes(x=x1, y=x2), 
                             shape=21, size=1, color=colInliers, 
                             fill=colInliers, alpha=.5, na.rm = TRUE) +
                  geom_point(data=rqsup_t,aes(x=x1, y=x2), 
                             shape=23, size=1, color=colOutliers, 
                             fill=colOutliers, alpha=.5, na.rm = TRUE) +
                  labs(x=x1l, y=x2l,
                       fill = paste("SDM-NEO\n",reponse[rep,3],sep="")) +
                  guides(alpha = "none") +
                  Scalf_rq2d()#; print(dp2dF_plot_list[[sdmname]])

                dp3d_plot_t[[t]] <- plot_ly(showlegend=F, scene=paste("scene",t,sep="")) %>% 
                  add_surface(x = gridx1x2mat[["x"]], 
                              y = gridx1x2mat[["y"]], 
                              z = RqMod_mat,
                              opacity = 0.9, colorscale = list(c(0,1),c(blank,colRQ[t])),
                              colorbar=list(title=list(
                                  text=paste("SDM-NEO\n",
                                             reponse[rep,3],"\nTau=",taus[t],sep="")))) %>%
                  add_trace(x = rqsup$x1, 
                            y = rqsup$x2, 
                            z = rqsup$y, 
                            mode = "markers", type = "scatter3d",
                            marker = list(size = 2, color = colOutliers, 
                                          opacity = 0.5, symbol = "diamond"))%>%
                  add_trace(x = rqinf$x1, 
                            y = rqinf$x2, 
                            z = rqinf$y, 
                            mode = "markers", type = "scatter3d",
                            marker = list(size = 2, color = colInliers, 
                                          opacity = 0.5, symbol = "circle")) %>%
                  layout(title = titleG, scene = myscene); dp3d_plot_t[[t]]

                
                dp3dS_plot_list[[sdmname]] <- dp3dS_plot_list[[sdmname]] %>% 
                  add_surface(x = gridx1x2mat[["x"]], 
                              y = gridx1x2mat[["y"]], 
                              z = RqMod_mat,
                              opacity = 0.8, colorscale = list(c(0,1),c(blank,colRQ[t])))
              } # taus
              dp2d_plot_list[[sdmname]] <- dp2d_plot_t
              dp3d_plot_list[[sdmname]] <- dp3d_plot_t

              # scene <- list.append(myscene,domain=list(x=c(0,0.33),y=c(0,0.5)))
              # scene2 <- list.append(myscene,domain=list(x=c(0.34,0.66),y=c(0,0.5)))
              # scene3 <- list.append(myscene,domain=list(x=c(0.67,1),y=c(0,0.5)))
              # scene4 <- list.append(myscene,domain=list(x=c(0,0.33),y=c(0.5,1)))
              # scene5 <- list.append(myscene,domain=list(x=c(0.34,0.66),y=c(0.5,1)))
              # scene6 <- list.append(myscene,domain=list(x=c(0.67,1),y=c(0.5,1)))
              # dp3dF_plot_list <- subplot(dp3dF_plot_list, shareX = TRUE, shareY = TRUE, nrows = 2) %>%
              #   layout(scene = scene, scene2 = scene2, scene3 = scene3, 
              #          scene4 = scene4, scene5 = scene5, scene6 = scene6)
              # saveWidget(plot_dp3dF,sprintf("%s%s/%s/RQ2add/%s_%s_RQ2add3dFacet_%s_%s_%s_%s.html",
              #                          wdgraphEx,speciesMP$SPCourt[sp],
              #                          analysis,etude,speciesMP$SPCourt[sp],
              #                          saison[sa,2],yt,x1t,x2t),
              #            selfcontained = F, libdir = "lib")
              
              # 3D plot with all taus models surface
              dp3dS_plot_list[[sdmname]] <- dp3dS_plot_list[[sdmname]] %>% 
                add_trace(x = rqsup$x1, 
                          y = rqsup$x2, 
                          z = rqsup$y, 
                          mode = "markers", type = "scatter3d",
                          marker = list(size = 2, color = colOutliers, 
                                        opacity = 0.5, symbol = "diamond")) %>%
                add_trace(x = rqinf$x1, 
                          y = rqinf$x2, 
                          z = rqinf$y, 
                          mode = "markers", type = "scatter3d",
                          marker = list(size = 2, color = colInliers, 
                                        opacity = 0.5, symbol = "circle")) %>%
                hide_colorbar()
              saveWidget(dp3dS_plot_list[[sdmname]],
                         sprintf("%s%s/%s/RQ2add/%s_%s_RQ2add3dStack_%s_%s_%s_%s.html",
                                       wdgraphEx,speciesMP$SPCourt[sp],
                                       analysis,etude,speciesMP$SPCourt[sp],
                                       saison[sa,2],yt,x1t,x2t),
                         selfcontained = F, libdir = "lib")

              smrq<-NULL; modelq<-NULL        
            },error = function(e) {print(e)})#,finally = {})
            
            # Two factors loop Interaction----
            sdmname<-sprintf("RQ2int_%g%g%g%g%g",sp,sa,rep,k,k2)
            # RQ calculation with error management
            tryCatch({
              modelq<-rq(y~x1*x2, tau=taus); 
              xt<-paste(x1t,x2t,sep=paste(' ',modelq$formula[[3]][[1]],' ')); 
              xl<-paste(x1l,x2l,sep=paste(' ',modelq$formula[[3]][[1]],' '))
              subtitleG <- sprintf("%s vs %s",yl,xl)
              modelq0<-rq(y ~ 1, tau=taus)
              
              # Summary calculation and pimp
              smrq_t <- list(); 
              smrq <- summary(modelq, se="boot")
              for (t in 1:length(taus)){
                smrq_t[[t]] <- as.data.frame(smrq[[t]][["coefficients"]]) %>%
                  mutate(Var=rownames(.)) %>% relocate(Var) %>% remove_rownames(.) %>%
                  mutate(AIC=round(AIC(modelq)[t],1),
                         Rone=round((1 - modelq$rho[t]/modelq0$rho[t]),5)) %>%
                  mutate(tau=taus[t], 
                         taust=sprintf("tau= %s",tau),
                         sdmname=sdmname,
                         mode=paste("Linear",modelq$formula[[3]][[1]],"2var",sep=' '),
                         formula=format(modelq$formula),
                         type=as.character(modelq$formula[[3]][[1]]),
                         tau=taus[t],
                         Sp=speciesMP$SPCourt[sp],Season=saison[sa,2],
                         reponse=yl,reponset=yt,
                         Predictor1=pred_red[k,1],Predictor2=pred_red[k2,1],
                         predict=xl,predictt=xt)
              }

              # Out of limits points calculation
              rqlim_t <- as.data.frame(predict(modelq)) %>%
                rename_all(~sprintf("%.3f",taus)) %>% 
                mutate(across(everything(), 
                              function(x){replace(x, which(x<0), NA)})) %>% 
                bind_cols(tibble(x1=x1),tibble(x2=x2),tibble(y=y))
              rqlim_t <- melt(rqlim_t,id.vars =c("x1","x2","y"),
                              variable.name = "tau",value.name = "RqLim") %>% 
                mutate(sdmname=sdmname,
                       taust=sprintf("tau= %s",tau),
                       tau=as.numeric(as.character(tau)))
              rqsup_t <- rqlim_t %>% 
                mutate(across(c(x1,x2,y),
                              ~replace(.x, which(y<RqLim), NA))) %>% 
                drop_na()
              rqinf_t <- rqlim_t %>% 
                mutate(across(c(x1,x2,y),
                              ~replace(.x, which(y>=RqLim), NA))) %>% 
                drop_na()
              # Surface model calculation
              RqMod_t <- as.data.frame(predict(modelq,newdata=gridx1x2)) %>%
                rename_all(~sprintf("%.3f",taus)) %>% 
                mutate(across(everything(), 
                              function(x){replace(x, which(x<0), NA)})) %>% 
                bind_cols(gridx1x2) 
              RqMod_t <- melt(RqMod_t,id.vars =c("x1","x2"),
                              variable.name = "tau",value.name = "RqMod") %>% 
                mutate(sdmname=sdmname,
                       taust=sprintf("tau= %s",tau),
                       tau=as.numeric(as.character(tau)))

              modelq_list[[sdmname]] <- modelq
              smrq_t_list[[sdmname]] <- smrq_t <- smrq_t %>% map_dfr(., ~.)
              RqMod_list[[sdmname]] <- RqMod_t
              rqlim_t_list[[sdmname]] <- rqlim_t
              rqsup_t_list[[sdmname]] <- rqsup_t
              rqinf_t_list[[sdmname]] <- rqinf_t

              # GRAPHIC PART ----
              # SUMMARY GRAPHS 
              png(file=
                    sprintf("%s%s/%s/RQ2int/%s_%s_RQ2int_sm_%s_%s_%s_%s.tiff",
                            wdgraph,speciesMP$SPCourt[sp],
                            analysis,etude,speciesMP$SPCourt[sp],
                            saison[sa,2],yt,x1t,x2t),
                  width=4800, height=3600, res = 600)
              plot(smrq, main=c("Beta 0",x1l,x2l,xl),cex=.7,pch=19,
                   lcol=colSum[1],col=colSum[2:3])
              title(sub = titleG)
              dev.off()
              # QQPLOT RESIDUALS OF RQ MODEL
              rqResid <- melt(residuals(modelq)) %>% 
                rename(tau=Var2, residuals=value) ##plot(density(rqResid$value)) #rqFit <- fitted(modelq)
              qqres_plot_list[[sdmname]] <- ggplot(rqResid, aes(sample=residuals, color=tau)) + 
                stat_qq() + stat_qq_line() + 
                facet_wrap(~tau) +
                Scalc_rq() +
                theme(legend.position="bottom")+
                labs(title=titleG,
                     subtitle=paste("Residuals for ",subtitleG,sep=""),
                     color="Quantile"); qqres_plot_list[[sdmname]]
              ggsave(sprintf("%s%s/%s/RQ2int/%s_%s_RQ2intResid_%s_%s_%s_%s.tiff",
                             wdgraph,speciesMP$SPCourt[sp],
                             analysis,etude,speciesMP$SPCourt[sp],
                             saison[sa,2],yt,x1t,x2t),
                     plot = qqres_plot_list[[sdmname]], width = 12, height = 8, dpi=600)
              
              # RQ GRAPHS
              # 2D static graphic : RASTER
              dp2d_plot_list[[sdmname]] <- ggplot(RqMod_t) +
                geom_raster(aes(x = x1, y = x2, fill = RqMod),alpha=0.8) + 
                geom_point(data=rqinf_t,aes(x=x1, y=x2), 
                           shape=21, size=1, color=colInliers, 
                           fill=colInliers, alpha=.5, na.rm = TRUE) +
                geom_point(data=rqsup_t,aes(x=x1, y=x2), 
                           shape=23, size=1, color=colOutliers, 
                           fill=colOutliers, alpha=.5, na.rm = TRUE) +
                labs(title=titleG,x=x1l, y=x2l,fill = 
                       paste("SDM-NEO\n",reponse[rep,3],sep="")) +
                guides(alpha = "none") +
                Scalf_rq2d() +
                facet_wrap(taust ~ .); print(dp2d_plot_list[[sdmname]])
              ggsave(sprintf("%s%s/%s/RQ2int/%s_%s_RQ2int2d_%s_%s_%s_%s.tiff",
                             wdgraph,speciesMP$SPCourt[sp],
                             analysis,etude,speciesMP$SPCourt[sp],
                             saison[sa,2],yt,x1t,x2t),
                     plot = dp2d_plot_list[[sdmname]], width = 10, height = 8, dpi=600)
              
              # Experimental points on 3D graphic
              # define standard layout scene
              myscene<-list(camera = list(
                eye = list(x = -1.5, y = 1.5, z = 0.3)), 
                aspectmode='cube', # define standard layout scene
                xaxis = list(title = x1l), 
                yaxis = list(title = x2l), 
                zaxis = list(title = yl))
              
              dp3dS_plot_list[[sdmname]] <- plot_ly(showlegend=FALSE) %>%
                add_trace(data=dfrq, x = x1, y = x2, z = y,
                          mode = "markers", type = "scatter3d",
                          marker = list(size = .5, 
                                        color = "blue", 
                                        opacity = 0.5, 
                                        symbol = 104)) %>%
                layout(title = titleG, scene = myscene)
              
              # Define surfaces for each tau
              dp2d_plot_t <- list(); dp3d_plot_t <- list()
              for (t in 1:length(taus)){ # t=6 1:length(taus)
                RqMod_mat <- RqMod_t %>% 
                  filter(tau==taus[t])
                RqMod_mat <- array(RqMod_mat$RqMod,
                                   # dim=c(length(x1mod),length(x1mod)))
                                   dim=c(graphfine,graphfine))
                rqlim <- rqlim_t %>% 
                  filter(tau==taus[t]) %>% 
                  select(RqLim)
                rqsup<-rqsup_t %>% 
                  filter(tau==taus[t])
                rqinf<-rqinf_t %>% 
                  filter(tau==taus[t])
                
                dp2d_plot_t[[t]] <- ggplot(RqMod_t %>% 
                  filter(tau==taus[t])) +
                  geom_raster(aes(x = x1, y = x2, fill = RqMod), alpha=0.7) + 
                  geom_point(data=rqinf_t,aes(x=x1, y=x2), 
                             shape=21, size=1, color=colInliers, 
                             fill=colInliers, alpha=.5, na.rm = TRUE) +
                  geom_point(data=rqsup_t,aes(x=x1, y=x2), 
                             shape=23, size=1, color=colOutliers, 
                             fill=colOutliers, alpha=.5, na.rm = TRUE) +
                  labs(x=x1l, y=x2l,
                       fill = paste("SDM-NEO\n",reponse[rep,3],sep="")) +
                  guides(alpha = "none") +
                  Scalf_rq2d()#; print(dp2dF_plot_list[[sdmname]])

                dp3d_plot_t[[t]] <- plot_ly(showlegend=F, 
                                     scene=paste("scene",t,sep="")) %>% 
                  add_surface(x = gridx1x2mat[["x"]], 
                              y = gridx1x2mat[["y"]], 
                              z = RqMod_mat,
                              opacity = 0.9, 
                              colorscale = list(c(0,1),c(blank,colRQ[t])),
                              colorbar=list(title=list(
                                text=paste("SDM-NEO\n",
                                           reponse[rep,3],"\nTau=",taus[t],sep="")))) %>%
                  add_trace(x = rqsup$x1, 
                            y = rqsup$x2, 
                            z = rqsup$y, 
                            mode = "markers", type = "scatter3d",
                            marker = list(size = 2, 
                                          color = colOutliers, 
                                          opacity = 0.5, 
                                          symbol = "diamond")) %>%
                  add_trace(x = rqinf$x1, 
                            y = rqinf$x2, 
                            z = rqinf$y, 
                            mode = "markers", 
                            type = "scatter3d",
                            marker = list(size = 2, 
                                          color = colInliers, 
                                          opacity = 0.5, 
                                          symbol = "circle")) %>%
                  layout(title = titleG, scene = myscene)#; print(dp3d_plot_t[[t]])

                dp3dS_plot_list[[sdmname]] <- dp3dS_plot_list[[sdmname]] %>% 
                  add_surface(x = gridx1x2mat[["x"]], 
                              y = gridx1x2mat[["y"]], 
                              z = RqMod_mat,
                              opacity = 0.8, 
                              colorscale = list(c(0,1),c(blank,colRQ[t])))
              } # taus
              dp2d_plot_list[[sdmname]] <- dp2d_plot_t
              dp3d_plot_list[[sdmname]] <- dp3d_plot_t

              # scene <- list.append(myscene,domain=list(x=c(0,0.33),y=c(0,0.5)))
              # scene2 <- list.append(myscene,domain=list(x=c(0.34,0.66),y=c(0,0.5)))
              # scene3 <- list.append(myscene,domain=list(x=c(0.67,1),y=c(0,0.5)))
              # scene4 <- list.append(myscene,domain=list(x=c(0,0.33),y=c(0.5,1)))
              # scene5 <- list.append(myscene,domain=list(x=c(0.34,0.66),y=c(0.5,1)))
              # scene6 <- list.append(myscene,domain=list(x=c(0.67,1),y=c(0.5,1)))
              # plot_dp3dF <- subplot(dp3d_plot_t, shareX = TRUE, shareY = TRUE, nrows = 2) %>%
              #   layout(scene = scene, scene2 = scene2, scene3 = scene3, 
              #          scene4 = scene4, scene5 = scene5, scene6 = scene6)
              # saveWidget(plot_dp3dF,
              #            sprintf("%s%s/%s/RQ2int/%s_%s_RQ2int3dFacet_%s_%s_%s_%s.html",
              #                          wdgraphEx,speciesMP$SPCourt[sp],
              #                          analysis,etude,speciesMP$SPCourt[sp],
              #                          saison[sa,2],yt,x1t,x2t),
              #            selfcontained = F, libdir = "lib")
              
              # 3D plot with all taus models surface
              dp3dS_plot_list[[sdmname]] <- dp3dS_plot_list[[sdmname]] %>%
                add_trace(x = rqsup$x1, y = rqsup$x2, z = rqsup$y,
                          mode = "markers", type = "scatter3d",
                          marker = list(size = 2, color = colOutliers,
                                        opacity = 0.5, symbol = "diamond")) %>%
                add_trace(x = rqinf$x1, y = rqinf$x2, z = rqinf$y,
                          mode = "markers", type = "scatter3d",
                          marker = list(size = 2, color = colInliers,
                                        opacity = 0.5, symbol = "circle")) %>%
                hide_colorbar()
              saveWidget(dp3dS_plot_list[[sdmname]],
                         sprintf("%s%s/%s/RQ2int/%s_%s_RQ2int3dStack_%s_%s_%s_%s.html",
                                       wdgraphEx,speciesMP$SPCourt[sp],
                                       analysis,etude,speciesMP$SPCourt[sp],
                                       saison[sa,2],yt,x1t,x2t),
                         selfcontained = F, libdir = "lib")

              smrq<-NULL; modelq<-NULL        
            },error = function(e) {print(e)})#,finally = {})
          } # two factors loop ----
          
          # THREE FACTORS LOOP ----
          for (k2 in explo[-c(1:k)]) {
            for (k3 in explo[-c(1:k2)]) { #k3=4
              x2t = sprintf("%s%s",pred_red[k2,1],saison[sa,1])
              x2l = sprintf("%s%s (%s)",pred_red[k2,2],saison[sa,1],pred_red[k2,3])
              x3t = sprintf("%s%s",pred_red[k3,1],saison[sa,1])
              x3l = sprintf("%s%s (%s)",pred_red[k3,2],saison[sa,1],pred_red[k3,3])
              dfrq <-df[,c(yt,x1t,x2t,x3t)]
              dfrq <-as.data.frame(na.omit(dfrq)) %>% rename(x1=!!x1t,x2=!!x2t,x3=!!x3t,y=!!yt)
              y <- dfrq$y <- jitter(dfrq$y)
              x1 <- dfrq$x1 <- jitter(dfrq$x1) 
              x2 <- dfrq$x2 <- jitter(dfrq$x2)
              x3 <- dfrq$x3 <- jitter(dfrq$x3)
              
              # Three factors loop Addition----
              sdmname<-sprintf("RQ3add_%g%g%g%g%g%g",sp,sa,rep,k,k2,k3)
              # RQ calculation with error management
              tryCatch({
                modelq<-rq(y~x1+x2+x3, tau=taus)
                modelq_list[[sdmname]] <- modelq
                xt<-paste(x1t,x2t,x3t,sep=paste(' ',modelq$formula[[3]][[1]],' ')); 
                xl<-paste(x1l,x2l,x3l,sep=paste(' ',modelq$formula[[3]][[1]],' '))
                subtitleG <- sprintf("%s vs %s",yl,xl)
                modelq0<-rq(y ~ 1, tau=taus)
                
                # Summary calculation and pimp
                smrq_t <- list(); 
                smrq <- summary(modelq, se="boot")
                for (t in 1:length(taus)){
                  smrq_t[[t]] <- as.data.frame(smrq[[t]][["coefficients"]]) %>%
                    mutate(Var=rownames(.)) %>% 
                    relocate(Var) %>% 
                    remove_rownames(.) %>%
                    mutate(AIC=round(AIC(modelq)[t],1),
                           Rone=round((1 - modelq$rho[t]/modelq0$rho[t]),5)) %>%
                    mutate(tau=taus[t], taust=sprintf("tau= %s",tau),
                           sdmname=sdmname,
                           mode=paste("Linear",modelq$formula[[3]][[1]],"3var",sep=' '),
                           formula=format(modelq$formula),
                           type=as.character(modelq$formula[[3]][[1]]),
                           tau=taus[t],
                           Sp=speciesMP$SPCourt[sp],Season=saison[sa,2],
                           reponse=yl,reponset=yt,
                           Predictor1=pred_red[k,1],Predictor2=pred_red[k2,1],
                           predict=xl,predictt=xt)
                }
                smrq_t_list[[sdmname]] <- smrq_t <- smrq_t %>% map_dfr(., ~.)
                
                smrq<-NULL; modelq<-NULL        
              },error = function(e) {print(e)})#,finally = {})
              
              # Three factors loop Interaction----
              sdmname<-sprintf("RQ3int_%g%g%g%g%g%g",sp,sa,rep,k,k2,k3)
              # RQ calculation with error management
              tryCatch({
                modelq<-rq(y~x1*x2*x3, tau=taus)
                modelq_list[[sdmname]] <- modelq
                xt<-paste(x1t,x2t,x3t,sep=paste(' ',modelq$formula[[3]][[1]],' ')); 
                xl<-paste(x1l,x2l,x3l,sep=paste(' ',modelq$formula[[3]][[1]],' '))
                subtitleG <- sprintf("%s vs %s",yl,xl)
                modelq0<-rq(y ~ 1, tau=taus)
                
                # Summary calculation and pimp
                smrq_t <- list(); 
                smrq <- summary(modelq, se="boot")
                for (t in 1:length(taus)){
                  smrq_t[[t]] <- as.data.frame(smrq[[t]][["coefficients"]]) %>%
                    mutate(Var=rownames(.)) %>% 
                    relocate(Var) %>% 
                    remove_rownames(.) %>%
                    mutate(AIC=round(AIC(modelq)[t],1),
                           Rone=round((1 - modelq$rho[t]/modelq0$rho[t]),5)) %>%
                    mutate(tau=taus[t], taust=sprintf("tau= %s",tau),
                           sdmname=sdmname,
                           mode=paste("Linear",modelq$formula[[3]][[1]],"3var",sep=' '),
                           formula=format(modelq$formula),
                           type=as.character(modelq$formula[[3]][[1]]),
                           tau=taus[t],
                           Sp=speciesMP$SPCourt[sp],Season=saison[sa,2],
                           reponse=yl,reponset=yt,
                           Predictor1=pred_red[k,1],Predictor2=pred_red[k2,1],
                           predict=xl,predictt=xt)
                }
                smrq_t_list[[sdmname]] <- smrq_t <- smrq_t %>% map_dfr(., ~.)
                
                smrq<-NULL; modelq<-NULL        
              },error = function(e) {print(e)})#,finally = {})
            }
          } # three factors loop ----
          
          # FOUR FACTORS LOOP ----
          for (k2 in explo[-c(1:k)]) {
            for (k3 in explo[-c(1:k2)]) {
              for (k4 in explo[-c(1:k3)]) { #k4=4
                x2t = sprintf("%s%s",pred_red[k2,1],saison[sa,1])
                x2l = sprintf("%s%s (%s)",pred_red[k2,2],saison[sa,1],pred_red[k2,3])
                x3t = sprintf("%s%s",pred_red[k3,1],saison[sa,1])
                x3l = sprintf("%s%s (%s)",pred_red[k3,2],saison[sa,1],pred_red[k3,3])
                x4t = sprintf("%s%s",pred_red[k4,1],saison[sa,1])
                x4l = sprintf("%s%s (%s)",pred_red[k4,2],saison[sa,1],pred_red[k4,3])
                dfrq <-df[,c(yt,x1t,x2t,x3t,x4t)]
                dfrq <-as.data.frame(na.omit(dfrq)) %>%
                  rename(x1=!!x1t,x2=!!x2t,x3=!!x3t,x4=!!x4t,y=!!yt)
                y <- dfrq$y <- jitter(dfrq$y)
                x1 <- dfrq$x1 <- jitter(dfrq$x1)
                x2 <- dfrq$x2 <- jitter(dfrq$x2)
                x3 <- dfrq$x3 <- jitter(dfrq$x3)
                x4 <- dfrq$x4 <- jitter(dfrq$x4)
                
                # Four factors loop Addition----
                sdmname<-sprintf("Rq4add_%g%g%g%g%g%g%g",sp,sa,rep,k,k2,k3,k4)
                # RQ calculation with error management
                tryCatch({
                  modelq<-rq(y~x1+x2+x3+x4, tau=taus)
                  modelq_list[[sdmname]] <- modelq
                  xt<-paste(x1t,x2t,x3t,x4t,sep=paste(' ',modelq$formula[[3]][[1]],' ')); 
                  xl<-paste(x1l,x2l,x3l,x4l,sep=paste(' ',modelq$formula[[3]][[1]],' '))
                  subtitleG <- sprintf("%s vs %s",yl,xl)
                  modelq0<-rq(y ~ 1, tau=taus)
                  
                  # Summary calculation and pimp
                  smrq_t <- list(); 
                  smrq <- summary(modelq, se="boot")
                  for (t in 1:length(taus)){
                    smrq_t[[t]] <- as.data.frame(smrq[[t]][["coefficients"]]) %>%
                      mutate(Var=rownames(.)) %>% relocate(Var) %>% remove_rownames(.) %>%
                      mutate(AIC=round(AIC(modelq)[t],1),
                             Rone=round((1 - modelq$rho[t]/modelq0$rho[t]),5)) %>%
                      mutate(tau=taus[t], taust=sprintf("tau= %s",tau),
                             sdmname=sdmname,
                             mode=paste("Linear",modelq$formula[[3]][[1]],"4var",sep=' '),
                             formula=format(modelq$formula),
                             type=as.character(modelq$formula[[3]][[1]]),
                             tau=taus[t],
                             Sp=speciesMP$SPCourt[sp],Season=saison[sa,2],
                             reponse=yl,reponset=yt,
                             Predictor1=pred_red[k,1],Predictor2=pred_red[k2,1],
                             predict=xl,predictt=xt)
                  }
                  smrq_t_list[[sdmname]] <- smrq_t <- smrq_t %>% map_dfr(., ~.)
                  
                  smrq<-NULL; modelq<-NULL
                },error = function(e) {print(e)})#,finally = {})
                
                # Four factors loop Interaction----
                sdmname<-sprintf("RQ4int_%g%g%g%g%g%g%g",sp,sa,rep,k,k2,k3,k4)
                # RQ calculation with error management
                tryCatch({
                  modelq<-rq(y~x1*x2*x3*x4, tau=taus)
                  modelq_list[[sdmname]] <- modelq
                  xt<-paste(x1t,x2t,x3t,x4t,sep=paste(' ',modelq$formula[[3]][[1]],' ')); 
                  xl<-paste(x1l,x2l,x3l,x4l,sep=paste(' ',modelq$formula[[3]][[1]],' '))
                  subtitleG <- sprintf("%s vs %s",yl,xl)
                  modelq0<-rq(y ~ 1, tau=taus)
                  
                  # Summary calculation and pimp
                  smrq_t <- list(); 
                  smrq <- summary(modelq, se="boot")
                  for (t in 1:length(taus)){
                    smrq_t[[t]] <- as.data.frame(smrq[[t]][["coefficients"]]) %>%
                      mutate(Var=rownames(.)) %>% relocate(Var) %>% remove_rownames(.) %>%
                      mutate(AIC=round(AIC(modelq)[t],1),
                             Rone=round((1 - modelq$rho[t]/modelq0$rho[t]),5)) %>%
                      mutate(tau=taus[t], taust=sprintf("tau= %s",tau),
                             sdmname=sdmname,
                             mode=paste("Linear",modelq$formula[[3]][[1]],"4var",sep=' '),
                             formula=format(modelq$formula),
                             type=as.character(modelq$formula[[3]][[1]]),
                             tau=taus[t],
                             Sp=speciesMP$SPCourt[sp],Season=saison[sa,2],
                             reponse=yl,reponset=yt,
                             Predictor1=pred_red[k,1],Predictor2=pred_red[k2,1],
                             predict=xl,predictt=xt)
                  }
                  smrq_t_list[[sdmname]] <- smrq_t <- smrq_t %>% map_dfr(., ~.)
                  
                  smrq<-NULL; modelq<-NULL
                },error = function(e) {print(e)})#,finally = {})
              }
            }
          } # four factors loop ----
        } # explo
      } # reponse
    } # saison
  } # speciesMP
  # ----
  
  smrq_l <- smrq_t_list %>% map_dfr(., ~.)
  sdmlist_name <- unique(smrq_l$sdmname)
  # Graphic representation of model performance
  dp_plot_list[[sdmname]]<-smrq_l %>% filter(!str_detect(mode, "3var")) %>% filter(!str_detect(mode, "4var")) %>%
    group_by(tau) %>%
    ggplot(aes(x=reorder(predict,-AIC),y=AIC,color = factor(tau),shape=reponse))+
    geom_point(alpha=0.8, size= 5) +
    scale_color_manual(values=colRQ) +
    labs(title="Quantile Regression AIC scores : Linear", x="Model",y="AIC",color="Quantile",shape="Biologic") +
    theme(axis.text = element_text(size=10,face="bold")) +
    coord_flip();dp_plot_list[[sdmname]]
  ggsave(sprintf("%s%s/%s/%s_%s_RQ_AIC_scores.tiff",
                 wdgraph,speciesMP$SPCourt[sp],analysis,etude,speciesMP$SPCourt[sp]), 
         plot = dp_plot_list[[sdmname]], width = 16, height = 8, dpi=600)

  # Sauvegarde des outputs ----
  rm(list=c("smrq_tmp","smrq","rqsup","rqinf","rqlim","RqMod_mat","RqMod_t",
            "modelq","modelq0","df","dfrq","tmp",
            "CSLN","CSLN_pur","Mars_csv","Mars_csv_sf"))
  rm(list=ls(pattern="scene"))
  save.image(rdataout)
} #end if choixttt 1 ----
```

# Final actions and save

Rdata are saved in `rdataout`. An excel file collects data in `r binderout` , with sheets for :

-   data : contains whatever

```{r}
#| label: finalsave_xls
wb <- loadWorkbook(binderout) # addWorksheet(wb, sheetName = "rql")
writeData(wb, sheet = "Rq_coeff", x = smrq_l, 
          startCol = 1, startRow = 1,withFilter = FALSE)
saveWorkbook(wb,file=binderout, overwrite = TRUE)

```

```{r}
#| label: finalsave_r

save.image(rdataout)
beepr::beep(2)
```


# References {.unnumbered}
