---
title: "My title"
author: "Amélie Lehuen"
date: "2022-12-14"
description: "add info on script"
toc: true # table of content at beginning of document
number-sections: true
highlight-style: pygments
format: 
  html: 
    self-contained: true
    code-fold: false
    html-math-method: katex # displays equations
  pdf:
    code-line-numbers: true
    geometry: 
      - top=20mm
      - left=20mm
  docx: 
    reference-doc: default

bibliography: references.bib

editor: visual
execute:
  eval: true
  echo: false
  message: false
  warning: false
  output: true
  include: true
  cache: false #create cache of results of chunks
---

# My title

## Introduction

Describe action of the script.

::: {.callout-important collapse="true"}
## To improve

Don't forget at some point to read about quarto projects https://quarto.org/docs/projects/quarto-projects.html.
:::

## Script preparation

### Packages

```{r}
#| label: load-packages
#| code-summary: "Packages"
#| include: false

library(readxl) ; library(openxlsx); 
library(beepr); library(clipr) # Edition d'un fichier Excel
library(tidyverse); #The toolbox indispensable 
library(broom); library(purrr) # tidy,glance,augment
library(scales) ; library(grafify) ; library(ggthemes) # show_col & viridis ; colorblind palettes ; scale_color_colorblind (8)
library(RColorBrewer) ; library(colorspace); library(ggsci) ; library(wesanderson) # Palettes; hcl_palettes; id; 
library(ggpubr);library(gridExtra) ; library(grid) # Mozaic of graphs tools
library(lubridate)
library(knitr)
```

### Working Environment

```{r}
#| label: workenvir
#| echo: true

rm(list=ls())
spec <- "CERED"
flume <- "Qfluff"
Run <- "Lehuen_2022"
wdtask <- "./DATA"
wdsource <- paste(wdtask,spec,"/",sep="") #,"/",Run
wdres <-  paste(wdtask,"Resultats/",sep="")
wdmat <-  paste(wdtask,"Matrices/",sep="")
wdgraph <-  paste(wdtask,"Graphiques/",sep="")
setwd("./")
```

### Graphic charter

```{r}
#| label: graphchart
theme_set(theme_bw(base_size = 16)) # theme_gray() theme_bw() theme_light() theme_dark()
pal_clb <- function(x) {graf_col_palette(palette = "fishy")(x)}; # show_col(pal_clb(6))
colBin<-c(pal_clb(6)[1],pal_clb(6)[2]); colMisc<-pal_clb(6)[6]# show_col(colBin)
Scalc_bin<- function() {scale_colour_manual(values=colBin)}
Scalc_lon<- function(x) {scale_colour_manual(values=pal_clb(x))}
# Scale_brew <- function() {scale_colour_brewer(palette="Spectral",aesthetics=c("colour","fill"))}
```

### Home made functions

```{r}
#| label: functmade
#| echo: true
scale <- function(x) {(x - max(x, na.rm = TRUE))} # (x - x[1])
```

## Load of External data and Basic Variables

### External data

"./" for files in same folder, "../" folder parent folder

```{r}
#| label: externdata
#| echo: true
binderin <- sprintf("./Model Base Script Data.xlsx")
data <- as.data.frame(read_excel(binderin,sheet = "Exp_Plan", na = ""))
rdatain <- sprintf("%sCSLN_Mars_BDD.RData",wdwork)
load(rdatain)
```

### Output binder and data

```{r}
#| label: outfiles
#| echo: true
binderout <- sprintf("binderout.xlsx")
rdataout  <- sprintf("dataout.xlsx")
# if exists
load(rdataout)
```

### Basic variables

```{r}
#| label: basicvar
#| echo: true
discarded_run <- c("")
nelrow<-2; nelcol<-1
```

# First block of action

A litterature ref is done by adding an "Insert anything" tool (CTRL+/) : "/Citation". Points to Zotero, looks like [@cozzoli2014], and creates the reference table. Can also inserted by the button insert.

You can add an equation @eq-one to illustrate the calculation : "/Inline Math" or "Display math"

$$
price = \hat{\beta}_0 + \hat{\beta}_1 \times area + \epsilon
$$ {#eq-one}

::: callout-note
Other way in Citedrive, or direct ref from Pubmed... to explore
:::

## The big calculations, loops or stuff

```{r}
#| label: calculation_1
#| include : false # necessary to avoid fluffy doc

1+1
```

## The results choices, and graphs or table out

### Refer to the @fig-scatterplot built to comment. You can add inline results, such as `r nrow(mpg)` observations in data.

```{r}
#| label: fig-scatterplot
#| fig-cap: "This is the legend."
#| fig-alt: "This is the figure description."
#| fig-width: 6
#| fig-height: 3.5

ggplot(mpg, aes(x = hwy, y = cty, color = cyl)) +
  geom_point(alpha = 0.5, size = 2) +
  scale_color_viridis_c() +
  theme_minimal()
```

### Multiple Figures

Two plots in the same @fig-mpg . In @fig-mpg-1 and @fig-mpg-2 side by side.

```{r}
#| label: fig-mpg
#| fig-cap: "Global title."
#| fig-subcap:
#|   - "sub title of graph1"
#|   - "sub title of graph2"
#| layout-ncol: 2
#| column: page

ggplot(mpg, aes(x = hwy, y = cty, color = cyl)) +
  geom_point(alpha = 0.5, size = 2) +
  scale_color_viridis_c() +
  theme_minimal()

ggplot(mpg, aes(x = hwy, y = cty, color = displ)) +
  geom_point(alpha = 0.5, size = 2) +
  scale_color_viridis_c(option = "E") +
  theme_minimal()
```

```{r}
#| label: calculation_2
#| echo: false
mean_cty <- round(mean(mpg$cty), 2)
mean_hwy <- round(mean(mpg$hwy), 2)
```

The average city mileage of the cars in our data is `r mean_cty` and the average highway mileage is `r mean_hwy`.

### Table

Put a nice table to sum up @tbl-stats

```{r}
#| label: tbl-stats
#| tbl-cap: "table legend"

mpg %>%
  summarise(
    `Median price` = median(hwy),
    `IQR price` = IQR(hwy),
    `Median area` = median(cty),
    `IQR area` = IQR(cty),
    `Correlation, r` = cor(hwy, cty)
    ) %>%
  kable(digits = c(0, 0, 0, 0, 2))
```

# Final actions and save

Rdata are saved in `rdataout`. An excel file collects data in `r binderout` , with sheets for :

-   data : contains whatever

```{r}
#| label: finalsave_xls
wb <- copyWorkbook(loadWorkbook(binderin))
if (!("AllMeas" %in% names(wb))) {addWorksheet(wb=wb, sheetName = "AllMeas")}
writeData(wb, sheet = "AllMeas", x = data, startCol = 1, startRow = 1,withFilter = FALSE)
saveWorkbook(wb,file=binderout, overwrite = TRUE)

```

```{r}
#| label: finalsave_r

save.image(file = rdataout)
beepr::beep(2)
```

# References {.unnumbered}
