---
title: "MELTING POTES"
author: "Amélie Lehuen"
date: "`r format(Sys.time(), '%B %Y')`"
description: "Task A : SDM-NEO Data analysis"
format: 
  html: 
    self-contained: true
    code-fold: false
    df-print: kable
    html-math-method: katex # displays equations
  docx: 
    reference-doc: customreference_nonb.docx
    
editor: visual
execute:
  eval: true
  echo: false
  message: false
  warning: false
  output: true
  include: false
  cache: false
---

```{r}
#| label: load-packages
#| code-summary: "Packages"

library(knitr)
library(readxl) ; library(openxlsx)
library(beepr); library(clipr)
library(tidyverse); library(data.table)
library(broom); library(purrr) # tidy,glance,augment
library(lubridate)
library(rstatix); library(Hmisc)  # corr and pvalue calculation

# SIG
library(sf); library(sfheaders)
library(rnaturalearth) # library(raster)
library(tmap)
library(tmaptools) # tmap_mode; for static and interactive maps
library(ggspatial)

# Graphics packages
library(scales) ; library(grafify)
library(RColorBrewer) ; library(wesanderson); library(colorspace)
library(ggpubr);library(GGally)#; library(patchwork)
library(metR)
library(rayshader)
```

```{r}
#| label: workenvir

rm(list=ls())
wdsource <- "Sources/Faune/CSLN/"
wdmat <- "Matrices/"
wdgraph <- "Graphiques/"
wdres <- "Resultats/"

pc <- "C:/Users/lehuen201/Nextcloud/" # "E:/" #
wdGIS <- paste(pc,"Melting Pot/SIG/",sep="");
wdscript <- paste(pc,"Melting Pot/BDD/Scripts/",sep="")
wdmsr <- (paste(wdscript,"MSR/MSR.R",sep=""))

```

```{r}
#| label: graphchart

theme_set(theme_bw(base_size = 14))
colDarj <- function(x) {wes_palette("Darjeeling2",x, type = "continuous")}
colZiss <- function(x) {wes_palette("Zissou1",x, type = "continuous")}
colSpec <- colorRampPalette(brewer.pal(8, "Spectral")); 
colDark <- colorRampPalette(brewer.pal(8, "Dark2"));
Scale_col <- function(x) {scale_colour_manual(values=colDarj(x))}
Scale_fill <- function(x) {scale_fill_manual(values=colDarj(x))}
Scale_brew <- function() {scale_colour_brewer(palette="Spectral",aesthetics=c("colour","fill"))}

pal_cspx <- function(x) {divergingx_hcl(x,palette = "Zissou 1")};
colBin<-c(pal_cspx(6)[3],pal_cspx(6)[5]); 
colRQ<-pal_cspx(4); blank<-alpha("#dae8ed",0.1); # show_col(colBin)
Scalc_rq<- function() {scale_colour_manual(values=colRQ)}
Scalf_rq2d <- function() {scale_fill_gradientn(colours=colRQ)} 

```

```{r}
#| label: func

loadRData <- function(fileName){
#loads an RData file, and returns it
    load(here::here(fileName))
    mget(ls()[ls() != "fileName"])
}

# ggpairs custom colors on corr
corr_col <- function(data, mapping, method="p", use="pairwise", ...){
  # grab data
  x <- eval_data_col(data, mapping$x)
  y <- eval_data_col(data, mapping$y)
  # calculate correlation
  corr <- cor(x, y, method=method, use=use)
  # calculate colour based on correlation value
  colFn <- colorRampPalette(c("blue", "white", "red"), interpolate ='spline')
  fill <- colFn(100)[findInterval(corr, seq(-1, 1, length=100))]
  ggally_text(
    label = as.character(round(corr, 2)),
    mapping = aes(),
    xP = 0.5, yP = 0.5,
    ...) + #  ggally_cor(data = data, mapping = mapping, ...) + 
    theme_void() +
    theme(panel.background = element_rect(fill=fill))

} #wrap(cor_func,method = 'spearman', symbol = "Corr:\n")

# Function for corr ad pvalue table x is a matrix containing the data
# method : correlation method. "pearson"" or "spearman"" is supported
# removeTriangle : remove upper or lower triangle
# results :  if "html" or "latex"
# the results will be displayed in html or latex format
corstars <-function(x, method=c("pearson", "spearman"), 
                    removeTriangle=c("upper", "lower"),
                     result=c("none", "html", "latex")){
    #Compute correlation matrix
    require(Hmisc)
    x <- as.matrix(x)
    correlation_matrix<-rcorr(x, type=method[1])
    R <- correlation_matrix$r # Matrix of correlation coeficients
    p <- correlation_matrix$P # Matrix of p-value 
    
    ## Define notions for significance levels; spacing is important.
    mystars <- ifelse(p < .0001, "****", 
                      ifelse(p < .001, "***", 
                             ifelse(p < .01, "**", 
                                    ifelse(p < .05, "*", " "))))
    
    ## trunctuate the correlation matrix to two decimal
    R <- format(round(cbind(rep(-1.11, ncol(x)), R), 2))[,-1]
    
    ## build a new matrix that includes the correlations with their apropriate stars
    Rnew <- matrix(paste(R, mystars, sep=""), ncol=ncol(x))
    diag(Rnew) <- paste(diag(R), " ", sep="")
    rownames(Rnew) <- colnames(x)
    colnames(Rnew) <- paste(colnames(x), "", sep="")
    
    ## remove upper/lower triangle of correlation matrix
    if(removeTriangle[1]=="upper"){
      Rnew <- as.matrix(Rnew)
      Rnew[upper.tri(Rnew, diag = TRUE)] <- ""
    }
    else if(removeTriangle[1]=="lower"){
      Rnew <- as.matrix(Rnew)
      Rnew[lower.tri(Rnew, diag = TRUE)] <- ""
    }
    
    ## remove last column and return the correlation matrix
    Rnew <- as.data.frame(Rnew)
    Rnew <- cbind(Rnew[1:length(Rnew)-1])
    if (result[1]=="none") return(Rnew)
    else{
      if(result[1]=="html") print(xtable(Rnew), type="html")
      else print(xtable(Rnew), type="latex") 
    }
} 
```

```{r}
#| label: externdata

rdatain  <- sprintf("%sCSLN_Mars_RQ_BDD.RData",wdmat)
# load(rdatain)
rqdata<-loadRData(rdatain)
rdatain  <- sprintf("%sCSLN_Mars_nlRQ_BDD.RData",wdmat)
# load(rdatain)
nlrqdata<-loadRData(rdatain)
rdatain  <- sprintf("%sCSLN_Mars_RQbs_BDD.RData",wdmat)
# load(rdatain)
rqbsdata<-loadRData(rdatain)
rdatain  <- sprintf("%sCSLN_Mars_BDD.RData",wdmat)
load(rdatain) #load(here::here(rdatain))

```

```{r}
#| label: outfiles
binderout <- sprintf("%sA_SDM_NEO_Results.xlsx",wdres)
rdataout  <- sprintf("%sA_SDM_NEO_Results.RData",wdmat)
# if exists
# load(rdataout)

```

```{r}
#| label: basicvar

etude <- "CSLN_Mars"
espece <- "CERED"
sp <- 1
sa <- 1
# rep <- 1
pred_red <- rqdata$pred_red
taus <- rqdata$taus

model_types<- data.frame(rbind(c(analysis="RQ Linear",code="RQ2int_",bdd="rqdata"),
                    c(analysis="RQ Nonlinear",code="nlRQ2_",bdd="nlrqdata"),
                    c(analysis="RQ BSpline",code="RQbs2_",bdd="rqbsdata")))

factors<-predict

```

# Materials and Methods

All data treatment has been conducted with `r R.version.string` except for MARS3D pre-treatment on Matlab 2019a. Significance levels are tagged for p \< .0001 with "\*\*\*\*", p \< .001 with "\*\*\*", p \< .01 with "\*\*", p \< .05 with "\*".

## Areas map

<!-- ::: {.content-visible unless-format="html"} -->

The Seine estuary has been studied regarding several historically known areas that have either habitat or communities differences. These areas are mainly mudflats and subtidal areas, are represented in @fig-map.

```{r}
#| label: fig-map
#| include: true
#| fig-cap: "Maps of habitats areas defined for the study"

# Base of bounding boxes
boxMP<-c(xmin=-0.1, ymin=49.3, xmax=0.45, ymax=49.65)
boxchannel<-c(xmin=-2.5, ymin=49.0, xmax=1.5, ymax=51.0)

bay <- ne_states(country = c("France","united kingdom"), 
                 returnclass = "sf") %>%
  st_crop(boxchannel)
bbbay<-bb(bay)

df <- ES_Areas %>% 
  dplyr::filter(!Zone  %in% c("Bay", "Ilot Oiseaux")) %>%
  st_transform(4326)

# osm_df <- read_osm(df, ext=1.1)
# save(osm_df, file=sprintf("Sources/OSM_extraction.RData"))
load(sprintf("Sources/OSM_extraction.RData"))

boxestuary<-bb(df)
tm_bay<-
  tm_shape(bay) + 
    tm_fill(col="grey30") +
  tm_shape(st_as_sfc(boxestuary)) + 
    tm_borders("red", lwd = 2) +
  tm_layout(bg.color = "skyblue")

tmap_mode("plot") # "plot" "view"
tm_Bio<-
  # tm_shape(estuary, bbox = boxestuary) + 
  #   tm_polygons() +
  tm_shape(osm_df) +  
    tm_rgb() +
  tm_shape(df) +  
    tm_fill(col = "Zone", palette = "Accent", alpha = 0.7) +
    # tm_text("Zone", size = .8, col= "black") +
    tm_borders("grey", lwd = 1) +
  tm_shape(CSLN_sf) +
    tm_dots(col="grey30",size=0.001, alpha = 0.5) +
  tm_scale_bar(position = c("LEFT", "BOTTOM"), width = 0.15) + #SCALE
  tm_compass(position = c("RIGHT", "BOTTOM"), size = 3) +          #NORTH COMPASS
  tm_layout(legend.position= c("LEFT","TOP"))
tm_Bio

vp <- grid::viewport(0.95, 0.95, width = 0.25, just=c(1, 1))
print(tm_bay, vp = vp)

tmap_save(tm_Bio,filename=
            sprintf("%sfig-map.png",wdres),
          insets_tm=tm_bay, insets_vp=vp,
          dpi=600,
          height=10, units="cm") 

```

<!-- ::: -->

<!-- # ::: {.content-visible when-format="html"} -->

<!-- The Seine estuary has been studied regarding several historically known areas that have either habitat or communities differences. These areas are mainly mudflats and subtidal areas, are represented in @fig-map_htlm. -->

```{r}
#| label: fig-map_html
#| eval: false
#| include: false
#| fig-cap: "Maps of habitats areas defined for the study"

tmap_mode("view") # "plot" "view"
tm_Bio<-
  tm_basemap(leaflet::providers$OpenStreetMap.HOT) +
  tm_shape(df) +
  tm_fill(col = "Zone", palette = "Accent", alpha = 0.7) +
  tm_borders("grey", lwd = 1) +
  tm_shape(df_stations) +
  tm_dots(col="grey30",size=0.001, alpha = 0.5) +
  tm_layout(legend.outside = TRUE)
# tmlf_Bio<-tmap_leaflet(tm_Bio) # conversion to leaflet object, quicker??
tm_Bio
tmap_save(tm_Bio, filename = "Estuary_map.html")

```

<!-- ::: -->

## Biologic dataset

```{r}
#| label: biostats_1

rawnb <-count(CSLN_raw) %>%
  pull() %>% 
  formatC(., format="f", big.mark=",", digits=0)
cleannb <-count(CSLN) %>% 
  pull() %>% 
  formatC(., format="f", big.mark=",", digits=0)
yearsrange <- CSLN %>% 
  mutate(Annee = as.numeric(levels(Annee))[Annee]) %>% # reconvert factors
  arrange(Annee) %>% 
  slice(c(1,n())) %>% 
  pull(Annee) %>%
  paste(collapse = " to ")
moisyear <- CSLN %>% group_by(Annee,Station_originelle) %>% 
  summarise(nbcamp=n_distinct(Mois)) %>% 
  group_by(nbcamp) %>%
  summarise(nbstat=n_distinct(Station_originelle)) %>% 
  arrange(desc(nbstat)) %>%
  slice(1:2) %>% #slice(c(1,n())) %>%
  pull(nbcamp) %>%
  paste(collapse = " to ")
nbstattot <- CSLN %>% 
  summarise(n_distinct(Station_originelle)) %>% 
  pull()
nbstatyr <- CSLN %>% group_by(Annee) %>% 
  summarise(nbstat=n_distinct(Station_originelle)) %>%
  summarise(nbstat=round(mean(nbstat),0)) %>% 
  pull()
period <- CSLN %>% group_by(Mois) %>% 
  mutate(Mois = month.abb[as.numeric(levels(Mois))[Mois]]) %>%
  summarise(nbstat=n_distinct(Station_originelle)) %>%
  arrange(desc(nbstat)) %>% 
  slice(1:3) %>% 
  pull(Mois) %>% 
  paste(collapse = ", ")
recbef2000 <- CSLN_Mars %>% 
  filter(!Zone  %in% c("Channel","Bay")) %>% # at minimum
  filter(Period == "1996-1999") %>%
  filter(!SPCourt %in% c("SEMBA","AMPIM","AUSMO",
                         "BALCR","MYTED","BIVAL","ANNEL")) %>%
  # filter(SPCourt %in% c(espece)) %>% 
  # summarise(n_distinct(idStationUnique)) %>% 
  count() %>% 
  pull()
richspe <- CSLN %>% 
  summarise(nbstat=n_distinct(Taxon_SNa)) %>% 
  pull()
cerednb <- CSLN %>%
  filter(SPCourt %in% c(espece)) %>% 
  summarise(nbstat=n_distinct(idStationUnique)) %>% 
  pull()
cereddetzon <- CSLN %>% 
  filter(SPCourt %in% c(espece)) %>% 
  group_by(Zone) %>% 
  summarise(nbstat=n_distinct(idStationUnique)) %>% 
  unite(Zonenb,Zone,nbstat,sep=" (n= ") %>% 
  pull() %>% 
  paste(collapse = "), ")

```

The raw data (n= `r rawnb`) has been harmonised and grouped to obtain a single database of `r cleannb` observations, and `r nbstattot` sampling stations in total (with some variation in coordinates from year to year), with an average of `r nbstatyr` stations per campaign, occurred mainly in the months of `r period`. Along the all period covered by the dataset, from `r yearsrange`, discarding before 2000 (too few observations n=`r recbef2000`), with `r moisyear` sampling campaigns per year, a series of 5-years periods has been defined, 2000-2005 the building of 'Port 2000' with high disruption in the estuary area; 2006-2010; 2011-2015; 2016-2019. `r richspe` different species has been founded, and *C. edule* counts `r cerednb` observations. Focus for following treatment is made on mudflats only for their biologic relevance for *C. edule* (`r cereddetzon`)).

## HMS dataset

```{r}
#| label: abiostats_1

marsweb <- Mars_csv %>% 
  summarise(nbstat=n_distinct(NINJ)) %>% 
  pull() %>% 
  formatC(., format="f", big.mark=",", digits=0)
marsxy <- Mars_csv %>% select(NINJ) %>% 
  separate(NINJ, c("Y", "X"),sep="_") %>%
  mutate(X=as.numeric(X), Y=as.numeric(Y)) %>% 
  summarise(across(.cols=everything(),~(max(.x)-min(.x)))) %>% # sapply(., max)
  unite(fus,Y,X,sep=" by ") %>%
  pull(fus)
marslonlat <- Mars_csv %>% 
  select(Lon,Lat) %>%
  summarise(across(.cols=everything(),
                   ~(sprintf("%g to %g",
                             round(min(.x),4),
                             round(max(.x),4))))) %>%
  unite(fus,Lon,Lat,sep=" and Latitude ") %>%
  pull(fus)
  
marsmaillenb <-CSLN_Mars %>% 
  select(NINJ) %>% 
  drop_na() %>% 
  summarise(nbstat=n_distinct(NINJ)) %>% 
  pull() %>% 
  formatC(., format="f", big.mark=",", digits=0)
marsmailledet <- CSLN_Mars %>% 
  select(NINJ,Tidal_level) %>% 
  drop_na() %>% 
  group_by(Tidal_level) %>% 
  summarise(nbstat=n_distinct(NINJ)) %>% 
  unite(Tidalnb,Tidal_level,nbstat,sep=" (n= ") %>% 
  pull() %>% 
  paste(collapse = "), ")
```

All Seine Bay MARS3D model is calculated in a grid of `r formatC(700*112, format="f", big.mark=",", digits=0)` mesh (700 by 112), focused on the Seine Estuary with a grid of 'only' `r marsweb` mesh (`r marsxy`), with the coordinates of Longitude `r marslonlat`. Mars mesh corresponding to the stations localisation in the biologic dataset are at total `r marsmaillenb` with the distinction in the position in the tidal area as `r marsmailledet`).

```{r}
#| label: globcorr

df <- CSLN_mud %>% 
  filter(SPCourt == speciesMP$SPCourt[sp]) %>%
  select(paste(factors[,1],saison[sa,1],sep=""),reponse[1:2,1])
# colnames(df)<-c(factors[,2],reponse[,2])
dfcornew <- corstars(df, result="none",removeTriangle="") %>% 
  mutate(variable=rownames(.)) %>% 
  relocate(variable)
# dftest<-dfcornew %>% mutate(across(.cols = everything(), str_replace_all, ".*\\s", ""))

```

## Predictors calculation and selection

All variables have been observed and a correlation study has been conducted to select the more pertinent factors and avoid autocorrelation between factors.

-   **Mud content** \[`r factors[14,3]`\] is chosen over Sediment total conc \[`r factors[13,3]`\] to link with other references (Corr = `r dfcornew %>% filter(variable %in% "mudrate_m") %>% select(Csed_m) %>% pull()`).
-   **Current Speed daily max** \[`r factors[2,3]`\] is chosen over Current_Speed \[`r factors[1,3]`\] to reflect hydrodynamics occurring on mudflats (Corr = `r dfcornew %>% filter(variable %in% "flow_mxd") %>% select(flow_m) %>% pull()`).
-   **Inundation time** \[`r factors[3,3]`\] is chosen rather than True bathymetry \[`r factors[9,3]`\], to be more informative on tidal position (Corr = `r dfcornew %>% filter(variable %in% "bathy") %>% select(inunt) %>% pull()`).
-   **Salinity daily range** \[`r factors[5,3]`\] is preferred to Salinity \[`r factors[4,3]`\] to emphasize more the tidal influence (Corr = `r dfcornew %>% filter(variable %in% "sal_dtd") %>% select(sal_m) %>% pull()`).
-   **Temperature** \[`r factors[6,3]`\] is picked upon Temperature daily range \[`r factors[7,3]`\] for its accessibility (Corr = `r dfcornew %>% filter(variable %in% "temp_dtd") %>% select(temp_m) %>% pull()`) and the range being reflected by the Inundation time (Corr = `r dfcornew %>% filter(variable %in% "temp_dtd") %>% select(inunt) %>% pull()`).
-   **Bed shear stress** \[`r factors[11,3]`\] is preferred to MES mud \[`r factors[8,3]`\] to focus on erosion phenomena more than MTZ processes (Corr = `r dfcornew %>% filter(variable %in% "MESmud_m") %>% select(tenfon_m) %>% pull()`).
-   **Bed shear stress daily max** is poorly correlated to Bed shear stress \[`r factors[12,3]`\] (Corr = `r dfcornew %>% filter(variable %in% "tenfon_mxd") %>% select(tenfon_m) %>% pull()`) but the former is preferred to indicate the extreme local conditions that can lead to erosion of sediment.
-   Yearly sediment budget \[`r factors[10,3]`\] is not correlated to any other variable but its distribution makes it not relevant.

## Scenarios of interest

Models are built with two abiotic factors to increase their reliability. The couples tested are selected to apply them depending on data available and the focus on abiotic conditions, which can be in particular impacted by the global climate change:

1.  Current Speed daily max \[`r factors[2,3]`\] and Inundation time \[`r factors[3,3]`\]: to compare results with [@cozzoli2014], and these variables are easily retrieved with some high frequency measurement (COASTHF) . They have also an interest because they contain information on the position in the tidal area that could evolve with sea level rise and the hydrological conditions impact by the river flowrate change due to climate change even though they have a significant correlation (Corr = `r dfcornew %>% filter(variable %in% "flow_mxd") %>% select(inunt) %>% pull()`)
2.  Mud content \[`r factors[14,3]`\] and Bed shear stress \[`r factors[11,3]`\]: Those variables are determinant in a building of an erosion model such as MARS3D (Corr = `r dfcornew %>% filter(variable %in% "mudrate_m") %>% select(tenfon_m) %>% pull()`).
3.  Temperature \[`r factors[6,3]`\] and Salinity daily range \[`r factors[5,3]`\]: These factors are easily measurable on high frequency also (Somlit). They are both bearing the climate change for the global temperature change and the evolution of the river regime that would impact the salinity profile of the estuary (Corr = `r dfcornew %>% filter(variable %in% "temp_m") %>% select(sal_dtd) %>% pull()`).
4.  Temperature \[`r factors[6,3]`\] and Inundation time \[`r factors[3,3]`\]: (Corr = `r dfcornew %>% filter(variable %in% "temp_m") %>% select(inunt) %>% pull()`).

```{r}
#| label: fig-corrmatrix
#| include: true
#| fig-cap: "Here is the matrix plot" 

# df <- CSLN_mud %>% 
#   filter(SPCourt == speciesMP$SPCourt[sp]) %>%
#   select(paste(factors[,1],saison[sa,1],sep=""),
#          reponse[1:2,1])
# colnames(df)<-c(factors[,2],reponse[,2])
# tp <- ggpairs(df, #title=titreG,
#               progress=FALSE,
#               upper = list(continuous = corr_col),
#               lower = list(continuous= 
#                              wrap("smooth", size = .5, 
#                                   alpha = 0.4, color = "navyblue")))
# print(tp)

```

# Results

## Biological data set description

```{r}
#| label: prefig-cered_struc
y_lab <- "Biomass (gAFDW/m2)" # "MSR (mW/m2)"
CSLN_df <- CSLN %>%
  filter(grepl("Mudflat", Zone)) %>% # focus on mudflat only
  filter(SPCourt %in% c(espece)) %>%
  filter(!is.na(Biomass_gAFDWm2)) %>%
  mutate(across(where(is.factor), fct_drop)) %>% # recode factors when filter has remove levels
  group_by(idStationUnique,Zone,Tidal_level,Period,Annee) %>% 
  summarise(iZTPAS = sum(Biomass_gAFDWm2, na.rm =TRUE)) %>% 
  group_by(Zone,Period,Tidal_level,Annee) %>% #,Season
  summarise(ZTPAS = mean(iZTPAS, na.rm =TRUE)) %>% 
  mutate(Period=factor(Period, 
                      labels = str_replace_all(unique(Period),"-"," - ")))

cereddetper <- CSLN %>% 
  filter(SPCourt %in% c(espece)) %>% 
  group_by(Period) %>% 
  summarise(nbstat=n_distinct(idStationUnique)) %>% 
  unite(Periodnb,Period,nbstat,sep=" (n= ") %>% 
  pull() %>% 
  paste(collapse = "), ")

stat.test <- CSLN_df %>% 
  group_by(Zone) %>% 
  wilcox_test(ZTPAS ~ Period) %>% 
  add_xy_position(x = "Period",group = "Zone",
                  dodge = 0,step.increase = 0.05) %>%
  filter(p.adj.signif!="ns")
gpA <- ggplot(CSLN_df) + 
  geom_boxplot(aes(x=Period, y = ZTPAS, fill = Period)) + 
  facet_grid(. ~ Zone, 
             scales = "free_y",
             labeller = label_wrap_gen(width = 10)) + 
  labs(y = y_lab, x="") +
  stat_pvalue_manual(stat.test,label = "p.adj.signif", tip.length = 0.02) +
  Scale_brew() +
  # guides(fill=guide_legend(nrow=2,byrow=TRUE)) +
  theme(legend.position="bottom",
    axis.text.x = element_blank(), axis.ticks = element_blank())
print(gpA)

stat.test <- CSLN_df %>% 
  group_by(Period) %>%
  wilcox_test(ZTPAS ~ Zone) %>% 
  add_xy_position(x = "Zone",group = "Period",
                  dodge = 0,step.increase = 0.05) %>%
  filter(p.adj.signif!="ns")

gpB <- ggplot(CSLN_df) + 
  geom_boxplot(aes(x=Zone, y = ZTPAS, fill = Zone)) + 
  facet_grid(. ~ Period, 
             scales = "free_y",
             labeller = label_wrap_gen(width = 10)) + 
  labs(y = y_lab, x="") +
  stat_pvalue_manual(stat.test,label = "p.adj.signif", tip.length = 0.02) +
  Scale_brew() +
  guides(fill=guide_legend(nrow=2,byrow=TRUE)) +
  theme(legend.position="bottom",
    axis.text.x = element_blank(), axis.ticks = element_blank())
print(gpB)

titre <- paste("C. edule population in Seine estuary",sep="")
bp <- ggarrange(gpA,gpB, ncol=1, nrow=2, 
                labels="AUTO", legend="bottom", 
                common.legend = FALSE)
bp <- annotate_figure(bp,
                      ) + #top = text_grob(titre, face = "bold", size = 14)
      bgcolor("white")

```

With a closer look on the data for *C. edule*, the differences of biomass \[gAFDW/m²\] are represented regarding periods and areas (@fig-cered_struc) with their significative differences. The observations are split into periods as `r cereddetper`). They are seldom in North Upstream Mudflat, where the water is more brackish.

```{r}
#| label: fig-cered_struc
#| include: true
#| fig-cap: "C. edule population biomass [gAFDW/m²] in the Seine estuary, by Period for each area (A) and by areas for each Period (B) with significant differences"
#| fig-width: 6
#| fig-height: 8
print(bp + 
        theme(title = element_blank(),
          text = element_text(size = 12)))

ggsave(sprintf("%sfig-cered_struc.tiff",wdres),
       plot = bp, width = 10, height = 8, dpi=600)
```


## SDM_NEO

```{r}
#| label: sdmchoices

sdmname_l<- matrix(nrow = 4,ncol = 2)
for (repi in 1:2){
  sdmname_l[1,repi]<-sprintf("%g%g%g%g%g",sp,sa,repi,
                    which(pred_red$Var %in% c("flow_mxd")),
                    which(pred_red$Var %in% c("inunt")))
  sdmname_l[2,repi]<-sprintf("%g%g%g%g%g",sp,sa,repi,
                    which(pred_red$Var %in% c("mudrate_m")),
                    which(pred_red$Var %in% c("tenfon_mxd")))
  sdmname_l[3,repi]<-sprintf("%g%g%g%g%g",sp,sa,repi,
                    which(pred_red$Var %in% c("sal_dtd")),
                    which(pred_red$Var %in% c("temp_m")))
  sdmname_l[4,repi]<-sprintf("%g%g%g%g%g",sp,sa,repi,
                    which(pred_red$Var %in% c("inunt")),
                    which(pred_red$Var %in% c("temp_m")))
}

```

### Linear QR type selection

```{r}
#| label: aic_res

repi<-1
sdi<-1
df <- rqdata$smrq_l %>% 
  filter(sdmname %in% c(paste0("RQ2add_",sdmname_l[sdi,repi]),
                        paste0("RQ2int_",sdmname_l[sdi,repi]))) %>% 
  group_by(sdmname, factorsor1, factorsor2) %>% 
  summarise(AIC_m=round(mean(AIC),2),
            AIC_sd=round(sd(AIC),2))

```

Using the first factor couple as example, the comparison between linear QR with or without interaction (@fig-add_int) shows there is a difference of result that does not shows on the AIC result (@fig-aic with biomass, simple addition AIC=`r df$AIC_m[1]`+/-`r df$AIC_sd[1]` , with interaction AIC=`r df$AIC_m[2]`+/-`r df$AIC_sd[2]`). The linear model with interaction is preferred for biologic reasons: the fauna reacts to the conjunction of factors.

```{r}
#| label: fig-aic
#| fig-cap: "SDM Linear Quantile Regression"
#| include: true

repi<-1
sdi<-1
titleG <- sprintf("%s and %s",unique(df$factorsor1),unique(df$factorsor2))
AIC_plot<-rqdata$smrq_l %>% 
  filter(sdmname %in% c(paste0("RQ2add_",sdmname_l[sdi,]),
                        paste0("RQ2int_",sdmname_l[sdi,]))) %>% 
  group_by(tau) %>%
  ggplot(aes(x=reponse, #reorder(factors,-AIC),
             y=AIC,
             color = factor(tau),
             shape = mode))+
  geom_point(size= 5) +
  scale_shape_manual(values=c(1, 8))+
  scale_color_manual(values=colRQ) +
 # facet_grid(.~reponse) +
  labs(#title=titleG, 
       x="Biologic",y="AIC",
       color="Quantile",shape="Model") +
  theme(axis.text = element_text(size=10,face="bold")) +
  coord_flip(); AIC_plot
  
ggsave(sprintf("%sfig-aic.tiff",wdres),
       plot = AIC_plot,
       width = 8, height = 8, dpi=600)

```

```{r}
#| label: fig-add_int
#| fig-cap: "SDM bifactorial Linear Quantile Regression comparision of two calculation models: simple addition of the two factors (A), with interaction of the factors together (B). All negatives results of model are NA and appear in grey."
#| fig-width: 12
#| fig-height: 6
#| include: true

repi<-1
sdi<-1
titleG <- sprintf("Linear QR comparison addition and interacton")
p1<-rqdata$dp2dF_plot_list[[paste0("RQ2add_",sdmname_l[sdi,repi])]] + 
      labs(title="Addition")
p2<-rqdata$dp2dF_plot_list[[paste0("RQ2int_",sdmname_l[sdi,repi])]] + 
      labs(title="Interaction")

plot_tmp <- ggarrange(
  p1 + theme(axis.title = element_blank()),
  p2 + theme(axis.title = element_blank()),
  ncol=2, nrow=1,
  labels="AUTO",legend="right",
  common.legend = TRUE)
plot_tmp <- annotate_figure(plot_tmp,
                # top = text_grob(titleG, size=18),
                left = text_grob(p1$labels$y, rot = 90),
                bottom = text_grob(p1$labels$x)) +
        bgcolor("white")

plot_tmp
ggsave(sprintf("%sfig-add_int.tiff",wdres),
     plot = plot_tmp,
     width = 16, height = 8, dpi=600)
```

### Comparison of the three types of model
Still illustrated by the first couples of factors, the three model types are compared (@fig-3mod). The B-Spline methods shows that it covers the data where they are, but with a overfitting that mathematically is correct but does not allow biologic process modelisation.

```{r}
#| label: fig-3mod
#| fig-cap: "SDM QR model type comparison: linear with interaction (A), Gaussian non-linear (B) and Cubic B-Splines linear (C). All negatives results of model are NA and appear in grey."
#| fig-width: 12
#| fig-height: 6
#| include: true

repi<-1
sdi<-1
titleG <- sprintf("Linear QR comparison addition and interacton")
p1<-rqdata$dp2dF_plot_list[[paste0("RQ2int_",sdmname_l[sdi,repi])]] + 
      labs(title="Linear interaction",
           fill=NULL)
p2<-nlrqdata$dp2dF_plot_list[[paste0("nlRQ2_",sdmname_l[sdi,repi])]] + 
      labs(title="Gaussian non-linear",
           fill=NULL)
p3<-rqbsdata$dp2dF_plot_list[[paste0("RQbs2_",sdmname_l[sdi,repi])]] + 
      labs(title="B-Spline linear",
           fill=NULL)

plot_tmp <- ggarrange(
  p1 + theme(axis.title = element_blank(),
             legend.text = element_text(size=8)) +
      scale_fill_gradientn(colours=colRQ,limits=c(0,max(
        max(p1$data$RqMod,na.rm=TRUE),
        max(p2$data$RqMod,na.rm=TRUE)
        ))),
  p2 + theme(axis.title = element_blank(),
             legend.text = element_text(size=8)) +
      scale_fill_gradientn(colours=colRQ,limits=c(0,max(
        max(p1$data$RqMod,na.rm=TRUE),
        max(p2$data$RqMod,na.rm=TRUE)
        ))),
  p3 + theme(axis.title = element_blank(),
             legend.text = element_text(size=8)) +
      scale_fill_gradientn(colours=colRQ,limits=c(0,max(
        max(p1$data$RqMod,na.rm=TRUE),
        max(p2$data$RqMod,na.rm=TRUE)
        ))),
  ncol=3, nrow=1,
  labels="AUTO",
  legend="bottom", common.legend = FALSE)
plot_tmp <- annotate_figure(plot_tmp,
                # top = text_grob(titleG, size=18),
                left = text_grob(p1$labels$y, rot = 90),
                bottom = text_grob(p1$labels$x)) +
        bgcolor("white")

plot_tmp
ggsave(sprintf("%sfig-3mod.tiff",wdres),
     plot = plot_tmp,
     width = 16, height = 8, dpi=600)
```

### Quantile choice

```{r}
#| label: quant_choice
t <- which(taus %in% c(0.975))  # c(0.5,0.9,0.95,0.975)

```

The quantile is chosen as a compromise between having a high quantile without over representing some high observations. The quantile chosen is \tau = `r t`, and is used for all further analyses.

```{r}
#| label: graphrq
#| output: false
# Linear RQ Graphics generation
analysis <- "RQ Linear"
rqname<- matrix(paste0("RQ2int_",sdmname_l),nrow=4,ncol=2)

plot_rq <- list()  
for (repi in 1:2) { # repi=1
  tmp<-list()
  for (sdi in 1:nrow(sdmname_l)) { # sdi=1
    tmp[[sdi]] <- rqdata$dp2d_plot_list[[rqname[sdi,repi]]][[t]] + 
      labs(fill=NULL)
      # geom_contour_tanaka(aes(x=x1,y=x2,z = RqMod))
      # plot_gg(tmp[[sdi]], width = 5, height = 5, 
      #     raytrace = TRUE, multicore = TRUE, scale = 250, 
      #     zoom = 0.7, theta = 10, phi = 30, windowsize = c(800, 800))
    
    # tmpF[[sdi]] <- rqdata$dp2dF_plot_list[[rqname[sdi,repi]]][[t]] 

      }
  plot_rq[[repi]]<-tmp
  # plot_rqF[[repi]]<-tmpF
  plot_rq_t <- ggarrange(plotlist = plot_rq[[repi]], 
                         ncol=1, nrow=length(plot_rq[[repi]]),
                         labels="AUTO",legend="right",
                         common.legend = FALSE)
  plot_rq_t <- annotate_figure(plot_rq_t,
                        top = text_grob(analysis)) +
        bgcolor("white")
}


```

```{r}
#| label: graphnlrq
#| output: false
# Nonlinear RQ Graphics generation
analysis <- "RQ Nonlinear"
nlrq2name<- matrix(paste0("nlRQ2_",sdmname_l),nrow=4,ncol=2)

plot_nlrq <- list()
for (repi in 1:2){
  tmp <- list()
  for (sdi in 1:nrow(sdmname_l)) { # sdi=1
      tmp[[sdi]] <- nlrqdata$dp2d_plot_list[[nlrq2name[sdi,repi]]][[t]] + 
      labs(fill=NULL)
  }
  plot_nlrq[[repi]]<-tmp
  plot_nlrq_t <- ggarrange(plotlist = plot_nlrq[[repi]], 
                         ncol=1, nrow=length(plot_nlrq[[repi]]),
                         labels="AUTO",legend="right",
                         common.legend = FALSE)
  plot_nlrq_t <- annotate_figure(plot_nlrq_t,
                        top = text_grob(analysis)) +
        bgcolor("white")
}

```

```{r}
#| label: graphrqbs
#| eval: false
#| output: false
# BSplines RQ Graphics generation
analysis <- "RQ BSpline"
rqbs2name<- matrix(paste0("RQbs2_",sdmname_l),nrow=4,ncol=2)

plot_rqbs <- list()
for (repi in 1:2) {
  tmp <- list()
  for (sdi in 1:nrow(sdmname_l)) { # sdi=1
    tmp[[sdi]] <- rqbsdata$dp2d_plot_list[[rqbs2name[sdi,repi]]][[t]] + 
      labs(fill=NULL)
  }
  plot_rqbs[[repi]]<-tmp
  plot_rqbs_t <- ggarrange(plotlist = plot_rqbs[[repi]], 
                         ncol=1, nrow=length(plot_rqbs[[repi]]),
                         labels="AUTO",legend="right",
                         common.legend = FALSE)
  plot_rqbs_t <- annotate_figure(plot_rqbs_t,
                        top = text_grob(analysis)) +
        bgcolor("white")
}

```

### Comparison of linear and nonlinear QR

```{r}
#| label: prefig-graphboard
#| output: false
# Create a board of plot

titleG <- sprintf("%s in %s (\tau=%g)",
                    speciesMP$Taxon_SNa[sp],saison[sa,2],taus[t])
plot_t <- list()
leti<-c("A","B","C","D")
for (repi in 1:2) { # repi=1
  tmp<-list()
  subtitleG <- sprintf("SDM-NEO %s",reponse[repi,3])
  for (i in 1:length(plot_rq[[repi]])) { # i=2
    plot_rq[[repi]][[i]] <- plot_rq[[repi]][[i]] +
      labs(title=ifelse(i==1, "RQ Linear","")) +
      theme(axis.title = element_blank(),
            plot.title = element_text(size=16)) +
      scale_fill_gradientn(colours=colRQ,limits=c(0,max(
        max(plot_rq[[repi]][[i]]$data$RqMod,na.rm=TRUE),
        max(plot_nlrq[[repi]][[i]]$data$RqMod,na.rm=TRUE)
        )))
    plot_nlrq[[repi]][[i]] <- plot_nlrq[[repi]][[i]] +
      labs(title=ifelse(i==1, "RQ Nonlinear","")) +
      theme(axis.title = element_blank(),
            plot.title = element_text(size=16)) +
      scale_fill_gradientn(colours=colRQ,limits=c(0,max(
        max(plot_rq[[repi]][[i]]$data$RqMod,na.rm=TRUE),
        max(plot_nlrq[[repi]][[i]]$data$RqMod,na.rm=TRUE)
        )))  
    tmp[[i]] <- ggarrange(plot_rq[[repi]][[i]],
                          plot_nlrq[[repi]][[i]], 
                          ncol=2, nrow=1,
                          labels=paste0(leti[i],1:2), hjust=0, vjust=2,
                          legend="right", common.legend = TRUE)
    tmp[[i]] <- annotate_figure(tmp[[i]], 
                    left = text_grob(plot_rq[[repi]][[i]]$labels$y, rot = 90),
                    bottom = text_grob(plot_rq[[repi]][[i]]$labels$x))
          
  }
  plot_t[[repi]] <- ggarrange(plotlist = tmp,
                         ncol=1, nrow=4,
                         labels=NULL,#legend="right",
                         common.legend = FALSE)

  plot_t[[repi]] <- annotate_figure(plot_t[[repi]],
                        top = text_grob(titleG, size=18),
                        bottom = text_grob(subtitleG, size=18)) +
        bgcolor("white")
  ggsave(sprintf("%s_%s_%s_%s_%s.tiff",
                 etude,speciesMP$SPCourt[sp],
                 reponse[repi,3],saison[sa,2],taus[t]),
         plot = plot_t[[repi]], width = 8, height = 12, dpi=600)
}

```

The four couples of factors have been treated with the different ways of calculation: either linear with interaction, or non-linear with bifactorial gaussian equation with the biomass as biologic response @fig-graphboard. AIC index has been calculated for each couple, model type and quantile, and is available on plots.

#### 2D visualisation

```{r}
#| label: fig-graphboard
#| include: true
#| fig-cap: "Selected factors couples models computed with linear with interaction (left, numbered 1) and nonlinear with gaussian equation (right numbered 2), the biologic response being the biomass in Ash Free Dry Weight per square meter"
#| fig-width: 6
#| fig-height: 9
plot_t[[1]]

ggsave(sprintf("%sfig-graphboard.tiff",wdres),
       plot = plot_t[[1]], width = 6, height = 12, dpi=600)
```

The same board of plots with density are in @fig-graphboard2. *can be put in supplementary data*

```{r}
#| label: fig-graphboard2
#| include: true
#| fig-cap: "Selected factors couples models computed with linear with interaction (left, numbered 1) and nonlinear with gaussian equation (right numbered 2), the biologic response being the density in individuals per square meter"
#| fig-width: 6
#| fig-height: 9
plot_t[[2]]

ggsave(sprintf("%sfig-graphboard2.tiff",wdres),
       plot = plot_t[[2]], width = 6, height = 12, dpi=600)
```

#### 3D visualisation NOT WORKING

```{r}
#| label: graphrq3d
#| eval: false
#| output: false
# Linear RQ Graphics generation
analysis <- "RQ Linear"
rqname<- matrix(paste0("RQ2int_",sdmname_l),nrow=4,ncol=2)

plot_rq <- list()  
for (repi in 1:2) {
  tmp<-list()
  for (sdi in 1:nrow(sdmname_l)) { # sdi=1
    tmp[[sdi]] <- rqdata$dp3d_plot_list[[rqname[sdi,repi]]][[t]]
  }
  plot_rq[[repi]]<-tmp
  
# PLOTLY FACET

}

```

```{r}
#| label: graphnlrq3d
#| eval: false
#| output: false
# Nonlinear RQ Graphics generation
analysis <- "RQ Nonlinear"
nlrq2name<- matrix(paste0("nlRQ2_",sdmname_l),nrow=4,ncol=2)

plot_nlrq <- list()
for (repi in 1:2){
  tmp <- list()
  for (sdi in 1:nrow(sdmname_l)) { # sdi=1
      tmp[[sdi]] <- nlrqdata$dp3d_plot_list[[nlrq2name[sdi,repi]]][[t]]
  }
  plot_nlrq[[repi]]<-tmp
  
}

```

```{r}
#| label: graphrqbs3d
#| eval: false
#| output: false
# BSplines RQ Graphics generation
analysis <- "RQ BSpline"
rqbs2name<- matrix(paste0("RQbs2_",sdmname_l),nrow=4,ncol=2)

plot_rqbs <- list()
for (repi in 1:2) {
  tmp <- list()
  for (sdi in 1:nrow(sdmname_l)) { # sdi=1
    tmp[[sdi]] <- rqbsdata$dp3d_plot_list[[rqbs2name[sdi,repi]]][[t]]
  }
  plot_rqbs[[repi]]<-tmp
  
}

```

```{r}
#| label: fig-mpg
#| fig-cap: "Global title."
#| fig-subcap:
#|   - "sub title of graph1"
#|   - "sub title of graph2"
#| layout-ncol: 2
#| column: page
#| include: true
#| eval: false

titleG <- sprintf("%s in %s (t=%g)",
                    speciesMP$Taxon_SNa[sp],saison[sa,2],taus[t])
repi<-1; sdi<-1
# for (sdi in 1:length(plot_rq[[repi]])) { # i=2
#   print(plot_rq[[repi]][[sdi]])
#   print(plot_nlrq[[repi]][[sdi]])
# }

#widgetframe::frameWidget(plot_rq[[repi]][[sdi]])
```

```{r}
#| label: prefig-graphboard3d
#| output: false
#| eval: false
# Create a board of plot

titleG <- sprintf("%s in %s (t=%g)",
                    speciesMP$Taxon_SNa[sp],saison[sa,2],taus[t])
plot_t <- list()
leti<-c("A","B","C","D")
for (repi in 1:2) { # repi=1
  tmp<-list()
  subtitleG <- sprintf("SDM-NEO %s",reponse[repi,3])
  for (i in 1:length(plot_rq[[repi]])) { # i=2
    x1l<-pred_red[as.numeric(substr(sdmname_l[i],4,4)),2]
    x2l <- pred_red[as.numeric(substr(sdmname_l[i],5,5)),2]
    zl<-sprintf("%s (%s)",reponse[repi,2],reponse[repi,3])
    
    plot_rq[[repi]][[i]] <- plot_rq[[repi]][[i]] 
    
    plot_nlrq[[repi]][[i]] <- plot_nlrq[[repi]][[i]]

scene <- list(camera = list(eye = list(x = -1.5, y = 1.5, z = 0.3)),
              aspectmode='cube',
              xaxis = list(title = x1l), 
              yaxis = list(title = x2l),
              zaxis = list(title = zl),
              domain=list(x=c(0,0.5),y=c(0,1)))
scene2 <- list(camera = list(eye = list(x = -1.5, y = 1.5, z = 0.3)),
               aspectmode='cube',
               xaxis = list(title = x1l), 
               yaxis = list(title = x2l),
               zaxis = list(title = zl),
               domain=list(x=c(0.5,1),y=c(0,1)))

    tmp[[i]] <- plotly::subplot(list(plot_rq[[repi]][[i]],plot_nlrq[[repi]][[i]]),
                                 shareX = TRUE, shareY = TRUE, nrows = 2) %>%
      layout(scene = scene, scene2 = scene2)
    saveWidget(dp3dF_plot_list,
               sprintf("%s%s/%s/nlRQ2/%s_%s_nlRQ23dFacet_%s_%s_%s_%s.html",
                       wdgraphEx,speciesMP$SPCourt[sp],
                       analysis,etude,speciesMP$SPCourt[sp],
                       saison[sa,2],yt,x1t,x2t),
               selfcontained = F, libdir = "lib")
    
    tmp[[i]] <- ggarrange(plotlist =
                            list(plot_rq[[repi]][[i]],plot_nlrq[[repi]][[i]]), 
                         ncol=2, nrow=1,
                         labels=paste0(leti[i],1:2), hjust=0, vjust=2,
                         legend="right", common.legend = TRUE)
    tmp[[i]] <- annotate_figure(tmp[[i]], 
                    left = text_grob(plot_rq[[repi]][[i]][["labels"]][["y"]], rot = 90),
                    bottom = text_grob(plot_rq[[repi]][[i]][["labels"]][["x"]]))
          
  }
  plot_t[[repi]] <- ggarrange(plotlist = tmp,
                         ncol=1, nrow=4,
                         labels=NULL,#legend="right",
                         common.legend = FALSE)
  
  # plot_t <- ggarrange(plotlist = c(rbind(plot_rq, plot_nlrq)), 
  #                        ncol=2, nrow=length(plot_rq),
  #                        labels="AUTO",legend="right",
  #                        common.legend = FALSE)
  # plot_t <- ggarrange(plotlist = list(plot_rq_t, plot_nlrq_t), 
  #                        ncol=2, nrow=1,
  #                        labels=NULL,legend="right",
  #                        common.legend = FALSE)
  
  plot_t[[repi]] <- annotate_figure(plot_t[[repi]],
                        top = text_grob(titleG, size=18),
                        bottom = text_grob(subtitleG, size=18)) +
        bgcolor("white")
  ggsave(sprintf("%s%s/%s_%s_%s_%s_%s.tiff",
                 wdgraph,speciesMP$SPCourt[sp],
                 etude,speciesMP$SPCourt[sp],
                 reponse[repi,3],saison[sa,2],taus[t]),
         plot = plot_t[[repi]], width = 8, height = 12, dpi=600)
}

# subtitleG <- sprintf("%s, tau = %g",xl,taus[t])
# scene <- list(camera = list(eye = list(x = -1.5, y = 1.5, z = 0.3)), aspectmode='cube',
#             xaxis = list(title = x1l), yaxis = list(title = x2l),
#             zaxis = list(title = sprintf("%s (%s)",reponse[1,2],reponse[1,3])),
#             domain=list(x=c(0,0.5),y=c(0,1)))
# scene2 <- list(camera = list(eye = list(x = -1.5, y = 1.5, z = 0.3)), aspectmode='cube',
#              xaxis = list(title = x1l), yaxis = list(title = x2l),
#              zaxis = list(title = sprintf("%s (%s)",reponse[2,2],reponse[2,3])),
#              domain=list(x=c(0.5,1),y=c(0,1)))
# dp3dF <- subplot(plot_dp3dlist, shareX = TRUE, shareY = TRUE, nrows = 2) %>%
# layout(title = paste(titleG,subtitleG,sep="\n"),
#        scene = scene, scene2 = scene2)
# saveWidget(dp3dF,sprintf("%s%s/%s/%s_%s_RQ2int3dFacet_%s_%s_%s_%s.html",
#                        wdgraphEx,speciesMP$SPCourt[sp],analysis,etude,speciesMP$SPCourt[sp],
#                        saison[sa,2],x1t,x2t,taus[t]),
#          selfcontained = F, libdir = "lib")
# 
# rm(list=ls(pattern="scene")); rm(list=ls(pattern="plot"))

```

See https://communicate-data-with-r.netlify.app/docs/communicate/htmlwidgets-in-documents/

https://stackoverflow.com/questions/74379298/argument-selfcontained-deprecated-in-htmlwidgetssavewidget

## SDM-NEO Maps
```{r}
save.image(rdataout)
```


see https://bookdown.org/nicohahn/making_maps_with_r5/docs/tmap.html#static-maps-with-tmap to make maps with facets with all years and the gif animation easily

```{r}
#| label: fig-sdm_map
#| fig-cap: "SDM maps"
#| include: true
Mars_SDM<-Mars_dat_sf %>% dplyr::select(c(NINJ,Lon,Lat,Zone,Tidal_level,Annee,pred_red$Var))

sfpF<-list()
sfp<-list()
tmp_sf<-list()
tmp_gif<-list()
for (mt in 1:nrow(model_types)){ # mt=2
  for (repi in 1:2){
    for (sdi in 1:nrow(sdmname_l)) { # sdi=1
      mltp<-model_types[mt,]
      sdmn<-sdmname_l[sdi,repi]
      sdname<- paste0(mltp$code,sdmn)
      modelq <- get(mltp$bdd)$modelq_list[[sdname]]
      # la pb avec le nlrq qui se presente en list et non  en table comme pour le lineaire...
      smrq_l<-get(mltp$bdd)$smrq_l
      yl<-sprintf("%s (%s)",reponse[repi,2],reponse[repi,3])
      sdmname_tau<- sprintf("%s_t%g",sdname,taus[t])
  
      k<- substr(sdmn,4,4); k2<-substr(sdmn,5,5)
      zt<-paste(speciesMP[sp,1],sdi,saison[sa,1],sep="")
      x1t = sprintf("%s%s",pred_red[k,1],saison[sa,1])
      x1l = sprintf("%s%s (%s)",pred_red[k,2],saison[sa,1],pred_red[k,3])
      x2t = sprintf("%s%s",pred_red[k2,1],saison[sa,1])
      x2l = sprintf("%s%s (%s)",pred_red[k2,2],saison[sa,1],pred_red[k2,3])
  
      titleG <- sprintf("%s %s in %s",
                        mltp$analysis,speciesMP$Taxon_SNa[sp],saison[sa,2])
      subtitleG = sprintf("%s & %s", x1l,x2l)
      capt = sprintf("%s Quantile regression\n%s", 
                     smrq_l$mode, taus[t])
      
      # MARS3D SDM CALCULATION
      x1M <- as.data.frame(pull(Mars_dat_sf, x1t)) %>% rename(x1=1)
      x2M <- as.data.frame(pull(Mars_dat_sf, x2t)) %>% rename(x2=1)
      x1x2mars <- bind_cols(x1M,x2M)
  
      RqModMars <- as.data.frame(predict(modelq,newdata=x1x2mars)) %>%
        select(contains(sprintf("%g",taus[t]))) %>% 
        rename(RqMod=1) %>%
        mutate(across(everything(), function(x){replace(x, which(x<0), NA)})) %>% 
        mutate(tau=taus[t], 
               taust=sprintf("tau= %g",tau),
               sdmname=sdname) %>% 
        bind_cols(x1x2mars)
      Mars_SDM$SDM<-pull(RqModMars %>% select(RqMod))
      Mars_SDM<-Mars_SDM %>%
        select(-contains(sdname)) %>% 
        rename(!!sdmname_tau:=SDM) # !!'string': to interpret text as variable
      
      # SDM-NEO MAP BUILDING
      Mars_SDM_sf <- Mars_SDM %>% 
        dplyr::select(c(Annee, !!sdmname_tau)) # raster::select also exists!
      sfpF[[sdmname_tau]] <-ggplot(data = Mars_SDM_sf) +
        # annotation_map_tile(zoom = 13, 
        #                   cachedir = system.file("rosm.cache", 
        #                                          package = "ggspatial")) +
        # labs(caption = "\U00a9 OpenStreetMap contributors") +
        geom_sf(aes(fill = .data[[sdmname_tau]]), color = NA) + #
          labs(x="Latitude",
               y="Longitude",
               title = titleG,
               fill = sprintf("%s\n%s",
                              sdmname_tau,yl)) +
          theme(legend.title = element_text(size=10),
                legend.position = "bottom",
                plot.margin = margin(0.05,0.05,0.05,0.05, "cm")) +
          facet_wrap(~Annee) +
          scale_fill_distiller(palette = "Spectral") 
      ggsave(sprintf("%s%s/%s/%s_%s_%s.tiff",
                     wdgraph,speciesMP[sp,1],mltp$analysis,
                     etude,speciesMP[sp,1],sdmname_tau),
             plot = sfpF[[sdmname_tau]], width = 12, height = 12, dpi=600)

      anMars<-unique(Mars_SDM_sf$Annee)
      sfpa<-list()
      for (ann in 1:length(anMars)) { # ann=1 :length(anMars)
        Mars_SDM_sfi<-Mars_SDM_sf %>% filter(Mars_SDM_sf$Annee==anMars[ann])
        sfpa[[ann]]<-ggplot(data = Mars_SDM_sfi) +
            # annotation_map_tile(zoom = 13, 
            #                   cachedir = system.file("rosm.cache", 
            #                                          package = "ggspatial")) +
            # labs(caption = "\U00a9 OpenStreetMap contributors") +
            geom_sf(aes(fill = .data[[sdmname_tau]]), color = NA) + #
            labs(x="Latitude",
                 y="Longitude",
                 title = anMars[ann],
                 fill = sprintf("%s\n%s",
                                sdmname_tau,yl)) +
            theme(legend.title = element_text(size=10),
                  plot.margin = margin(0.05,0.05,0.05,0.05, "cm")) +
            scale_fill_distiller(palette = "Spectral") # + 
            # coord_sf(xlim = c(-88, -78), ylim = c(24.5, 33), expand = FALSE)
     } #year
      sfp[[sdmname_tau]]<-sfpa
      
      # nncol = 5; nnrow=ceiling(length(anMars)/nncol)
      # sfp <- ggarrange(plotlist=sfp, 
      #                  ncol=nncol, nrow=nnrow,
      #                  labels = "AUTO",
      #                  common.legend = TRUE, legend="bottom")+
      #   theme(plot.margin = margin(0.05,0.05,0.05,0.05, "cm")) 
      # sfp <- annotate_figure(sfp, 
      #                        top = text_grob(sprintf("%s\n%s",titleG,subtitleG), 
      #                                        face = "bold", size = 14),
      #                        bottom = capt)+
      #   bgcolor("white");
      # ggsave(sprintf("%s%s/%s/%s_%s_%s.tiff",
      #                wdgraph,speciesMP[sp,1],mltp$analysis,
      #                etude,speciesMP[sp,1],sdmname_tau),
      #        plot = sfp, width = 12, height = 12, dpi=600)
      
      
      # tmap_mode("view") # "plot" "view"
      # tmp_sf[[sdmname_tau]] <- 
      #   tm_basemap(leaflet::providers$OpenStreetMap.HOT) +
      #   tm_shape(Mars_SDM_sf) +
      #     tm_fill(col = sdmname_tau, 
      #             palette = "-Spectral", alpha = 0.7) +
      #   tm_facets(along = "Annee") +
      #   tm_layout(legend.outside = TRUE, # legend.position = c("right", "bottom"), # 
      #             title.size=10
      #             )
      # tmp_sf[[sdmname_tau]]
      # tmap_save(tmp_sf[[sdmname_tau]], filename = sprintf("%s%s/%s/%s_%s_%s.html",
      #                wdgraph,speciesMP[sp,1],mltp$analysis,
      #                etude,speciesMP[sp,1],sdmname_tau))
      
      # tmap_animation(
      #   tmp_sf[[sdmname_tau]], filename = sprintf("%s%s/%s/%s_%s_%s.gif",
      #                wdgraph,speciesMP[sp,1],mltp$analysis,
      #                etude,speciesMP[sp,1],sdmname_tau),
      #   fps = 1, width = 2000, height = 2000, dpi=600
      #   )
      
    }
  }
}
# st_write(Mars_SDM, sprintf("%sLayers made/SDM_NEO_RQ_%s.shp",wdGIS,speciesMP$SPCourt[sp]),append=FALSE)

```


# Final actions and save

Rdata are saved in `rdataout`. An excel file collects data in `r binderout` , with sheets for :

-   data : contains whatever

```{r}
#| label: finalsave_xsl
#| eval: false

wb <- copyWorkbook(loadWorkbook(binderin))
# copyworkbook to avoid any pb that can occur on the excel file
if (!("AllMeas" %in% names(wb))) {
  addWorksheet(wb=wb, sheetName = "AllMeas")}
writeData(wb, sheet = "AllMeas", x = data, 
          startCol = 1, startRow = 1,withFilter = FALSE)

saveWorkbook(wb,file=binderout, overwrite = TRUE)

```

```{r}
#| label: finalsave_rdata

# rm("rqdata","nlrqdata","rqbsdata")
# save.image(rdataout)
beepr::beep(2)

```

# References {.unnumbered}
