---
title: "TREATMENT OF BIOLOGIC DATA FROM CSLN BEFORE SDM TREATMENT"
author: "Amélie Lehuen"
description: "add info on script"
date: "`r format(Sys.time(), '%B %Y')`"
editor: source
execute:
  eval: true
  echo: false
  message: false
  warning: false
  output: true
  include: false
editor_options: 
  chunk_output_type: console
---

# TREATMENT OF BIOLOGIC DATA FROM CSLN BEFORE SDM TREATMENT

## Script preparation

### Packages

```{r}
#| label: load-packages
library(conflicted)
library(knitr)

library(readxl) ; library(openxlsx) # Edition d'un fichier Excel
library(tidyverse) #The tooltm_box indispensable
library(reshape2); library(rlist) # the one; melt; list.append
library(data.table)
library(rstatix); library(Hmisc)  # corr and pvalue calculation

library(ggpubr); library(GGally); # stat_compare_means ;
library(scales); library(RColorBrewer); library(wesanderson); library(grafify); library(colorspace); library(ggsci)# show_col and colors colors colors!
library(patchwork)
library(ggridges)
# GIS Packages#
library(sf); library(sfheaders); # st_as_sf ; sf_to_df
library(tmap); library(tmaptools)
library(htmlwidgets) # library(leaflet) # saveWidget ; for interactive maps
library(introdataviz) # geom_split_violin # devtools::install_github("psyteachr/introdataviz")

conflict_prefer("transpose", "purrr")
conflict_prefer("filter", "dplyr")
conflict_prefer("select", "dplyr")
conflict_prefer("summarise", "dplyr")
conflict_prefer("melt", "reshape2")
conflict_prefer("layout", "plotly")

options(scipen=999) # Prevents scientific display of numbers
```

### Working Environment

```{r}
#| label: workenvir

rm(list=ls())
wdsource <- "Data/Faune/CSLN/"
wdmat <- "Matrices/"
wdgraph <- "C:/Users/lehuen201/Back Up AMLH/Melting Potes/BDD/A_SDM_NEO/Plots/" #"Plots/"
wdres <- "Results/"
wdproj<-getwd()
pc <- "C:/Users/lehuen201/Nextcloud/" # "E:/" #
wdGIS <- paste(pc,"Melting Pot/SIG/",sep="")

```

### Home made functions

```{r}
#| label: functmade

source("Scripts/0.0_SDM_NEO_functions.R")

```

### Graphic charter

```{r}
#| label: graphchart

```

## Load of External data and Basic Variables

### External data

```{r}
#| label: externdata

rdatain <- sprintf("%sCSLN_Mars_BDD.RData",wdmat)
load(rdatain)
```

### Output binder and data

```{r}
#| label: outfiles

rdataout <- sprintf("%sCSLN_Mars_plot",wdmat)
# if exists
load(paste0(rdataout,".RData"))

```

### Basic variables

```{r}
#| label: basicvar

etude <- "CSLN_Mars"
espece <- "CERED"

sp <- which(speciesMP$SPCourt == espece)
sai <- which(saison$M_Def == "Year") #Year Winter Summer

```

# Material and methods

## Study area

The Seine estuary has been studied regarding several historically known areas that have either habitat or communities differences. These areas are mainly mudflats and subtidal areas, are represented in @fig-map.

```{r}
#| label: prefig-map

tmap_mode("plot") # "plot" "view"
# Extract data of interest
tm_areas <- ES_Areas %>% 
  st_transform(4326)

# Zoom tm_boxes
tm_boxestuary<-bb(tm_areas)
tm_box_mud <- ES_Areas %>%  
  dplyr::filter(grepl("Mudflat",ES_Areas$Zone)) %>%
  st_transform(4326) %>% bb(ext=2)
# tm_boxchannel<-bb(c(xmin=-2.5, ymin=49.0, xmax=1.5, ymax=51.0) )
tm_boxchannel<-bb(tm_boxestuary, ext=14)

# Extract Open street map background based on data

# osm_df <- read_osm(bb(Mars_SDM), ext=1, type= "osm") #stamen-watercolor osm
# osm_df_mud <- read_osm(tm_box_mud, ext=1, type= "osm") #stamen-watercolor osm
# osm_df_wat_bay <- read_osm(tm_boxchannel, ext=1, type= "stamen-watercolor")
# osm_df_wat_estu <- read_osm(tm_boxestuary, ext=1, type= "stamen-watercolor")
# save(osm_df,osm_df_mud,osm_df_wat_bay,osm_df_wat_estu, file=sprintf("SIG/OSM_extraction.RData"))
load(sprintf("SIG/OSM_extraction.RData"))

# create a map of context with red rect of area of interest
tm_bay<-
  tm_shape(osm_df_wat_bay) +  
    tm_rgb(saturation = .6, alpha= 0.8) +
  tm_shape(st_as_sfc(tm_boxestuary)) + 
    tm_borders("red", lwd = 2) +
  tm_layout(legend.show= FALSE) 
tm_estu<-tm_shape(osm_df_wat_estu) +  
    tm_rgb(saturation = .6, alpha= 0.8) +
  tm_shape(tm_areas) +  
    tm_fill(col = "Zone", title="Areas", palette = areas_pal, alpha = 0.5) +
  tm_shape(st_as_sfc(tm_box_mud)) + 
    tm_borders("red", lwd = 2) +
  tm_layout(legend.show= FALSE)

# Create main map
tm_Bio<-
  tm_shape(osm_df_mud) +  
    tm_rgb(saturation = .6, alpha= 0.8) +
  tm_shape(tm_areas) +  
    tm_fill(col = "Zone", title="Areas", palette = areas_pal, alpha = 0.5) +
    # tm_text("Zone", size = .8, col= "black") +
    tm_borders("grey", lwd = 1.5) +±
  tm_shape(CSLN_sf) +
    tm_dots(col="grey30", size=0.1, alpha = 0.3) +
  tm_scale_bar(position = c("LEFT", "BOTTOM"), width = 0.15) + #SCALE
  tm_compass(type = "8star",position = c("LEFT", "TOP"), size = 2) +  #NORTH COMPASS
  tm_layout(#legend.outside = TRUE,
    legend.position= c("center","BOTTOM"),
    legend.title.size=.8,legend.text.size=.5,
    legend.bg.color="white",legend.bg.alpha=0.6)

# arrange maps insets thanks to patchwork !
tm_Bio_g<-tmap_grob(tm_Bio)
tm_bay_g<-tmap_grob(tm_bay)
tm_estu_g<-tmap_grob(tm_estu)
design <- c(
  area(t= 1, b= 14, l= 1, r= 20),
  area(t= 4, b= 6, l= 18, r= 20),
  area(t= 8, b= 11, l= 17, r= 20) )
tm_study_map<- wrap_plots(tm_Bio_g,tm_bay_g,tm_estu_g,design = design)

ggsave(filename= sprintf("%sfig-map.tiff",wdgraph), 
       tm_study_map, dpi=600, width = 10,  height = 7) #
```

```{r}
#| label: map_html
#| include: true

tmap_mode("view") # "plot" "view"
tm_study_map_html<-
  tm_basemap(leaflet::providers$OpenStreetMap.HOT) +
  tm_shape(tm_areas) +  
    tm_fill(col = "Zone", palette = areas_pal, alpha = 0.5, legend.show = FALSE) +
    # tm_text("Zone", size = .8, col= "black") +
    tm_borders("grey", lwd = 1.5) +
  tm_shape(CSLN_sf) +
    tm_dots(col="Zone", title="Areas", palette = areas_pal, size=0.05, alpha = 0.7) +
  tm_layout(legend.show=FALSE)
tm_study_map_html
tmap_save(tm_study_map_html, filename = "SIG/Estuary_map.html")

```

## Predictors calculation and selection

```{r}
#| label: pl_mars_corr
#| include: true

df <- CSLN_mud %>%
  filter(SPCourt == espece) %>% 
  select(paste(predict[,1],saison[sai,1],sep=""),
         map(reponse_l,~.x$rvar) %>% unlist(., use.names=FALSE)) %>%
  select(-c("eptot_dt"))

pl_mars_corr <- ggpairs(df, 
                        labeller=label_wrap_gen(width = 10,multi_line = TRUE),
                        upper = list(continuous = f.corr_col),
                        lower = list(continuous=
                                       wrap("smooth", size = .3,
                                            alpha = 0.4, color = "aquamarine4")),
                        columnLabels = c(predict[!predict$Var %in% c("eptot_dt"),2],reponse[,3]),
                        progress=FALSE) +
      theme(axis.text = element_text(size=8),
            strip.text.y = element_text(size=8))

ggsave(sprintf("%sfig-mars_corr.tiff",wdgraph),
       plot = pl_mars_corr, width = 16, height = 9, dpi=400)

```

# Results

## Biological data set description

::: column-screen-inset-right

```{r}
#| label: biom_distr_bd
CSLN_df <- CSLN %>%
  filter(grepl("Mudflat", Zone)) %>% # focus on mudflat only
  filter(SPCourt %in% c(espece)) %>%
  drop_na(Biomass_gAFDWm2) %>%
  drop_na(Density_indm2) %>% 
  mutate(across(where(is.numeric), ~f.replace_outliers(.,3))) %>%
  mutate(across(where(is.factor), fct_drop)) %>% # recode predict when filter has remove levels
  mutate(Period=factor(Period, 
                      labels = str_replace_all(unique(Period),"-"," - ")))

pl_des <- map(reponse_l,
              ~f.data_descr(CSLN_df,.x$rvar))
pl_des_gt <- map(pl_des, ~{
                  .x$geog$plot / .x$time$plot & 
                    theme(text=element_text(size=12)) &
                    plot_annotation(tag_levels = c('A'))
                 } )

iwalk(pl_des_gt,
  ~ggsave(sprintf("%sfig-bio_struc_%s.tiff",wdgraph,.y),
       plot = .x, width = 14, height = 7, dpi=400) )

```
:::

## HMS data set description
```{r}
#| label: hms_distr

Mars_tmp<-Mars_SDM %>%
           st_drop_geometry %>%
           group_by(Zone) %>% 
           # mutate(across(where(is.numeric), ~f.replace_outliers(.,2))) %>% 
           mutate(across(all_of(c("bathy","tenfon_mxd","temp_m")), ~f.replace_outliers(.,1.5))) %>%
           filter(!Zone %in% c("Octeville","Cote Fleurie")) %>%
           mutate(across(where(is.factor), fct_drop))

pl_mars <- 
  map(setNames(as.list(Var_choosen),Var_choosen), \(vars)
          f.data_descr(Mars_tmp,
                     vars) )

pl_mars<- map(setNames(as.list(Var_choosen),Var_choosen),  \(vars)
  map(pl_mars[[vars]], ~{modify_at(.x,"plot", 
          ~{ .x + ylab(pred_red$whole[which(pred_red$Var == vars)])} )} ) ) 

pl_mars_gt <- map(pl_mars,
                  ~{.x$geog$plot / .x$time$plot  & 
    theme(text=element_text(size=12)) &
    plot_annotation(tag_levels = c('A')) } )

iwalk(pl_mars_gt,
  ~ggsave(sprintf("%sfig-hms_struc_%s.tiff",wdgraph,.y),
       plot = .x, width = 14, height = 7, dpi=400) )

# Done after saving file because of special characters
pl_mars<-pl_mars %>% 
  setNames(pred_red$whole)

pl_mars_gt<-pl_mars_gt %>% 
  setNames(pred_red$whole)

```

```{r}
#| label: hms_stat

Mars_tmp<- Mars_SDM_per %>%
           st_drop_geometry() %>%
           group_by(Zone) %>% 
           mutate(across(any_of(c("bathy","tenfon_mxd","temp_m")), ~f.replace_outliers(.x,1.5))) %>%
           filter(!Zone %in% c("Octeville","Cote Fleurie")) %>%
           mutate(across(where(is.factor), fct_drop))

# pl_mars_st_board <- Mars_tmp %>% 
#   st_drop_geometry() %>% 
#   pivot_longer(cols = !NINJ:Annee) %>% 
#   mutate(name=factor(name, levels = pred_red$Var, labels = pred_red$whole)) %>% 
#   ggplot() + 
#   geom_density_ridges(aes(y = Period , fill=Period,
#                           height = after_stat(count), # to have a comparative height between periods
#                           x=value), 
#                       stat="density", alpha=0.3, size = .2) +
#   facet_grid(name ~ Zone, 
#              scales = "free",
#              labeller = label_wrap_gen(width = 20) ) + 
#   labs(x="", y="") +
#   Scale_brew() + 
#   theme(legend.position="none",
#         strip.background = element_blank(),
#         text=element_text(size=10))

pl_mars_st_board <-
  pl_var_st_board(Mars_tmp,pred_red)

ggsave(sprintf("%s/mars_st_board.tiff", wdgraph),
                  plot = pl_mars_st_board$all,
                  width = 10, height = 12, dpi=600)

```

```{r}
#| label: hms_map

tmap_mode("plot") # "plot" "view"

Mars_tmp<-Mars_SDM_per %>%
           # st_drop_geometry() %>%
           group_by(Zone) %>% 
           mutate(across(all_of(c("bathy","tenfon_mxd","temp_m")), ~f.replace_outliers(.,1.5))) #%>%
           # filter(!Zone %in% c("Octeville","Cote Fleurie")) %>%
           # mutate(across(where(is.factor), fct_drop))

pl_mars_map <-
    map(pred_red %>% transpose %>% 
           setNames(.,pred_red$Var),
        ~f.pl_var_map(Mars_tmp,.x))


iwalk(pl_mars_map, ~tmap_save(.x$all, dpi=400, height=2, asp=0, #width=20,
                          filename=sprintf("%s/%s_map.tiff",
                              wdgraph, .y) ) )

tm_mars_gb <- map(pl_mars_map, ~ .x$all + 
                    tm_layout(main.title.size= .6,
                              panel.label.size=.6,
                              legend.title.size= .6,
                              legend.text.size= .6)) %>%
    tmap_arrange(., asp=NA, ncol=1,
                 sync = TRUE, outer.margins=0)

tmap_save(tm_mars_gb, dpi=600,
              filename= sprintf("%sfig-graphboard_map_mars.tiff",wdgraph) )

```

## Animated GIF Maps plots
```{r}
#| label: hms_map_gif

# Mars_tmp<-Mars_SDM %>%
#            # mutate(across(where(is.numeric), ~f.replace_outliers(.,2))) %>% 
#            # mutate(across(all_of(c("bathy","tenfon_mxd")), ~f.replace_outliers(.,1.5))) %>%
#            # filter(!Zone %in% c("Octeville","Cote Fleurie")) %>%
#            mutate(across(where(is.factor), fct_drop))

# pl_mars_map_gif <-
#     map(pred_red %>% transpose %>% 
#            setNames(.,pred_red$Var),
#         ~f.pl_var_map_gif(Mars_tmp,.x))

```

# Final actions and save

Rdata are saved in `rdataout`. An excel file collects data in `r binderout` , with sheets for :

-   data : contains whatever

```{r}
#| label: finalsave_r

# rm(list=lsf.str()) # remove all functions
rm(list=(ls()[grepl("tmp", ls())] ))
save(list=(setdiff(ls()[grep("^tm_|^pl_", ls())], lsf.str())), # select only variables that respect pattern, remove functions
     file = paste0(rdataout,".RData"))
beepr::beep(2)
```

# Supplementary data

## Software details

::: {.callout-tip collapse="true"}
## Expand for R Session Info

```{r, echo = FALSE}
#| include: true

library(sessioninfo)
pkg_sesh <- session_info(pkgs = "attached")
quarto_version <- system("quarto --version", intern = TRUE)
pkg_sesh$platform$quarto <- paste(
  system("quarto --version", intern = TRUE),
  "@",
  quarto::quarto_path()
  )
pkg_sesh
```
:::

# References {.unnumbered}