---
title: "TREATMENT OF BIOLOGIC DATA FROM CSLN BEFORE SDM TREATMENT"
author: "Amélie Lehuen"
description: "add info on script"
date: "`r format(Sys.time(), '%B %Y')`"
editor: source
execute:
  eval: true
  echo: false
  message: false
  warning: false
  output: true
  include: false
editor_options: 
  chunk_output_type: console
---

# TREATMENT OF BIOLOGIC DATA FROM CSLN BEFORE SDM TREATMENT

## Script preparation

### Packages

```{r}
#| label: load-packages

library(readxl) ; library(openxlsx) # Edition d'un fichier Excel
library(knitr)
library(tidyverse) #The tooltm_box indispensable
library(reshape2); library(rlist) # the one; melt; list.append
library(data.table)
library(rstatix); library(Hmisc)  # corr and pvalue calculation

library(ggpubr); #library(GGally); # stat_compare_means ;
library(scales); library(RColorBrewer); library(wesanderson); library(grafify); library(colorspace); library(ggsci)# show_col and colors colors colors!
library(patchwork)
library(ggridges)
# GIS Packages
library(sf); library(sfheaders); # st_as_sf ; sf_to_df
library(tmap); library(tmaptools)
library(htmlwidgets) # library(leaflet) # saveWidget ; for interactive maps
library(introdataviz) # geom_split_violin # devtools::install_github("psyteachr/introdataviz")

conflict_prefer("transpose", "purrr")
conflict_prefer("filter", "dplyr")
conflict_prefer("select", "dplyr")
conflict_prefer("summarise", "dplyr")
conflict_prefer("melt", "reshape2")
conflict_prefer("layout", "plotly")

options(scipen=999) # Prevents scientific display of numbers
```

### Working Environment

```{r}
#| label: workenvir

rm(list=ls())
wdsource <- "Data/Faune/CSLN/"
wdmat <- "Matrices/"
wdgraph <- "C:/Users/lehuen201/Back Up AMLH/Melting Potes/BDD/A_SDM_NEO/Plots/" #"Plots/"
wdres <- "Results/"
wdproj<-getwd()
pc <- "C:/Users/lehuen201/Nextcloud/" # "E:/" #
wdGIS <- paste(pc,"Melting Pot/SIG/",sep="")

```

### Definition of the graphic charter

```{r}
#| label: graphchart
theme_set(theme_bw(base_size = 14)) # theme_gray() theme_bw() theme_light()
colDarj <- function(x) {wes_palette("Darjeeling2",x, type = "continuous")}
colZiss <- function(x) {wes_palette("Zissou1",x, type = "continuous")}
colSpec <- colorRampPalette(brewer.pal(8, "Spectral")); # show_col(colSpec(2))
colDark <- colorRampPalette(brewer.pal(8, "Dark2"));
# Scale_col <- function(x) {scale_colour_manual(values=colDarj(x))}
# Scale_fill <- function(x) {scale_fill_manual(values=colDarj(x))}

pal_cspx <- function(x) {divergingx_hcl(x,palette = "Zissou 1")};
colBin<-c(pal_cspx(6)[3],pal_cspx(6)[5]); 
colRQ<-pal_cspx(4); blank<-alpha("#dae8ed",0.1); # show_col(colBin)
Scalc_rq<- function() {scale_colour_manual(values=colRQ)}
Scalf_rq2d <- function() {scale_fill_gradientn(colours=colRQ)} 
pal_cspx <- function(x) {divergingx_hcl(x,palette = "Zissou 1")}; # show_col(pal_cspx(6)) # colorspace
Scale_brew <- function()
  {scale_colour_brewer(palette="Spectral",aesthetics=c("colour","fill"))}
Scale_co <- function() 
  {scale_colour_manual(palette=colorRampPalette(brewer.pal(11, "Spectral")))}
Scale_fi <- function() 
  {scale_fill_manual(palette=colorRampPalette(brewer.pal(11, "Spectral")))}

pal_ggscc <- pal_material("teal"); # show_col(pal_ggscc(9)) # ggsci
colSum <- c(pal_ggscc(9)[3],pal_ggscc(9)[7],pal_ggscc(9)[1]) # show_col(colSum)
colInliers<-pal_material("teal")(10)[5] ; colOutliers = pal_material("deep-orange")(10)[5] # show_col(colOutliers)
Scalc_misc <- function() {scale_color_discrete_diverging(palette = "Blue-Red 3")}
Scale_map<- function() {scale_fill_distiller(palette = "Spectral")}
```

### Home made functions

```{r}
#| label: functmade

```
## Load of External data and Basic Variables
### External data

```{r}
#| label: externdata

rdatain <- sprintf("%sCSLN_Mars_BDD.RData",wdmat)
load(rdatain)
```

### Output binder and data

```{r}
#| label: outfiles

rdataout <- sprintf("%sCSLN_Mars_plot",wdmat)
# if exists
# load(rdataout)

```

### Basic variables

```{r}
#| label: basicvar

etude <- "CSLN_Mars"
espece <- "CERED"
analysis <- "RQ Linear"
sp <- which(speciesMP$SPCourt == espece)
sai <- which(saison$M_Def == "Year") #Year Winter Summer

```

## Study area

The Seine estuary has been studied regarding several historically known areas that have either habitat or communities differences. These areas are mainly mudflats and subtidal areas, are represented in @fig-map.

```{r}
#| label: prefig-map

# Extract data of interest
tm_areas <- ES_Areas %>% 
  dplyr::filter(!Zone  %in% c("Bay", "Ilot Oiseaux")) %>%
  st_transform(4326)

# Zoom tm_boxes
tm_boxestuary<-bb(tm_areas)
tm_box_mud <- ES_Areas %>%  
  dplyr::filter(grepl("Mudflat",ES_Areas$Zone)) %>%
  st_transform(4326) %>% bb(ext=1.4)
tm_boxchannel<-bb(c(xmin=-2.5, ymin=49.0, xmax=1.5, ymax=51.0) )
tm_boxchannel<-bb(tm_boxestuary, ext=12)

# Extract Open street map background based on data
# osm_df <- read_osm(tm_box_mud, ext=1, type= "osm") #stamen-watercolor osm
# osm_df_wat_bay <- read_osm(tm_boxchannel, ext=1, type= "stamen-watercolor")
# osm_df_wat_estu <- read_osm(tm_boxestuary, ext=1, type= "stamen-watercolor")
# save(osm_df,osm_df_wat_bay,osm_df_wat_estu, file=sprintf("SIG/OSM_extraction.RData"))
load(sprintf("SIG/OSM_extraction.RData"))

# create a map of context with red rect of area of interest
tm_bay<-
  tm_shape(osm_df_wat_bay) +  
    tm_rgb(saturation = .6, alpha= 0.8) +
  tm_shape(st_as_sfc(tm_boxestuary)) + 
    tm_borders("red", lwd = 2) +
  tm_layout(legend.show= FALSE) 
tm_estu<-tm_shape(osm_df_wat_estu) +  
    tm_rgb(saturation = .6, alpha= 0.8) +
  tm_shape(st_as_sfc(tm_box_mud)) + 
    tm_borders("red", lwd = 2) +
  tm_shape(tm_areas) +  
    tm_fill(col = "Zone", title="Areas", palette = "Set2", alpha = 0.7) +
  tm_layout(legend.show= FALSE)

# Create main map
tmap_mode("plot") # "plot" "view"
tm_Bio<-
  tm_shape(osm_df) +  
    tm_rgb(saturation = .6, alpha= 0.8) +
  tm_shape(tm_areas) +  
    tm_fill(col = "Zone", title="Areas", palette = "Set2", alpha = 0.6) +
    # tm_text("Zone", size = .8, col= "black") +
    tm_borders("grey", lwd = 1.5) +
  tm_shape(CSLN_sf) +
    tm_dots(col="grey30", size=0.1, alpha = 0.5) +
  tm_scale_bar(position = c("LEFT", "BOTTOM"), width = 0.15) + #SCALE
  tm_compass(type = "8star",position = c("LEFT", "TOP"), size = 3) +  #NORTH COMPASS
  tm_layout(legend.position= c("LEFT","center"),
            legend.bg.color="white",legend.bg.alpha=0.6)

# arrange maps insets thanks to patchwork !
tm_Bio_g<-tmap_grob(tm_Bio)
tm_bay_g<-tmap_grob(tm_bay)
tm_estu_g<-tmap_grob(tm_estu)
design <- c(
  area(t= 1, b= 4, l= 1, r= 6),
  area(t= 2, b= 2, l= 6, r= 6),
  area(t= 3, b= 3, l= 6, r= 6) )
tm_study_map<- wrap_plots(tm_Bio_g,tm_bay_g,tm_estu_g,design = design)

ggsave(filename= sprintf("%sfig-map.tiff",wdgraph), 
       tm_study_map, dpi=600, height = 8) #, width = 12
```

```{r}
#| label: map_html
#| eval: false

# tmap_mode("view") # "plot" "view"
# tm_Bio<-
#   tm_basemap(leaflet::providers$OpenStreetMap.HOT) +
#   tm_shape(df) +
#   tm_fill(col = "Zone", palette = "Accent", alpha = 0.7) +
#   tm_borders("grey", lwd = 1) +
#   tm_shape(df_stations) +
#   tm_dots(col="grey30",size=0.001, alpha = 0.5) +
#   tm_layout(legend.outside = TRUE)
# # tmlf_Bio<-tmap_leaflet(tm_Bio) # conversion to leaflet object, quicker??
# tm_Bio
# tmap_save(tm_Bio, filename = "Estuary_map.html")

```

# Results

```{r}
#| label: data_set_description

data_descr <- function (dataset,vars) {
    # dataset=Mars_SDM %>% st_drop_geometry %>% filter(!Zone %in% c("Bay","Octeville"))
      # vars ="inunt"
  descr<-list(geog=list(stat.test=NULL,plot=NULL),
              time=list(stat.test=NULL,plot=NULL))

  descr$geog$stat.test <- dataset %>% 
    group_by(Zone) %>% 
    wilcox_test(formula(paste0(vars," ~ Period"))) %>% 
    add_xy_position(x = "Period",group = "Zone",
                    dodge = 0,step.increase = 0.05) %>%
    filter(p.adj.signif!="ns")
  descr$geog$plot <- ggplot(dataset) + 
    # geom_boxplot(aes(x=Period, y = .data[[vars]], fill = Period)) +
    # geom_density_ridges(aes(y = Period , fill=Period,
    #                       x=.data[[vars]]), alpha=0.3) +
    geom_flat_violin(aes(x=Period, fill = Period,
                     y = .data[[vars]] ),
                     alpha = 0.5) +
    stat_summary(aes(x=Period, fill = Period,
                     y = .data[[vars]], color = Period),
                 fun.data = mean_se, show.legend = FALSE) +
    stat_pvalue_manual(descr$geog$stat.test,label = "p.adj.signif",
                       tip.length = 0.02, coord.flip = TRUE) +
    coord_flip() +
    facet_wrap( ~ Zone, 
               nrow = 1, #scales = "free_y", 
               labeller = label_wrap_gen(width = 10)) + 
    labs(y = vars, x="") +
    Scale_brew() +
    theme(legend.position="bottom")#,
      # axis.text.x = element_blank(), axis.ticks = element_blank())

  descr$time$stat.test <- dataset %>% 
    group_by(Period) %>%
    wilcox_test(formula(paste0(vars," ~ Zone"))) %>% 
    add_xy_position(x = "Zone",group = "Period",
                    dodge = 0,step.increase = 0.05) %>%
    filter(p.adj.signif!="ns")

  descr$time$plot <- ggplot(dataset) + 
    # geom_boxplot(aes(x=Zone, y = .data[[vars]], fill = Zone)) +
    geom_flat_violin(aes(x=Zone, fill = Zone,
                     y = .data[[vars]] ),
                     alpha = 0.5) +
    stat_summary(aes(x=Zone, fill = Zone,
                     y = .data[[vars]], color = Zone), 
                 fun.data = mean_se, show.legend = FALSE) +
    stat_pvalue_manual(descr$time$stat.test,label = "p.adj.signif", 
                       tip.length = 0.02, coord.flip = TRUE) +
    coord_flip() +
    facet_wrap( ~ Period, 
               nrow = 1, #scales = "free_y", 
               labeller = label_wrap_gen(width = 10)) + 
    labs(y = vars, x="") +
    Scale_brew() +
    guides(fill=guide_legend(nrow=2,byrow=TRUE)) +
    theme(legend.position="bottom")
  
  return(descr)
}

```

## Biological data set description

::: column-screen-inset-right

```{r}
#| label: biom_distr_b
CSLN_df <- CSLN %>%
  filter(grepl("Mudflat", Zone)) %>% # focus on mudflat only
  filter(SPCourt %in% c(espece)) %>%
  filter(!is.na(Biomass_gAFDWm2)) %>%
  mutate(across(where(is.factor), fct_drop)) %>% # recode predict when filter has remove levels
  group_by(idStationUnique,Zone,Tidal_level,Period,Annee) %>% 
  summarise(iZTPAS = sum(Biomass_gAFDWm2, na.rm =TRUE)) %>% 
  group_by(Zone,Period,Tidal_level,Annee) %>% #,Season
  summarise(Biomass_gAFDWm2 = mean(iZTPAS, na.rm =TRUE)) %>% 
  mutate(Period=factor(Period, 
                      labels = str_replace_all(unique(Period),"-"," - ")))

pl_des_b <- data_descr(CSLN_df,"Biomass_gAFDWm2")
```

```{r}
#| label: biom_distr_d
CSLN_df <- CSLN %>%
  filter(grepl("Mudflat", Zone)) %>% # focus on mudflat only
  filter(SPCourt %in% c(espece)) %>%
  filter(!is.na(Density_indm2)) %>%
  mutate(across(where(is.factor), fct_drop)) %>% # recode predict when filter has remove levels
  group_by(idStationUnique,Zone,Tidal_level,Period,Annee) %>% 
  summarise(iZTPAS = mean(Density_indm2, na.rm =TRUE)) %>% 
  group_by(Zone,Period,Tidal_level,Annee) %>% #,Season
  summarise(Density_indm2 = mean(iZTPAS, na.rm =TRUE)) %>% 
  mutate(Period=factor(Period, 
                      labels = str_replace_all(unique(Period),"-"," - ")))

pl_des_d <- data_descr(CSLN_df,"Density_indm2")

# pl_des_b_gt <- ggarrange(pl_des_b$geog$plot,pl_des_b$time$plot, ncol=1, nrow=2,
#                 labels="AUTO", legend="bottom",
#                 common.legend = FALSE)
# pl_des_b_gt <- annotate_figure(pl_des_b) + #top = text_grob(title, face = "bold", size = 14)
#       bgcolor("white")

pl_des_b_gt <- pl_des_b$geog$plot / pl_des_b$time$plot & 
    theme(legend.position="bottom",
          text=element_text(size=12)) &
    plot_annotation(tag_levels = c('A'))
pl_des_d_gt <- pl_des_d$geog$plot / pl_des_d$time$plot  & 
    theme(legend.position="bottom",
          text=element_text(size=12)) &
    plot_annotation(tag_levels = c('A'))

ggsave(sprintf("%sfig-cered_struc_biom.tiff",wdgraph),
       plot = pl_des_b_gt, width = 12, height = 8, dpi=600)
ggsave(sprintf("%sfig-cered_struc_dens.tiff",wdgraph),
       plot = pl_des_d_gt, width = 12, height = 8, dpi=600)
```
:::

## HMS data set description

```{r}
#| label: pl_mars_corr
#| include: true

# ggpairs custom colors on corr
corr_col <- function(data, mapping, method="p", use="pairwise", ...){
  # grab data
  x <- eval_data_col(data, mapping$x)
  y <- eval_data_col(data, mapping$y)
  # calculate correlation
  corr <- cor(x, y, method=method, use=use)
  # calculate colour based on correlation value
  colFn <- colorRampPalette(c(colSpec(2)[1], "white", colSpec(2)[2]), interpolate ='spline')
  fill <- colFn(100)[findInterval(corr, seq(-1, 1, length=100))]
  ggally_text(
    label = as.character(round(corr, 2)),
    mapping = aes(),
    xP = 0.5, yP = 0.5,
    ...) + #  ggally_cor(data = data, mapping = mapping, ...) + 
    theme_void() +
    theme(panel.background = element_rect(fill=fill))

} #wrap(cor_func,method = 'spearman', symbol = "Corr:\n")


df <- CSLN_mud %>%
  filter(SPCourt == espece) %>%
  select(paste(predict[,1],saison[sai,1],sep=""),
         map(reponse,~.x$rvar) %>% unlist(., use.names=FALSE))

pl_mars_corr <- ggpairs(df, labeller=label_wrap_gen(width = 10,multi_line = TRUE),# title=titreG,
              progress=FALSE,
              upper = list(continuous = corr_col),
              lower = list(continuous=
                             wrap("smooth", size = .3,
                                  alpha = 0.4, color = "aquamarine4"))) +
      theme(axis.text = element_text(size=8),
            strip.text.y = element_text(size=8))

ggsave(sprintf("%sfig-pl_mars_corr.tiff",wdgraph),
       plot = pl_mars_corr, width = 16, height = 9, dpi=600)

```

```{r}
#| label: hms_distr

pl_mars <- 
  map(setNames(as.list(Var_choosen),Var_choosen), \(vars)
          data_descr(Mars_SDM %>%
                       st_drop_geometry %>%
                       filter(!Zone %in% c("Bay","Octeville")),
                     vars) )

pl_mars_gt <- map(pl_mars,
                  ~{.x$geog$plot / .x$time$plot  & 
    theme(legend.position="bottom",
          text=element_text(size=12)) &
    plot_annotation(tag_levels = c('A')) } )

iwalk(pl_mars_gt,
  ~ggsave(sprintf("%sfig-hms_struc_%s.tiff",wdgraph,.y),
       plot = .x, width = 12, height = 8, dpi=600) )

```

# Final actions and save

Rdata are saved in `rdataout`. An excel file collects data in `r binderout` , with sheets for :

-   data : contains whatever

```{r}
#| label: finalsave_r
#| 
save(list=(ls()[grep("^tm_|^pl_", ls())]),
     file = paste0(rdataout,".RData"))
beepr::beep(2)
```

# Supplementary data

## Software details

::: {.callout-tip collapse="true"}
## Expand for R Session Info

```{r, echo = FALSE}
#| include: true

library(sessioninfo)
pkg_sesh <- session_info(pkgs = "attached")
quarto_version <- system("quarto --version", intern = TRUE)
pkg_sesh$platform$quarto <- paste(
  system("quarto --version", intern = TRUE),
  "@",
  quarto::quarto_path()
  )
pkg_sesh
```
:::

# References {.unnumbered}