---
title: "TREATMENT OF BIOLOGIC DATA FROM CSLN BEFORE SDM TREATMENT"
author: "Amélie Lehuen"
description: "add info on script"
date: "`r format(Sys.time(), '%B %Y')`"
editor: source
execute:
  eval: true
  echo: false
  message: false
  warning: false
  output: true
  include: false
editor_options: 
  chunk_output_type: console
---

# TREATMENT OF BIOLOGIC DATA FROM CSLN BEFORE SDM TREATMENT

## Script preparation

### Packages

```{r}
#| label: load-packages
library(conflicted)
library(knitr)

library(readxl) ; library(openxlsx) # Edition d'un fichier Excel
library(tidyverse) #The tooltm_box indispensable
library(reshape2); library(rlist) # the one; melt; list.append
library(data.table)
library(rstatix); library(Hmisc)  # corr and pvalue calculation

library(ggpubr); library(GGally); # stat_compare_means ;
library(scales); library(RColorBrewer); library(wesanderson); library(grafify); library(colorspace); library(ggsci)# show_col and colors colors colors!
library(patchwork)
library(ggridges)
# GIS Packages#
library(sf); library(sfheaders); # st_as_sf ; sf_to_df
library(tmap); library(tmaptools)
library(htmlwidgets) # library(leaflet) # saveWidget ; for interactive maps
library(introdataviz) # geom_split_violin # devtools::install_github("psyteachr/introdataviz")

conflict_prefer("transpose", "purrr")
conflict_prefer("filter", "dplyr")
conflict_prefer("select", "dplyr")
conflict_prefer("summarise", "dplyr")
conflict_prefer("melt", "reshape2")
conflict_prefer("layout", "plotly")

options(scipen=999) # Prevents scientific display of numbers
```

### Working Environment

```{r}
#| label: workenvir

rm(list=ls())
wdsource <- "Data/Faune/CSLN/"
wdmat <- "Matrices/"
wdgraph <- "C:/Users/lehuen201/Back Up AMLH/Melting Potes/BDD/A_SDM_NEO/Plots/" #"Plots/"
wdres <- "Results/"
wdproj<-getwd()
pc <- "C:/Users/lehuen201/Nextcloud/" # "E:/" #
wdGIS <- paste(pc,"Melting Pot/SIG/",sep="")

```

### Definition of the graphic charter

```{r}
#| label: graphchart
theme_set(theme_bw(base_size = 14)) # theme_gray() theme_bw() theme_light()
colDarj <- function(x) {wes_palette("Darjeeling2",x, type = "continuous")}
colZiss <- function(x) {wes_palette("Zissou1",x, type = "continuous")}
colSpec <- colorRampPalette(brewer.pal(8, "Spectral")); # show_col(colSpec(2))
colDark <- colorRampPalette(brewer.pal(8, "Dark2"));
# Scale_col <- function(x) {scale_colour_manual(values=colDarj(x))}
# Scale_fill <- function(x) {scale_fill_manual(values=colDarj(x))}

pal_cspx <- function(x) {divergingx_hcl(x,palette = "Zissou 1")};
colBin<-c(pal_cspx(6)[3],pal_cspx(6)[5]); 
colRQ<-pal_cspx(4); blank<-alpha("#dae8ed",0.1); # show_col(colBin)
Scalc_rq<- function() {scale_colour_manual(values=colRQ)}
Scalf_rq2d <- function() {scale_fill_gradientn(colours=colRQ)} 
pal_cspx <- function(x) {divergingx_hcl(x,palette = "Zissou 1")}; # show_col(pal_cspx(6)) # colorspace
Scale_brew <- function()
  {scale_colour_brewer(palette="Spectral",aesthetics=c("colour","fill"))}
Scale_co <- function() 
  {scale_colour_manual(palette=colorRampPalette(brewer.pal(11, "Spectral")))}
Scale_fi <- function() 
  {scale_fill_manual(palette=colorRampPalette(brewer.pal(11, "Spectral")))}

pal_ggscc <- pal_material("teal"); # show_col(pal_ggscc(9)) # ggsci
colSum <- c(pal_ggscc(9)[3],pal_ggscc(9)[7],pal_ggscc(9)[1]) # show_col(colSum)
colInliers<-pal_material("teal")(10)[5] ; colOutliers = pal_material("deep-orange")(10)[5] # show_col(colOutliers)
Scalc_misc <- function() {scale_color_discrete_diverging(palette = "Blue-Red 3")}
Scale_map<- function() {scale_fill_distiller(palette = "Spectral")}
```

### Home made functions

```{r}
#| label: functmade
replace_outliers <- function(x, coefv=1.5) {
  replace(x, x %in% boxplot.stats(x, coef = coefv)$out, NA) #mean(x)
}

```
## Load of External data and Basic Variables
### External data

```{r}
#| label: externdata

rdatain <- sprintf("%sCSLN_Mars_BDD.RData",wdmat)
load(rdatain)
```

### Output binder and data

```{r}
#| label: outfiles

rdataout <- sprintf("%sCSLN_Mars_plot",wdmat)
# if exists
load(paste0(rdataout,".RData"))

```

### Basic variables

```{r}
#| label: basicvar

etude <- "CSLN_Mars"
espece <- "CERED"

sp <- which(speciesMP$SPCourt == espece)
sai <- which(saison$M_Def == "Year") #Year Winter Summer

```

# Material and methods

## Study area

The Seine estuary has been studied regarding several historically known areas that have either habitat or communities differences. These areas are mainly mudflats and subtidal areas, are represented in @fig-map.

```{r}
#| label: prefig-map

tmap_mode("plot") # "plot" "view"
# Extract data of interest
tm_areas <- ES_Areas %>% 
  dplyr::filter(!Zone  %in% c("Bay", "Ilot Oiseaux")) %>%
  st_transform(4326)

# Zoom tm_boxes
tm_boxestuary<-bb(tm_areas)
tm_box_mud <- ES_Areas %>%  
  dplyr::filter(grepl("Mudflat",ES_Areas$Zone)) %>%
  st_transform(4326) %>% bb(ext=2)
# tm_boxchannel<-bb(c(xmin=-2.5, ymin=49.0, xmax=1.5, ymax=51.0) )
tm_boxchannel<-bb(tm_boxestuary, ext=14)

# Extract Open street map background based on data

# osm_df <- read_osm(bb(Mars_SDM), ext=1, type= "osm") #stamen-watercolor osm
# osm_df_mud <- read_osm(tm_box_mud, ext=1, type= "osm") #stamen-watercolor osm
# osm_df_wat_bay <- read_osm(tm_boxchannel, ext=1, type= "stamen-watercolor")
# osm_df_wat_estu <- read_osm(tm_boxestuary, ext=1, type= "stamen-watercolor")
# save(osm_df,osm_df_mud,osm_df_wat_bay,osm_df_wat_estu, file=sprintf("SIG/OSM_extraction.RData"))
load(sprintf("SIG/OSM_extraction.RData"))

# create a map of context with red rect of area of interest
tm_bay<-
  tm_shape(osm_df_wat_bay) +  
    tm_rgb(saturation = .6, alpha= 0.8) +
  tm_shape(st_as_sfc(tm_boxestuary)) + 
    tm_borders("red", lwd = 2) +
  tm_layout(legend.show= FALSE) 
tm_estu<-tm_shape(osm_df_wat_estu) +  
    tm_rgb(saturation = .6, alpha= 0.8) +
  tm_shape(tm_areas) +  
    tm_fill(col = "Zone", title="Areas", palette = "Set2", alpha = 0.7) +
  tm_shape(st_as_sfc(tm_box_mud)) + 
    tm_borders("red", lwd = 2) +
  tm_layout(legend.show= FALSE)

# Create main map
tm_Bio<-
  tm_shape(osm_df_mud) +  
    tm_rgb(saturation = .6, alpha= 0.8) +
  tm_shape(tm_areas) +  
    tm_fill(col = "Zone", title="Areas", palette = "Set2", alpha = 0.6) +
    # tm_text("Zone", size = .8, col= "black") +
    tm_borders("grey", lwd = 1.5) +
  tm_shape(CSLN_sf) +
    tm_dots(col="grey30", size=0.1, alpha = 0.5) +
  tm_scale_bar(position = c("LEFT", "BOTTOM"), width = 0.15) + #SCALE
  tm_compass(type = "8star",position = c("LEFT", "TOP"), size = 2) +  #NORTH COMPASS
  tm_layout(#legend.outside = TRUE,
    legend.position= c("center","BOTTOM"),
    legend.title.size=.8,legend.text.size=.5,
    legend.bg.color="white",legend.bg.alpha=0.6)

# arrange maps insets thanks to patchwork !
tm_Bio_g<-tmap_grob(tm_Bio)
tm_bay_g<-tmap_grob(tm_bay)
tm_estu_g<-tmap_grob(tm_estu)
design <- c(
  area(t= 1, b= 14, l= 1, r= 20),
  area(t= 4, b= 6, l= 18, r= 20),
  area(t= 8, b= 11, l= 17, r= 20) )
tm_study_map<- wrap_plots(tm_Bio_g,tm_bay_g,tm_estu_g,design = design)

ggsave(filename= sprintf("%sfig-map.tiff",wdgraph), 
       tm_study_map, dpi=600, width = 10,  height = 7) #
```

```{r}
#| label: map_html
#| eval: false

tmap_mode("view") # "plot" "view"
tm_study_map_html<-
  tm_basemap(leaflet::providers$OpenStreetMap.HOT) +
  tm_shape(tm_areas) +  
    tm_fill(col = "Zone", palette = "Set2", alpha = 0.6, legend.show = FALSE) +
    # tm_text("Zone", size = .8, col= "black") +
    tm_borders("grey", lwd = 1.5) +
  tm_shape(CSLN_sf) +
    tm_dots(col="Zone", title="Areas", palette = "Set2", size=0.05, alpha = 0.9) +
  tm_layout(legend.show=FALSE)
tm_study_map_html
tmap_save(tm_study_map_html, filename = "SIG/Estuary_map.html")

```

## Predictors calculation and selection

```{r}
#| label: pl_mars_corr
#| include: true

# ggpairs custom colors on corr
corr_col <- function(data, mapping, method="p", use="pairwise", ...){
  # grab data
  x <- eval_data_col(data, mapping$x)
  y <- eval_data_col(data, mapping$y)
  # calculate correlation
  corr <- cor(x, y, method=method, use=use)
  # calculate colour based on correlation value
  colFn <- colorRampPalette(c(colSpec(2)[1], "white", colSpec(2)[2]), interpolate ='spline')
  fill <- colFn(100)[findInterval(corr, seq(-1, 1, length=100))]
  ggally_text(
    label = as.character(round(corr, 2)),
    mapping = aes(),
    xP = 0.5, yP = 0.5,
    ...) + #  ggally_cor(data = data, mapping = mapping, ...) + 
    theme_void() +
    theme(panel.background = element_rect(fill=fill))

} #wrap(cor_func,method = 'spearman', symbol = "Corr:\n")


df <- CSLN_mud %>%
  filter(SPCourt == espece) %>%
  select(paste(predict[,1],saison[sai,1],sep=""),
         map(reponse_l,~.x$rvar) %>% unlist(., use.names=FALSE))

pl_mars_corr <- ggpairs(df, labeller=label_wrap_gen(width = 10,multi_line = TRUE),# title=titreG,
              progress=FALSE,
              upper = list(continuous = corr_col),
              lower = list(continuous=
                             wrap("smooth", size = .3,
                                  alpha = 0.4, color = "aquamarine4"))) +
      theme(axis.text = element_text(size=8),
            strip.text.y = element_text(size=8))

ggsave(sprintf("%sfig-pl_mars_corr.tiff",wdgraph),
       plot = pl_mars_corr, width = 16, height = 9, dpi=600)

```

# Results

```{r}
#| label: data_set_description

data_descr <- function (dataset,vars) {
    # dataset=Mars_SDM %>% st_drop_geometry %>% filter(!Zone %in% c("Bay","Octeville"))
      # vars ="inunt"
  descr<-list(geog=list(stat.test=NULL,plot=NULL),
              time=list(stat.test=NULL,plot=NULL))

  descr$geog$stat.test <- dataset %>% 
    group_by(Zone) %>% 
    wilcox_test(formula(paste0(vars," ~ Period"))) %>%  #wilcox_test  levene_test???
    add_xy_position(x = "Period",group = "Zone",
                    dodge = 0,step.increase = 0.05) %>%
    filter(p.adj.signif!="ns")
  descr$geog$plot <- ggplot(dataset) + 
    # geom_boxplot(aes(x=Period, y = .data[[vars]], fill = Period)) +
    # geom_density_ridges(aes(y = Period , fill=Period,
    #                       x=.data[[vars]]), alpha=0.3) +
    geom_flat_violin(aes(x=Period, fill = Period,
                     y = .data[[vars]] ),
                     alpha = 0.5,size = .1,show.legend = FALSE) +
    stat_summary(aes(x=Period, fill = Period,
                     y = .data[[vars]], color = Period), 
                 size=.3, fun.data = mean_se, show.legend = FALSE) +
    stat_pvalue_manual(descr$geog$stat.test,label = "p.adj.signif",
                       size = 3,tip.length = 0, coord.flip = TRUE) +
    coord_flip() +
    facet_wrap( ~ Zone, 
               nrow = 1, #scales = "free_y", 
               labeller = label_wrap_gen(width = 20)) + 
    labs(y = vars, x="") +
    Scale_brew() +
    theme(legend.position="none",
          strip.background = element_blank())

  descr$time$stat.test <- dataset %>% 
    group_by(Period) %>%
    wilcox_test(formula(paste0(vars," ~ Zone"))) %>% 
    add_xy_position(x = "Zone",group = "Period",
                    dodge = 0,step.increase = 0.05) %>%
    filter(p.adj.signif!="ns")

  descr$time$plot <- ggplot(dataset) + 
    # geom_boxplot(aes(x=Zone, y = .data[[vars]], fill = Zone)) +
    geom_flat_violin(aes(x=Zone, fill = Zone,
                     y = .data[[vars]] ),
                     alpha = 0.5,size = .1,show.legend = FALSE) +
    stat_summary(aes(x=Zone, fill = Zone,
                     y = .data[[vars]], color = Zone),  
                 size=.3, fun.data = mean_se, show.legend = FALSE) +
    stat_pvalue_manual(descr$time$stat.test,label = "p.adj.signif", 
                       size = 3,tip.length = 0, coord.flip = TRUE) +
    coord_flip() +
    facet_wrap( ~ Period, 
               nrow = 1, #scales = "free_y", 
               labeller = label_wrap_gen(width = 20)) + 
    labs(y = vars, x="") +
    Scale_brew() +
    guides(fill=guide_legend(nrow=2,byrow=TRUE)) +
    theme(legend.position="none",
          strip.background = element_blank())
  
  return(descr)
}

```

## Biological data set description

::: column-screen-inset-right

```{r}
#| label: biom_distr_bd
CSLN_df <- CSLN %>%
  filter(grepl("Mudflat", Zone)) %>% # focus on mudflat only
  filter(SPCourt %in% c(espece)) %>%
  filter(!is.na(Biomass_gAFDWm2)) %>%
  mutate(across(where(is.factor), fct_drop)) %>% # recode predict when filter has remove levels
  group_by(idStationUnique,Zone,Tidal_level,Period,Annee) %>% 
  summarise(iZTPAS = sum(Biomass_gAFDWm2, na.rm =TRUE)) %>% 
  group_by(Zone,Period,Tidal_level,Annee) %>% #,Season
  summarise(Biomass_gAFDWm2 = mean(iZTPAS, na.rm =TRUE)) %>% 
  mutate(Period=factor(Period, 
                      labels = str_replace_all(unique(Period),"-"," - ")))

pl_des_b <- data_descr(CSLN_df,"Biomass_gAFDWm2")

CSLN_df <- CSLN %>%
  filter(grepl("Mudflat", Zone)) %>% # focus on mudflat only
  filter(SPCourt %in% c(espece)) %>%
  filter(!is.na(Density_indm2)) %>%
  mutate(across(where(is.factor), fct_drop)) %>% # recode predict when filter has remove levels
  group_by(idStationUnique,Zone,Tidal_level,Period,Annee) %>% 
  summarise(iZTPAS = mean(Density_indm2, na.rm =TRUE)) %>% 
  group_by(Zone,Period,Tidal_level,Annee) %>% #,Season
  summarise(Density_indm2 = mean(iZTPAS, na.rm =TRUE)) %>% 
  mutate(Period=factor(Period, 
                      labels = str_replace_all(unique(Period),"-"," - ")))

pl_des_d <- data_descr(CSLN_df,"Density_indm2")

# pl_des_b_gt <- ggarrange(pl_des_b$geog$plot,pl_des_b$time$plot, ncol=1, nrow=2,
#                 labels="AUTO", legend="bottom",
#                 common.legend = FALSE)
# pl_des_b_gt <- annotate_figure(pl_des_b) + #top = text_grob(title, face = "bold", size = 14)
#       bgcolor("white")

pl_des_b_gt <- pl_des_b$geog$plot / pl_des_b$time$plot & 
    theme(legend.position="bottom",
          text=element_text(size=12)) &
    plot_annotation(tag_levels = c('A'))
pl_des_d_gt <- pl_des_d$geog$plot / pl_des_d$time$plot  & 
    theme(legend.position="bottom",
          text=element_text(size=12)) &
    plot_annotation(tag_levels = c('A'))

ggsave(sprintf("%sfig-cered_struc_biom.tiff",wdgraph),
       plot = pl_des_b_gt, width = 10, height = 5, dpi=400)
ggsave(sprintf("%sfig-cered_struc_dens.tiff",wdgraph),
       plot = pl_des_d_gt, width = 10, height = 5, dpi=400)
```
:::

## HMS data set description

```{r}
#| label: hms_distr

pl_mars <- 
  map(setNames(as.list(Var_choosen),Var_choosen), \(vars)
          data_descr(Mars_SDM %>%
                       st_drop_geometry %>%
                       # mutate(across(where(is.numeric), ~replace_outliers(.,2))) %>% 
                       mutate(across(all_of(c("bathy","tenfon_mxd")), ~replace_outliers(.,1.5))) %>%
                       filter(!Zone %in% c("Bay","Octeville","Ilot Oiseaux","Cote Fleurie")) %>%
                       mutate(across(where(is.factor), fct_drop)), # recode predict when filter has remove levels
                     vars) )

pl_mars_gt <- map(pl_mars,
                  ~{.x$geog$plot / .x$time$plot  & 
    theme(legend.position="bottom",
          text=element_text(size=12)) &
    plot_annotation(tag_levels = c('A')) } )

iwalk(pl_mars_gt,
  ~ggsave(sprintf("%sfig-hms_struc_%s.tiff",wdgraph,.y),
       plot = .x, width = 10, height = 5, dpi=400) )

```

# Final actions and save

Rdata are saved in `rdataout`. An excel file collects data in `r binderout` , with sheets for :

-   data : contains whatever

```{r}
#| label: finalsave_r

save(list=(ls()[grep("^tm_|^pl_", ls())]),
     file = paste0(rdataout,".RData"))
beepr::beep(2)
```

# Supplementary data

## Software details

::: {.callout-tip collapse="true"}
## Expand for R Session Info

```{r, echo = FALSE}
#| include: true

library(sessioninfo)
pkg_sesh <- session_info(pkgs = "attached")
quarto_version <- system("quarto --version", intern = TRUE)
pkg_sesh$platform$quarto <- paste(
  system("quarto --version", intern = TRUE),
  "@",
  quarto::quarto_path()
  )
pkg_sesh
```
:::

# References {.unnumbered}