---
title: "CREATION DE GRAPHIQUES DESCRIPTIFS DES DONNEES PROGRAMME REPONSES PREDICTEURS"
author: "Amélie Lehuen"
description: "add info on script"
date: last-modified
editor: source
execute:
  eval: true
  echo: false
  message: false
  warning: false
  output: true
  include: false
editor_options: 
  chunk_output_type: console
---

# CREATION DE GRAPHIQUES DESCRIPTIFS DES DONNEES PROGRAMME REPONSES PREDICTEURS

## Script preparation

### Packages

```{r}
#| label: load-packages
#| code-summary: "Packages"

library(tidyverse); library(beepr)
library(rstatix)
library(ggpubr) ; # stat_compare_means ggarrange
library(RColorBrewer) ; library(wesanderson) 
library(treemap)
library(knitr)
```

### Working Environment

```{r}
#| label: workenvir

rm(list=ls())

wdsource <- "Data/Faune/CSLN/"
wdmat <- "Matrices/"
wdgraph <- "Plots/"
wdres <- "Results/"

```

### Graphic charter

```{r}
#| label: graphchart

theme_set(theme_bw(base_size = 16)) # theme_gray() theme_bw() theme_light() theme_dark()
colDarj <- function(x) {wes_palette("Darjeeling2",x, type = "continuous")}
colZiss <- function(x) {wes_palette("Zissou1",x, type = "continuous")}
colSpec <- colorRampPalette(brewer.pal(8, "Spectral")); 
colDark <- colorRampPalette(brewer.pal(8, "Dark2"));
Scale_col <- function(x) {scale_colour_manual(values=colDarj(x))}
Scale_fill <- function(x) {scale_fill_manual(values=colDarj(x))}
Scale_brew <- function() {scale_colour_brewer(palette="Spectral",aesthetics=c("colour","fill"))}

```

### Home made functions

```{r}
#| label: functmade

# FONCTION boxplot with facets
boxplot_amlh <- function(df,x,y,fil,face,xlab,ylab,title,palette){
  ggplot(df) + 
    geom_boxplot(aes(x=x,y=y,fill=fil)) + 
    facet_wrap(face) + # scales="free_y", margins=TRUE 
    scale_fill_manual(values=palette) +
    labs(title=title, x=xlab, y=ylab, fill="") +
    theme(legend.position="bottom",
          axis.text.x=element_text(angle=-90, vjust=0.4,hjust=1))
}
#FONCTION Bargraph with error bars and facets
bargraph_sd <- function(df,x,y,sd,fil,face,xlab,ylab,title,palette){
  empilement = .9 # .9 pour des groupes par zone cote cote, .6 pour empilement
  ggplot(df, aes(x=x, y=y, fill=fil)) +
    geom_bar(stat="identity", position=position_dodge(width=empilement)) +
    geom_errorbar(aes(ymin=y-sd, ymax=y+sd),
                  width=.2,position=position_dodge(empilement)) +
    labs(title=title,x=xlab, y=ylab, fill="") +
    facet_wrap(face) +
    theme(legend.position="bottom",legend.title = element_blank()) +
    scale_fill_manual(values=palette)
}
```

## Load of External data and Basic Variables

### External data

```{r}
#| label: externdata

rdatain <- sprintf("%sCSLN_Mars_Zone_BDD.RData",wdmat)
load(rdatain)
```

### Output binder and data

```{r}
#| label: outfiles

rdataout <- sprintf("%sCSLN_Mars_Zone_BDD.RData",wdmat)

```

### Basic variables

```{r}
#| label: basicvar

prgm <- "CSLN"
etude <- paste(prgm,"_Zone",sep="")

spe <- 1#:nrow(speciesMP) # 1:CERED 2:CORVO 3:HEDDI 4:LIMBA 5:PERUL 6:SCRPL
reponse<-reponse[1:2,] # 1:Biomass_gAFDWm2 2:Density_indm2
answ <- 1#:nrow(reponse)

```

# GLOBAL VISION OF DATA

Nb record per speciesMP

```{r}
#| label: global_1
#| include: true

dp<-CSLN %>% count(SP,SPCourt,Taxon_SNa) %>% group_by(SP) %>%
              arrange(desc(n),.by_group = TRUE) %>% slice(1:20) %>%
              ggplot(aes(x=reorder(Taxon_SNa,n),y=n,fill = SP))+
              geom_col() +
              geom_text(aes(y=60, label = n)) + #
              scale_fill_manual(values=colDarj(CSLN_unique$Annee)) +
              labs(title="List of top 20 occurences speciesMP", 
                   x="Species",
                   y="Occurence") +
              coord_flip() +
              theme(axis.text = element_text(size=15,face="bold")) +
              theme_bw()
print(dp)
ggsave(sprintf("%s%s_Species_Occr_List20.tiff",wdgraph,etude), 
       plot = dp, width = 10, height = 8, dpi=400)

```

Treemap 2 levels : Tidal levels distribution

```{r}
#| label: global_2

titre = paste("Total Density (ind.m-2) of Species"," - ", etude,sep="")
Sumry<-CSLN_mud %>% 
  group_by(Zone,SPCourt) %>% #Tidal_level
  summarise(Smry=sum(Density_indm2,na.rm=TRUE),Nb=n()) #n_distinct(idStationUnique)
Sumry$label <- paste(Sumry$SPCourt, Sumry$Nb, sep = "\n")
png(filename=sprintf("%s%s_TreeMap2 Zone Level Density.png",
                     wdgraph,etude),
    width=6000, height=3600, res = 600)
treemap(Sumry,index=c("Zone","label"),
        vSize="Smry",vColor="Zone",type="index",
        title=titre, 
        palette=colSpec(CSLN_unique$Zone), 
        fontface.labels=c(2,1),
        border.col=c("black","white"), 
        border.lwds=c(3,1), bg.labels=0,
        fontsize.labels=c(12,9), 
        fontcolor.labels=c("black","black"),
        align.labels=list( c("left","bottom"), c("right","center")))
dev.off()

titre = paste("Total Biomass (gAFDW.m-2) of Species"," - ", etude,sep="")
Sumry<-CSLN_mud %>% 
  group_by(Zone,SPCourt) %>% 
  summarise(Smry=sum(Biomass_gAFDWm2,na.rm=TRUE),Nb=n()) #n_distinct(idStationUnique)
Sumry$label <- paste(Sumry$SPCourt, Sumry$Nb, sep = "\n")
png(filename=sprintf("%s%s_TreeMap2 Zone Level Biomass.png",
                     wdgraph,etude),
    width=6000, height=3600, res = 600)
treemap(Sumry,index=c("Zone","label"),
        vSize="Smry",vColor="Zone",type="index",
        title=titre, 
        palette=colSpec(CSLN_unique$Zone), 
        fontface.labels=c(2,1),
        border.col=c("black","white"), 
        border.lwds=c(3,1), bg.labels=0,
        fontsize.labels=c(12,9), fontcolor.labels=c("black","black"),
        align.labels=list( c("left","bottom"), c("right","center")))
dev.off()

```

Treemap 3 levels: Distribution spatial

```{r}
#| label: global_3

titre = paste("Total Density (ind.m-2) of Species"," - ", etude,sep="")
Sumry<-CSLN_mud %>% 
  group_by(Zone,Tidal_level,SPCourt) %>%
  summarise(Smry=sum(Density_indm2,na.rm=TRUE),Nb=n())
Sumry$label <- paste(Sumry$SPCourt, Sumry$Nb, sep = "\n")
png(filename=sprintf("%s%s_TreeMap3 Zone Density.png",
                     wdgraph,etude),
    width=6000, height=3600, res = 600)
treemap(Sumry,index=c("Zone","Tidal_level","label"),
        vSize="Smry",vColor="Zone",type="index",
        title=titre, 
        palette=colSpec(CSLN_unique$Zone), 
        fontface.labels=c(2,2,1),
        border.col=c("black","white","gray81"), 
        border.lwds=c(4,3,1),bg.labels=0,
        fontsize.labels=c(14,14,10), 
        fontcolor.labels=c("black","black","gray81"),
        align.labels=list( c("left","bottom"), c("left","top"), c("right","center")))
dev.off()

```

#### GRAPHICS OF MSR AND BPC PER PERIOD FACET MUDFLATS with significance stars

```{r}
#| label: global_4

stat.test <- CSLN_mud_BTA %>% 
  group_by(Zone) %>% 
  wilcox_test(MSR.ZTPAS ~ Period) %>% 
  add_xy_position(x = "Period",group = "Zone",
                  dodge = 0,step.increase = 0.05) %>% 
  mutate(xmin=xmin-1,xmax=xmax-1) %>%
  filter(p.adj.signif!="ns")
gpA <- ggplot(CSLN_mud_BTA) + 
  geom_boxplot(aes(x=Period, y = MSR.ZTPAS, fill = Period)) + 
  # coord_cartesian(ylim = c(0, 100)) +
  facet_grid(. ~ Zone, scales = "free_y") + #facet_grid(. ~ Zone # facet_wrap(vars(Zone)
  labs(y = "MSR (mW/m2)",x="") +
  stat_pvalue_manual(stat.test,label = "p.adj.signif", tip.length = 0.02) +
  Scale_brew() +
  theme(axis.text.x = element_blank(),axis.ticks = element_blank())
print(gpA)

stat.test <- CSLN_mud_BTA %>% 
  group_by(Zone) %>% 
  wilcox_test(BPc.ZTPAS ~ Period) %>% 
  add_xy_position(x = "Period",group = "Zone",
                  dodge = 0,step.increase = 0.05) %>% 
  mutate(xmin=xmin-1,xmax=xmax-1) %>% 
  filter(p.adj.signif!="ns")
gpB <- ggplot(CSLN_mud_BTA) + 
  geom_boxplot(aes(x=Period, y = BPc.ZTPAS, fill = Period)) + 
  # coord_cartesian(ylim = c(0, 100)) +
  facet_grid(. ~ Zone, scales = "free_y") + 
  labs(y = "BPc",x="") +
  stat_pvalue_manual(stat.test,label = "p.adj.signif", tip.length = 0.02) +
  Scale_brew() +
  theme(axis.text.x = element_blank(),axis.ticks = element_blank())
print(gpB)

titre <- paste("Life traits indexes",sep="")
bp <- ggarrange(gpA,gpB, ncol=1, nrow=2, 
                labels="AUTO", legend="bottom", 
                common.legend = TRUE)
bp <- annotate_figure(bp,top = text_grob(titre, face = "bold", size = 14)) + #,bottom="Period"
      bgcolor("white")
print(bp)
 
ggsave(sprintf("%s%s %s.tiff",wdgraph,etude,titre), 
       plot = bp, width = 10, height = 8, dpi=400)

# # GENERAL BILAN
# CSLN_mud_sm <- CSLN_mud %>%
#   group_by(Taxon_SNa) %>%
#   summarise(Density_indm2_m = mean(Density_indm2)) %>%
#   arrange(desc(Densite_moyenne)) %>%
#   slice(1:20)

```

BOXPLOT BY PERIOD AND FACET ZONE

```{r}
#| label: bxp_1

titre <- paste("Periodic Evolution of Specific Richness",sep="")
bp<-boxplot_amlh(CSLN_sm,CSLN_sm$Period,CSLN_sm$SR
                 ,CSLN_sm$Period,CSLN_sm$Zone,
                 "Years","Specific Richness",
                 titre,
                 colDarj(CSLN_unique$Zone))
print(bp)
ggsave(sprintf("%s%s %s.tiff",wdgraph,etude,titre),
       plot = bp, width = 12, height = 6, dpi=400)
bp1<-bp + 
  labs(title="Specific Richness",x="",y="") + 
  theme(legend.position='none')

titre <- paste("Periodic Evolution of Pielou",sep="")
bp<-boxplot_amlh(CSLN_sm,CSLN_sm$Period,CSLN_sm$Pielou,
                 CSLN_sm$Period,CSLN_sm$Zone,
                 "Years","Index of Equitability",titre,
                 colDarj(CSLN_unique$Zone))
print(bp) 
ggsave(sprintf("%s%s %s.tiff",wdgraph,etude,titre), 
       plot = bp, width = 12, height = 6, dpi=400)
bp2<-bp + 
  labs(title="Index of Equitability",x="",y="") + 
  theme(legend.position='none')

titre <- paste("Periodic Evolution of Total Biomass",sep="")
bp<-boxplot_amlh(CSLN_sm,CSLN_sm$Period,CSLN_sm$Biomass_t,
                 CSLN_sm$Period,CSLN_sm$Zone,
                 "Years","Total Biomass (gAFDW/m2)",
                 titre,
                 colDarj(CSLN_unique$Zone)) +
                ylim(0,750)
print(bp) 
ggsave(sprintf("%s%s %s.tiff",wdgraph,etude,titre),
       plot = bp, width = 12, height = 6, dpi=400)
bp3<-bp + 
  labs(title="Total Biomass (gAFDW/m2)",x="",y="") + 
  theme(legend.position='none')

titre <- paste("Periodic Evolution of Metabolic Rate",sep="")
bp<-boxplot_amlh(CSLN_sm,CSLN_sm$Period,CSLN_sm$MSRtot_t,
                 CSLN_sm$Period,CSLN_sm$Zone,
                 "Years","Total MSRtot (mW/m2)",
                 titre,
                 colDarj(CSLN_unique$Zone))
print(bp) 
ggsave(sprintf("%s%s %s.tiff",wdgraph,etude,titre),
       plot = bp, width = 12, height = 6, dpi=400)
bp4<-bp + 
  labs(title="Total MSRtot (mW/m2)",x="",y="") + 
  theme(legend.position='none')

titre <- paste("Periodic Evolution of Biodiversity",sep="")
bp <- ggarrange(bp1,bp2,bp3,bp4, ncol=2, nrow=2, 
                labels="AUTO", legend="bottom", 
                common.legend = TRUE)
bp <- annotate_figure(bp,top = text_grob(titre, face = "bold", size = 14),
                      bottom=xlab) + 
  bgcolor("white")
print(bp)
ggsave(sprintf("%s%s %s.tiff",wdgraph,etude,titre),
       plot = bp, width = 15, height = 15, dpi=400)


```

BARPLOT BY PERIOD AND FACET ZONE

```{r}
#| label: brp_1

# df  <-  CSLN_sm %>% unite(AS,Annee,Season,sep = "_", remove = FALSE)
df <- CSLN_sm %>% 
  group_by(Zone,Period) %>%
  summarise(SR_m=median(SR,na.rm=TRUE),
            SR_sd=sd(SR,na.rm=TRUE),
            Biomass_t_m=median(Biomass_t,na.rm=TRUE),
            Biomass_t_sd=sd(Biomass_t,na.rm=TRUE),
            Pielou_m=median(Pielou,na.rm=TRUE),
            Pielou_sd=sd(Pielou,na.rm=TRUE),
            MSRtot_t_m=median(MSRtot_t,na.rm=TRUE),
            MSRtot_t_sd=sd(MSRtot_t,na.rm=TRUE))
xlab <- ""
titre <- paste("Temporal Evolution of Specific Richness",sep="")
bp <-bargraph_sd(df,df$Period,df$SR_m,df$SR_sd,
                 df$Period,df$Zone,
                 xlab,"Specific Richness",
                 titre,
                 colDarj(CSLN_unique$Zone)) + 
    theme(axis.text.x=element_text(angle=90, 
                                   vjust=0.4,hjust=1))
print(bp)
ggsave(sprintf("%s%s %s.tiff",wdgraph,etude,titre), 
       plot = bp, width = 12, height = 6, dpi=400)
bp1<-bp + 
  labs(title="Specific Richness",y="") + 
  theme(legend.position='none')

titre <- paste("Temporal Evolution of Density",sep="")
bp <-bargraph_sd(df,df$Period,df$Biomass_t_m,df$Biomass_t_sd,
                 df$Period,df$Zone,
                 xlab,"Total Biomass (gAFDW/m2)",
                 titre,
                 colDarj(CSLN_unique$Zone)) + 
  theme(axis.text.x=element_text(angle=90, 
                                 vjust=0.4,hjust=1))
print(bp)
ggsave(sprintf("%s%s %s.tiff",wdgraph,etude,titre),
       plot = bp, width = 12, height = 6, dpi=400)
bp2<-bp + 
  labs(title="Total Biomass (gAFDW/m2)",y="") + 
  theme(legend.position='none')

titre <- paste("Temporal Evolution of Pielou index",sep="")
bp <-bargraph_sd(df,df$Period,df$Pielou_m,df$Pielou_sd,
                 df$Period,df$Zone,
                 xlab,"Index of Equitability",
                 titre,
                 colDarj(CSLN_unique$Zone)) + 
  theme(axis.text.x=element_text(angle=90, 
                                 vjust=0.4,hjust=1))
print(bp)
ggsave(sprintf("%s%s %s.tiff",wdgraph,etude,titre),
       plot = bp, width = 12, height = 6, dpi=400)
bp3<-bp + 
  labs(title="Index of Equitability",y="") + 
  theme(legend.position='none')

titre <- paste("Temporal Evolution of metabolic rate",sep="")
bp <-bargraph_sd(df,df$Period,df$MSRtot_t_m,df$MSRtot_t_sd,
                 df$Period,df$Zone,
                 xlab,"Total MSRtot (mW/m2)",
                 titre,
                 colDarj(CSLN_unique$Zone)) + 
  theme(axis.text.x=element_text(angle=90, 
                                 vjust=0.4,hjust=1))
print(bp)
ggsave(sprintf("%s%s %s.tiff",wdgraph,etude,titre),
       plot = bp, width = 12, height = 6, dpi=400)
bp4<-bp + 
  labs(title="Total MSRtot (mW/m2)",y="") + 
  theme(legend.position='none')

# Un plot pour les 3 differents graphiques
titre=paste("Temporal Evolution of Benthic fauna",sep="")
bp <- ggarrange(bp1,bp2,bp3,bp4, ncol=2, nrow=2, 
                labels="AUTO", legend="bottom", common.legend = TRUE)
bp <- annotate_figure(bp,top = text_grob(titre, face = "bold", size = 14),
                      bottom=xlab) + 
  bgcolor("white")
print(bp)
ggsave(sprintf("%s%s %s.tiff",wdgraph,etude,titre), 
       plot = bp, width = 15, height = 15, dpi=400)


```

## CHOSEN SPECIES FOCUS

```{r}
#| label: focus_1

CSLN_mud_spm <- CSLN_mud_spm %>% 
  filter(SPCourt %in% speciesMP$SPCourt)
titre <- paste("Periodic Evolution of Biomass for chosen speciesMP",sep="")
bp<-ggplot(CSLN_mud_spm) +
  geom_boxplot(aes(x=Period,y=Biomass_m,fill=Period)) + 
  facet_grid(SPCourt~Zone) + # Period, scales="free_y", margins=TRUE
  scale_y_continuous(trans = "log10") + 
  theme(legend.position="bottom",
        axis.text.x=element_blank()) + #element_text(angle=-90, vjust=0.4,hjust=1)) +
  labs(title=titre,y="Biomass (gAFDW/m2)") +
  Scale_fill(CSLN_unique$Annee)
print(bp)
ggsave(sprintf("%s%s %s.tiff",wdgraph,etude,titre), 
       plot = bp, width = 10, height = 10, dpi=400)


```

Boxplot des reponses par annee pour 1 esp avec facet

```{r}
#| label: focus_2

for (sp in spe){ #sp=1
  df <- CSLN_mud[which(CSLN_mud$SPCourt == speciesMP$SPCourt[sp]),]
  mlist<-vector(mode = "list", length = length(answ))
  for (rep in answ){ # rep=3
    y <- reponse[rep,1]
    titreG <- sprintf("%s time evolution for %s - %s",
                      y,speciesMP$Taxon_SNa[sp], etude)
    bp <- ggplot(df, aes(Annee,.data[[y]],fill=Period)) +
      geom_boxplot() + 
      facet_wrap(~Zone) + #,scales="free_y"
      theme(axis.text.x=element_text(angle=-90, vjust=0.4,hjust=1)) +
      labs(title=titreG, fill="")+
      theme(legend.position="bottom",
            legend.title = element_blank())+
      scale_y_continuous(trans = "log10") + 
      annotation_logticks(sides="lr") +
      Scale_fill(CSLN_unique$Annee)
    print(bp)
    ggsave(sprintf("%s%s/%s_Bxp_%s_%s.tiff",
                   wdgraph,speciesMP$SPCourt[sp],etude,
                   speciesMP$SPCourt[sp],y), 
           plot = bp, width = 12, height = 6, dpi=400)
    bp <- bp + 
      labs(title=sprintf("%s (%s)",reponse[rep,2],reponse[rep,3]), y="") +
      theme(legend.position='none')
    
    mlist[[rep]] <- bp + theme(legend.position="none");
  }
  
  if (length(answ)>1 ){
    titreG <- sprintf("Responses time evolution for %s - %s",
                      speciesMP$Taxon_SNa[sp], etude)
    nncol = length(answ); nnrow=ceiling(length(answ)/nncol)
    bp <- ggarrange(plotlist=mlist, ncol=nncol, nrow=nnrow, 
                    labels="AUTO",legend="bottom",common.legend = TRUE)
    bp <- annotate_figure(bp, 
                          top = text_grob(titreG, face = "bold", size = 14))+
      bgcolor("white")
    print(bp)
    
    ggsave(sprintf("%s%s/%s_Bxp_%s_Ans.tiff",
                   wdgraph,speciesMP$SPCourt[sp],
                   etude,speciesMP$SPCourt[sp]),
           plot = bp, width = 16, height = 6, dpi=400)
  }}

```

Boxplot des Densites moy par annee pour 1 esp avec facet

```{r}
#| label: focus_3

# for (sp in spe){ 
#   df <- CSLN_sp[which(CSLN_sp$SPCourt == speciesMP[sp,1]),]
#   titreG <- paste("Time evolution for ",speciesMP[sp,2], " - ", etude,sep="")
#   bp <- ggplot(df, aes(Annee,Density_indm2,fill=Period)) + geom_boxplot() + 
#     facet_wrap(~Zone) +
#     labs(title=titreG) + 
#     scale_y_continuous(trans = "log10") + annotation_logticks(sides="lr") + 
#     theme(axis.text.x=element_text(angle=-90, vjust=0.4,hjust=1)) +
#     Scale_fill(CSLN_unique$Annee); bp
#   ggsave(paste(wdgraph,speciesMP[sp,1],"/",etude,"_Bxp1Zone_",speciesMP[sp,1],"_","Density_indm2",".png",sep=""), plot = bp, width = 12, height = 6, dpi=400)
# }

```

# Final actions and save

Rdata are saved in `rdataout`.

```{r}
#| label: finalsave_xls

```

```{r}
#| label: finalsave_r

beepr::beep(2)
```

# Supplementary data

## Software details

::: {.callout-tip collapse="true"}
## Expand for R Session Info

```{r, echo = FALSE}
#| include: true

library(sessioninfo)
pkg_sesh <- session_info(pkgs = "attached")
quarto_version <- system("quarto --version", intern = TRUE)
pkg_sesh$platform$quarto <- paste(
  system("quarto --version", intern = TRUE),
  "@",
  quarto::quarto_path()
  )
pkg_sesh
```
:::

# References {.unnumbered}
