---
title: "CREATION DE CARTES DU LIEU D'ETUDE ET DES RESULTATS"
author: "Amélie Lehuen"
description: "add info on script"
date: "`r format(Sys.time(), '%B %Y')`"
editor: source
execute:
  eval: true
  echo: false
  message: false
  warning: false
  output: true
  include: false
editor_options: 
  chunk_output_type: console
---

# CREATION DE CARTES DU LIEU D'ETUDE ET DES RESULTATS

## Script preparation

### Packages

```{r}
#| label: load-packages
#| code-summary: "Packages"
#| include: false

library(tidyverse) ; library(RColorBrewer)
library(sf) ; library(sfheaders); #library(raster)
library(dplyr);library(ggpubr) ; library(grid) ; library(gridExtra)
library(plotly); library(plot3D);  
# library(leaflet) # for interactive maps
library(tmap) # tmap_mode; for static and interactive maps
library(htmlwidgets) # library(leaflet) # saveWidget ; for interactive maps
library(rnaturalearth) # ne_states install_github("ropensci/rnaturalearthhires")
# library(spData) ; library(spDataLarge) #remotes::install_github("Nowosad/spDataLarge")
# library(raster) 

options(scipen=999) # Empeche affichage scientifique des nombres
library(knitr)
```

### Working Environment

```{r}
#| label: workenvir
#| echo: true

rm(list=ls())
# tsk <- "A_SDM_NEO/"
wdtask <- "./" #paste(pc,"Melting Pot/BDD/",tsk,sep="")
wdwork <- paste(wdpath,"Matrices/",sep="")
wdgraph <- paste(wdpath,"Graphiques/",sep="")
wdres <- paste(wdpath,"Resultats/",sep="")
setwd(wdpath)

pc <- "C:/Users/lehuen201/Nextcloud/" # "E:/" #
wdgraphEx<-wdgraph #paste(pc,"Copie-HD/Melting Potes/",tsk,"Graphiques/",sep="")
wdGIS <- paste(pc,"Melting Pot/SIG/",sep="")
```

### Graphic charter

```{r}
#| label: graphchart
theme_set(theme_bw(base_size = 16)) # theme_gray() theme_bw() theme_light() theme_dark()
colorsS <- colorRampPalette(brewer.pal(8, "Spectral")); coldiscS <- function(x) {scale_colour_manual(values=colorsS(x))}
colorsD <- colorRampPalette(brewer.pal(8, "Dark2")); coldiscD <- function(x) {scale_colour_manual(values=colorsD(x))}

```

### Home made functions

```{r}
#| label: functmade
#| echo: true

```

## Load of External data and Basic Variables

### External data

```{r}
#| label: externdata
#| echo: true
binderin <- sprintf("./Model Base Script Data.xlsx")
data <- as.data.frame(read_excel(binderin,sheet = "Exp_Plan", na = ""))
rdatain <- sprintf("%sCSLN_Mars_BDD.RData",wdwork)
load(rdatain)
```

### Output binder and data

```{r}
#| label: outfiles
#| echo: true
binderout <- sprintf("binderout.xlsx")
rdataout  <- sprintf("dataout.xlsx")
# if exists
load(rdataout)
```

### Basic variables

```{r}
#| label: basicvar
#| echo: true

prgm <- "CSLN" # 1:CSLN 2:Mabes 3:Geco 4:Beaug
etude <- "CSLN_Mars"


```

# SHP CREATION OF FAUNA & MARS DATA

GIS HTML MAP CREATION \-\-\--

https://thinkr.fr/sil-te-plait-dessine-moi-carte-r/

https://thinkr.fr/cartographie-interactive-comment-visualiser-mes-donnees-spatiales-de-maniere-dynamique-avec-leaflet/

https://thinkr.fr/cartographie-interactive-avec-r-la-suite/

https://cran.r-project.org/web/packages/tmap/vignettes/tmap-getstarted.html

https://bookdown.org/nicohahn/making_maps_with_r5/docs/tmap.html

```{r}
#| label: calculation_1
#| include : false # necessary to avoid fluffy doc

CSLN_sf <- st_as_sf(CSLN_pur,coords= c("longitude","latitude"),crs=4326,remove = FALSE) #c("x","y"),crs=2154,remove = FALSE) %>% st_transform(4326) #"Lon","Lat" in WGS=4326
Sel_col<-c(facto_col,"Density_indm2","Biomass_gAFDWm2",
           "IndBodySize_gAFDW","MSR_mW","MSRtot") #,"Itot"
CSLN_sf <- CSLN_sf %>% select(all_of(Sel_col)) # lighten the data to help build the map

# boxbds=c(xmin=-0.1, ymin=49.3, xmax=0.45, ymax=49.7)
# bds<-st_crop(ES_Areas,boxbds, crs = st_crs(4326)) %>% # decoupe zone interet
#   st_transform(2154) # transform to planar as required by st_intersection()
# bds_points <- cbind(bds, st_coordinates(st_centroid(bds$geometry))) # def centroides pour placer les noms
# # pol1 = st_polygon(list(rbind(c(0,0),c(1,0),c(1,1),c(0,1),c(0,0))))

```

```{r}
#| label: calculation_2
#| include : false # necessary to avoid fluffy doc

tmap_mode("view") #tmap_mode("plot") # ttm() toggle #
tm_Bio<-tm_basemap("OpenStreetMap.HOT") + # OpenStreetMap.HOT .Mapnik .France - Stamen.Watercolor #https://leaflet-extras.github.io/leaflet-providers/preview/
  # tm_logo(paste(wdlogos,"logo-MIE.png",sep=""), height = 2) +
  # tm_logo(c(paste(wdlogos,"logoMP.png",sep=""),
  #           paste(wdlogos,"Logo Borea.png",sep="")), height = 2) +
  tm_scale_bar(position = c("left", "bottom"), width = 0.15)+ #SCALE
  tm_compass(position = c("left", "top"), size = 2)+          #NORTH COMPASS
  tm_shape(ES_Areas) +
  tm_fill(col = "Zone", palette = "Spectral", alpha = 0.6) +
  tm_borders("white", lwd = 1) +
  tm_shape(CSLN_sf) +
  # tm_bubbles(size="Density_indm2") +
  tm_dots(size=0.001) +
  tm_layout(legend.outside = TRUE)
# tm_Bio # WARNING THIS OPERATION CAN LAST LONG
tmlf_Bio<-tmap_leaflet(tm_Bio) # conversion to leaflet object, quicker??
# tmlf_Bio

# # Simple and basic map of France
# france <- ne_states(country = "France", returnclass = "sf") %>%
#   filter(region %in% c("Haute-Normandie","Basse-Normandie"))
# map_france<-tm_shape(france) + tm_polygons()#+
# # tm_shape(mybb)+tm_borders("white", lwd = 1) # ajout du cadre rouge sur la zone interet
# print(map_france, vp = grid::viewport(0.9, 0.7, width = 0.2, height = 0.2))
# # map_france<-qtm("france") # creation tres rapide d'une carte

```

### GIS SAVE

```{r}
#| label: calculation_3
#| include : false # necessary to avoid fluffy doc

write.csv(CSLN_Mars,file=paste(wdres,"CSLN_Mars", ".csv",sep=""), na = "",row.names = FALSE)
write.csv(CSLN_pur,file=paste(wdres,"CSLN_data", ".csv",sep=""), na = "",row.names = FALSE)
# WARNING FOLLOWING OPERATION CAN LAST LONG
tmap_save(tm_Bio, filename = paste(wdres,'CSLN_Mars Map.html',sep=""))
# saveWidget(tmlf_Bio, paste(wdres,'CSLN_Mars Map.html',sep=""), selfcontained = TRUE)

```

OPTION LEAFLET

```{r}
#| label: calculation_4
#| include : false # necessary to avoid fluffy doc

# pal <- colorFactor(
#   palette = "viridis",na.color = NA,
#   levels = factor(bds$Type))
# map_leaflet <- leaflet() %>%
#   addProviderTiles("OpenStreetMap.HOT") %>% #addTiles()
#   # setView(lng = 2.80, lat = 46.80, zoom = 5) %>%
#   # addMarkers(data = CSLN_sf) %>%  # ATTENTION HYPER LONG !!!
#   addPolygons(data = bds,
#               label = ~Type, # En passant la souris
#               popup = ~Zone, # En cliquant sur l'icone
#               fill = TRUE,
#               fillColor = ~pal(Type),
#               fillOpacity = 0.8,
#               highlightOptions = highlightOptions(color = "white", weight = 2)) %>%
#    addRectangles(
#     lng1 = boxbds[1], lat1 = boxbds[2],
#     lng2 = boxbds[3], lat2 = boxbds[4],
#     color = "green",
#     fill = FALSE) %>%
#   addLegend(
#     title = "Zones",
#     pal = pal, values = bds$Type)
# map_leaflet
# saveWidget(map_leaflet, 'test_leaflet.html', selfcontained = TRUE)
```

## MAPS BUILDING

### Linear Two factor Quantile Regression

```{r}
#| label: rq_1
#| include : false # necessary to avoid fluffy doc

rdatain <- sprintf("%sCSLN_Mars_RQ_BDD.RData",wdwork)
load(rdatain)

spe <- 1#:nrow(species) # 1:CERED 2:CORVO 3:HEDDI 4:LIMBA 5:PERUL 6:SCRPL
# reponse<-reponse[1:3,] # 1:Biomass_gAFDWm2 2:Density_indm2 3:MSRtot 4:Itot  
answ <- 1#:nrow(reponse)
sai <- 1#:nrow(saison) # 1:Year 2:Winter 3:Summer
sdm <-1:nrow(sdmlist)
foldtxt<-"RQ Lineaire"; filtxt<-"Rq"
# Mars_SDM<-Mars_SDM %>% filter(!is.na(Lat) & !is.na(Lon) & !is.na(flow_mxd))

```

```{r}
#| label: rq_2
#| include : false # necessary to avoid fluffy doc

for (sp in spe) { # sp=1
  mlist<-vector(mode = "list", length = length(sdm))
  for (sdi in sdm) { # sdi=1
    x1t<-sdmlist$x1t[sdi]; x2t<-sdmlist$x2t[sdi]
    yt<-sdmlist$yt[sdi]
    x1l<-pred_red$Desc[which(pred_red$Var==x1t)]
    x2l<-pred_red$Desc[which(pred_red$Var==x2t)]
    yl<-reponse$rdescr[which(reponse$rvar==yt)]
    ylu<-reponse$runit[which(reponse$rvar==yt)]
    zt<-paste(species[sp,1],sdi,saison[sa,1],sep="")
    subtitre = sprintf("%s & %s", x1l,x2l)
    capt = sprintf("%s Quantile regression", sdmlist$SDM_desc[sdi])
    Mars_SDM_sf <- Mars_SDM %>% dplyr::select(c(Annee, !!zt)) # raster::select also exists!
    anMars<-unique(Mars_SDM_sf$Annee)
    for (ann in 1:length(anMars)) { # ann=1
      for (sai in sai) { # sai=1
        titreG = sprintf("SDM-NEO for %s in %s",species[sp,2],anMars[ann])
        Mars_SDM_sfi<-Mars_SDM_sf %>% filter(Mars_SDM_sf$Annee==anMars[ann])
        sfp<-ggplot(data = Mars_SDM_sfi) +
          geom_sf(aes(fill = .data[[zt]]), color = NA) + #
          labs(x="Latitude",y="Longitude",
               fill = sprintf("SDM-NEO\n%s\n(%s)\nTau=%s",yl,ylu,sdmlist$SDM_tau[sdi])) +
          theme(plot.margin = margin(0.05,0.05,0.05,0.05, "cm")) +
          scale_fill_distiller(palette = "Spectral") # + coord_sf(xlim = c(-88, -78), ylim = c(24.5, 33), expand = FALSE)
        assign(paste('sfp', ann, sep=''), 
               sfp + theme(legend.position="none",
                           axis.title.x=element_blank(), axis.text.x=element_blank(),axis.ticks.x=element_blank(),
                           axis.title.y=element_blank(), axis.text.y=element_blank(),axis.ticks.y=element_blank()) + labs(title=anMars[ann]))

        mlist[[ann]] <- eval(parse(text = paste('sfp', ann, sep='')))
        
        sfp <- sfp + labs(title=titreG, subtitle=subtitre, caption=capt)
        ggsave(sprintf("%s%s/%s/Map Detail/%s_%s_%s_SDM%s_%s_%s_%s.png",
                       wdgraphEx,species[sp,1],foldtxt,prgm,species[sp,1],filtxt,sdi,yt,paste(x1t,x2t,sep="&"),anMars[ann]),
                       plot = sfp, width = 10, height = 9)
      } #sai sfp1
    } #year
    titreG = sprintf("SDM-NEO for %s in %s (%s)",species[sp,2],yl,ylu)
    nncol = 5; nnrow=ceiling(length(anMars)/nncol)
    sfp <- ggarrange(plotlist=mlist, ncol=nncol, nrow=nnrow, common.legend = TRUE, legend="bottom")+
          theme(plot.margin = margin(0.05,0.05,0.05,0.05, "cm")) 
    sfp <- annotate_figure(sfp, top = text_grob(titreG, face = "bold", size = 14))+
          bgcolor("white"); #print(sfp)
    ggsave(sprintf("%s%s/%s/%s_%s_%s_SDM%s_%s_%s.png",
                   wdgraphEx,species[sp,1],foldtxt,prgm,species[sp,1],filtxt,sdi,yt,paste(x1t,x2t,sep="&")), plot = sfp, width = 10, height = 9)
  } # sdm
} #spe

```

### Non-linear One factor Gaussian Quantile Regression

```{r}
#| label: nlrq_1
#| include : false # necessary to avoid fluffy doc

rdatain <- sprintf("%sCSLN_Mars_NLRQ_BDD.RData",wdwork)
load(rdatain)

spe <- 1#:nrow(species) # 1:CERED 2:CORVO 3:HEDDI 4:LIMBA 5:PERUL 6:SCRPL
# reponse<-reponse[2:4,] # 1:Itot 2:MSRtot 3:Density_indm2 4:Biomass_gAFDWm2
answ <- 4#1:nrow(reponse)
sai <- 1#:nrow(saison) # 1:Year 2:Winter 3:Summer
sdm<-1:nrow(sdmlist)
foldtxt<-"RQ Nonlineaire"; filtxt<-"Nlrq"
# Mars_SDM<-Mars_SDM %>% filter(!is.na(Lat) & !is.na(Lon) & !is.na(flow_mxd))

```

```{r}
#| label: nlrq_2
#| include : false # necessary to avoid fluffy doc

for (sp in spe) { # sp=1
  mlist<-vector(mode = "list", length = length(sdm))
  for (sdi in sdm) { # sdi=1
    x1t<-sdmlist$x1t[sdi]; yt<-sdmlist$yt[sdi]
    x1l<-pred_red$Desc[which(pred_red$Var==x1t)];
    yl<-reponse$rdescr[which(reponse$rvar==yt)]; ylu<-reponse$runit[which(reponse$rvar==yt)]
    zt<-paste(species[sp,1],sdi,saison[sa,1],sep="")
    subtitre = sprintf("%s", x1l)
    capt = sprintf("%s Quantile regression", sdmlist$SDM_desc[sdi])
    Mars_SDM_sf <- Mars_SDM %>% dplyr::select(c(Annee, !!zt))
    anMars<-unique(Mars_SDM_sf$Annee)
    for (ann in 1:length(anMars)) { # ann=1
      for (sai in saison) { # sai=1
        titreG = sprintf("SDM-NEO for %s in %s",species[sp,2],anMars[ann])
        Mars_SDM_sfi<-Mars_SDM_sf %>% filter(Mars_SDM_sf$Annee==anMars[ann])
        sfp<-ggplot(data = Mars_SDM_sfi) +
          geom_sf(aes(fill = .data[[zt]]), color = NA) +
          labs(x="Latitude",y="Longitude",
               fill = sprintf("SDM-NEO\n%s\n(%s)\nTau=%s",yl,ylu,sdmlist$SDM_tau[sdi])) +
          theme(plot.margin = margin(0.05,0.05,0.05,0.05, "cm")) +
          scale_fill_distiller(palette = "Spectral") # + coord_sf(xlim = c(-88, -78), ylim = c(24.5, 33), expand = FALSE)
        assign(paste('sfp', ann, sep=''), 
               sfp + theme(legend.position="none",
                           axis.title.x=element_blank(), axis.text.x=element_blank(),axis.ticks.x=element_blank(),
                           axis.title.y=element_blank(), axis.text.y=element_blank(),axis.ticks.y=element_blank()) + labs(title=anMars[ann]))
        
        mlist[[ann]] <- eval(parse(text = paste('sfp', ann, sep='')))
        
        sfp <- sfp + labs(title=titreG, subtitle=subtitre, caption=capt)
        ggsave(sprintf("%s%s/%s/Map Detail/%s_%s_%s_SDM%s_%s_%s_%s.png",
                       wdgraphEx,species[sp,1],foldtxt,prgm,species[sp,1],filtxt,sdi,yt,x1t,anMars[ann]),plot = sfp, width = 10, height = 9)
      } #sai sfp1
    } #year
    titreG = sprintf("SDM-NEO for %s in %s (%s)\n%s",species[sp,2],yl,ylu,capt)
    nncol = 5; nnrow=ceiling(length(anMars)/nncol)
    sfp <- ggarrange(plotlist=mlist, ncol=nncol, nrow=nnrow, common.legend = TRUE, legend="bottom")+
      theme(plot.margin = margin(0.05,0.05,0.05,0.05, "cm")) 
    sfp <- annotate_figure(sfp, top = text_grob(titreG, face = "bold", size = 14))+
      bgcolor("white"); #print(sfp)
    ggsave(sprintf("%s%s/%s/%s_%s_%s_SDM%s_%s_%s.png",
                   wdgraphEx,species[sp,1],foldtxt,prgm,species[sp,1],filtxt,sdi,yt,x1t), plot = sfp, width = 10, height = 9)
  } # sdm
} #spe
```

### Non-linear Two factor Gaussian Quantile Regression

```{r}
#| label: nlrq2_1
#| include : false # necessary to avoid fluffy doc

rdatain <- sprintf("%sCSLN_Mars_NLRQ2d_BDD.RData",wdwork)
load(rdatain)

spe <- 1#:nrow(species) # 1:CERED 2:CORVO 3:HEDDI 4:LIMBA 5:PERUL 6:SCRPL
# reponse<-reponse[2:4,] # 1:Itot 2:MSRtot 3:Density_indm2 4:Biomass_gAFDWm2
answ <- 4#1:nrow(reponse)
sai <- 1#:nrow(saison) # 1:Year 2:Winter 3:Summer
sdm<-1:nrow(sdmlist)
foldtxt<-"RQ Nonlineaire"; filtxt<-"Nlrq2d"
# Mars_SDM2d<-Mars_SDM2d %>% filter(!is.na(Lat) & !is.na(Lon) & !is.na(flow_mxd))

```

```{r}
#| label: nlrq2_2
#| include : false # necessary to avoid fluffy doc

for (sp in spe) { # sp=1
  mlist<-vector(mode = "list", length = length(sdm))
  for (sdi in sdm) { # sdi=1
    x1t<-sdmlist$x1t[sdi]; x2t<-sdmlist$x2t[sdi]; yt<-sdmlist$yt[sdi]
    x1l<-pred_red$Desc[which(pred_red$Var==x1t)]; x2l<-pred_red$Desc[which(pred_red$Var==x2t)]
    yl<-reponse$rdescr[which(reponse$rvar==yt)]; ylu<-reponse$runit[which(reponse$rvar==yt)]
    zt<-paste(species[sp,1],sdi,saison[sa,1],sep="")
    subtitre = sprintf("%s & %s", x1l,x2l)
    capt = sprintf("%s Quantile regression", sdmlist$SDM_desc[sdi])
    Mars_SDM_sf <- Mars_SDM2d %>% dplyr::select(c(Annee, !!zt))
    anMars<-unique(Mars_SDM_sf$Annee)
    for (ann in 1:length(anMars)) { # ann=1
      for (sai in saison) { # sai=1
        titreG = sprintf("SDM-NEO for %s in %s",species[sp,2],anMars[ann])
        Mars_SDM_sfi<-Mars_SDM_sf %>% filter(Mars_SDM_sf$Annee==anMars[ann])
        sfp<-ggplot(data = Mars_SDM_sfi) +
          geom_sf(aes(fill = .data[[zt]]), color = NA) +
          labs(x="Latitude",y="Longitude",
               fill = sprintf("SDM-NEO\n%s\n(%s)\nTau=%s",yl,ylu,sdmlist$SDM_tau[sdi])) +
          theme(plot.margin = margin(0.05,0.05,0.05,0.05, "cm")) +
          scale_fill_distiller(palette = "Spectral") # + coord_sf(xlim = c(-88, -78), ylim = c(24.5, 33), expand = FALSE)
        assign(paste('sfp', ann, sep=''), 
               sfp + theme(legend.position="none",
                           axis.title.x=element_blank(), axis.text.x=element_blank(),axis.ticks.x=element_blank(),
                           axis.title.y=element_blank(), axis.text.y=element_blank(),axis.ticks.y=element_blank()) + labs(title=anMars[ann]))
        
        mlist[[ann]] <- eval(parse(text = paste('sfp', ann, sep='')))
        
        sfp <- sfp + labs(title=titreG, subtitle=subtitre, caption=capt)
        ggsave(sprintf("%s%s/%s/Map Detail/%s_%s_%s_SDM%s_%s_%s_%s.png",
                       wdgraphEx,species[sp,1],foldtxt,prgm,species[sp,1],filtxt,sdi,yt,paste(x1t,x2t,sep="&"),anMars[ann]),plot = sfp, width = 10, height = 9)
      } #sai sfp1
    } #year
    titreG = sprintf("SDM-NEO for %s in %s (%s)\n%s",species[sp,2],yl,ylu,capt)
    nncol = 5; nnrow=ceiling(length(anMars)/nncol)
    sfp <- ggarrange(plotlist=mlist, ncol=nncol, nrow=nnrow, common.legend = TRUE, legend="bottom")+
      theme(plot.margin = margin(0.05,0.05,0.05,0.05, "cm")) 
    sfp <- annotate_figure(sfp, top = text_grob(titreG, face = "bold", size = 14))+
      bgcolor("white"); #print(sfp)
    ggsave(sprintf("%s%s/%s/%s_%s_%s_SDM%s_%s_%s.png",
                   wdgraphEx,species[sp,1],foldtxt,prgm,species[sp,1],filtxt,sdi,yt,paste(x1t,x2t,sep="&")), plot = sfp, width = 10, height = 9)
  } # sdm
} #spe
```

```{r}
#| label: nlrq2_3
#| include : false # necessary to avoid fluffy doc

# wdshp <- paste("D:/Melting Potes/SIG/SIG sources externes/Communes 2015 5m/");
# fdc<-st_read(paste(wdshp,"communes-20150101-5m.shp",sep=""),quiet=TRUE) ; #str(cadre)
# bds<-st_crop(fdc,boxbds, crs = st_crs(4326)) # decoupe zone interet
# bds_points <- cbind(bds, st_coordinates(st_centroid(bds$geometry))) # def centroides pour placer les noms

# #________________________________________________________________
# # Ajout fond de carte des points de prelevement ----
# tmap_mode("view") #tmap_mode("plot") # ttm() toggle #
# tm_Bio<-tm_basemap("OpenStreetMap.HOT") + # OpenStreetMap.HOT .Mapnik .France - Stamen.Watercolor #https://leaflet-extras.github.io/leaflet-providers/preview/
#   # tm_logo(paste(wdlogos,"logo-MIE.png",sep=""), height = 2) +
#   # tm_logo(c(paste(wdlogos,"logoMP.png",sep=""),
#             # paste(wdlogos,"Logo Borea.png",sep="")), height = 2) +
#   tm_scale_bar(position = c("left", "bottom"), width = 0.15)+ #SCALE
#   tm_compass(position = c("left", "top"), size = 2)+          #NORTH COMPASS
#   # tm_shape(ES_Areas) +
#   # annotation_map_tile() +
#   tm_fill(col = "Zone", palette = "Spectral", alpha = 0.6) +
#   tm_borders("white", lwd = 1) +
#   tm_shape(Mars_SDM_sfi) +
#   tm_dots(size=0.001) +
#   tm_layout(legend.outside = TRUE)
# # tm_Bio # WARNING THIS OPERATION CAN LAST LONG
# tmlf_Bio<-tmap_leaflet(tm_Bio) # conversion to leaflet object, quicker??

```

# Final actions and save

Rdata are saved in `rdataout`. An excel file collects data in `r binderout` , with sheets for :

-   data : contains whatever

```{r}
#| label: finalsave_xls

```

```{r}
#| label: finalsave_r

beepr::beep(2)
```

# References {.unnumbered}
